### 2022-11-03 austins_analysis.rmd
First Commit Imported data, initialized Austin's .rmd and the main .rmd

# Analysis of the data by Austin
```{r}
# Import packages
library(tidyverse)
library(ggplot2)
```
```{r}
# Load the data
data <- read.csv("breast_cancer_diagnostic_data.csv")
```

```{r}
# Initial data exploration

# View the first 6 rows of the data
head(data)
# Available variables
names(data)
```

### 2022-11-03 main.rmd
First Commit Imported data, initialized Austin's .rmd and the main .rmd

# This r-markdown file is used to import the data and all of our own personal models as a final document

### 2022-11-12 main.rmd
Split Test and Training data with Group

@@
 # This r-markdown file is used to import the data and all of our own personal models as a final document
+
+```{r}
+library(tidyverse)
+library(ggplot2)
+library(dplyr)
+library(caret)
+```
+
+```{r}
+set.seed(420)
+data <- read.csv("breast_cancer_diagnostic_data.csv")
+data <- subset(data, select = -c(X))
+split <- sample(c(rep(0, 0.6 * nrow(data)), rep(1, 0.4 * nrow(data))))
+train <- data[split == 0, ]
+test <- data[split== 1, ]
+#write.csv(train,"training.csv", row.names = FALSE)
+#write.csv(test,"test.csv", row.names = FALSE)
+```

### 2022-12-01 ethan_analysis.Rmd
Add files via upload

---
title: "ethan_analysis"
author: "Ethan Powers"
date: '2022-11-12'
output: html_document
---

```{r}
#install.packages("tidymodels")
#install.packages("fs")
#install.packages("caret")
#install.packages("caTools")
#install.packages("leaps")
#install.packages("glmnet")
#install.packages("pls")
#install.packages("forecast")
```
```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}

setwd("~/Desktop/DSCI/DSCI445/Project")
library(tidyverse)
library(ggplot2)
library(dplyr)
library(caret)
library(leaps)
library(caTools)
library(tidymodels)
library(glmnet)
library(pls)
library(forecast)

set.seed(420)

dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
dataTot <- dataTemp %>% select_if(~ !any(is.na(.)))
testTot <- read.csv("test.csv")
trainingTot <- read.csv("training.csv")

```

#Dividing the data, test, and training data between malignant and benign
```{r}
#Malignant
tempM <- dataTot[dataTot$diagnosis == 'M', ]
dataM <- tempM[, !names(tempM) %in% c("diagnosis")]
splitM <- sample(c(rep(0, 0.6 * nrow(dataM)), rep(1, 0.4 * nrow(dataM))))
trainingM <- dataM[splitM == 0, ]
testM <- dataM[splitM == 1, ]

#Benign
tempB <- dataTot[dataTot$diagnosis == 'B', ]
dataB <- tempB[, !names(tempB) %in% c("diagnosis")]
splitB <- sample(c(rep(0, 0.6 * nrow(dataB)), rep(1, 0.4 * nrow(dataB))))
trainingB <- dataB[splitB == 0, ]
testB <- dataB[splitB == 1, ]
```

#Predicting the mean of the radius of the cancer cells between both Malignant & Benign combined and seperate
```{r}
#The prediction on the Total testing data set
reg_trainTot <- lm(radius_mean~., data = trainingTot)
pred_testTot <- predict(reg_trainTot, newdata = testTot)
error_testTot <- mean((testTot[, "radius_mean"] - pred_testTot)^2)

#The prediction on the Malignant testing data set
reg_trainM <- lm(radius_mean~., data = trainingM)
pred_testM <- predict(reg_trainM, newdata = testM)
error_testM <- mean((testM[, "radius_mean"] - pred_testM)^2)

#The prediction on the Benign testing data set
reg_trainB <- lm(radius_mean~., data = trainingB)
pred_testB <- predict(reg_trainB, newdata = testB)
error_testB <- mean((testB[, "radius_mean"] - pred_testB)^2)

print("Total test set error:")
print(error_testTot)
print('Malignant test set error:')
print(error_testM)
print('Benign test set error:')
print(error_testB)
```

#ridge regression
```{r}
#Total
training_matxTot <- model.matrix(radius_mean ~ . , data = trainingTot)[, -1 ]
test_matxTot <- model.matrix(radius_mean ~ . , data = testTot)[,-1]

cv_finalTot = cv.glmnet(training_matxTot , trainingTot$radius_mean, nfolds = 10, alpha=0)
lamTot = cv_finalTot$lambda.min

ridge_regTot = glmnet(training_matxTot , trainingTot$radius_mean, nfolds = 10, alpha = 0)
ridge_preTot = predict(ridge_regTot, s=lamTot, newx = test_matxTot)
error_test_ridTot <- mean((ridge_preTot - testTot$radius_mean)^2)
```
```{r}
#Malignant
training_matxM <- model.matrix(radius_mean ~ . , data = trainingM)[, -1 ]
test_matxM <- model.matrix(radius_mean ~ . , data = testM)[,-1]

cv_finalM = cv.glmnet(training_matxM , trainingM$radius_mean, nfolds = 10, alpha=0)
lamM = cv_finalM$lambda.min

ridge_regM = glmnet(training_matxM , trainingM$radius_mean, nfolds = 10, alpha = 0)
ridge_preM = predict(ridge_regM, s=lamM, newx = test_matxM)
error_test_ridM <- mean((ridge_preM - testM$radius_mean)^2)
```
```{r}
#Benign
training_matxB <- model.matrix(radius_mean ~ . , data = trainingB)[, -1 ]
test_matxB <- model.matrix(radius_mean ~ . , data = testB)[,-1]

cv_finalB = cv.glmnet(training_matxB , trainingB$radius_mean, nfolds = 10, alpha=0)
lamB = cv_finalB$lambda.min

ridge_regB = glmnet(training_matxB , trainingB$radius_mean, nfolds = 10, alpha = 0)
ridge_preB = predict(ridge_regB, s=lamB, newx = test_matxB)
error_test_ridB <- mean((ridge_preB - testB$radius_mean)^2)
```
```{r}
print("Total")
print(error_test_ridTot)
print("Malignant")
print(error_test_ridM)
print("Benign")
print(error_test_ridB)
```

#lasso
```{r}
#Total
lasso_setTot = glmnet(training_matxTot, trainingTot$radius_mean, nfolds = 10, alpha=1)
lasso_preTot = predict(lasso_setTot , s= lamTot, newx=test_matxTot)
error_test_lassoTot <- mean((lasso_preTot - testTot$radius_mean)^2)
```
```{r}
#Maligant
lasso_setM = glmnet(training_matxM, trainingM$radius_mean, nfolds = 10, alpha=1)
lasso_preM = predict(lasso_setM , s= lamM, newx=test_matxM)
error_test_lassoM <- mean((lasso_preM - testM$radius_mean)^2)
```
```{r}
#Benign
lasso_setB = glmnet(training_matxB, trainingB$radius_mean, nfolds = 10, alpha=1)
lasso_preB = predict(lasso_setB , s= lamB, newx=test_matxB)
error_test_lassoB <- mean((lasso_preB - testB$radius_mean)^2)
```
```{r}
print("Total")
print(error_test_lassoTot)
print("Malignant")
print(error_test_lassoM)
print("Benign")
print(error_test_lassoB)
```

#PCR
```{r}
#Total
trainingTot2 = dataTot %>% sample_frac(0.6)
testTot2 = dataTot %>% setdiff(trainingTot2)
pcr_fitTot2 = pcr(radius_mean ~ ., data = trainingTot2 , scale = TRUE, validation = "CV")
summary(pcr_fitTot2)
```
```{r}
#Total cont M is 31
x_trainingTot = model.matrix(radius_mean~., trainingTot2)[,-1]
x_testTot = model.matrix(radius_mean~., testTot2)[,-1]

y_trainingTot = trainingTot2 %>%
  select(radius_mean) %>%
  unlist() %>%
  as.numeric()

y_testTot = testTot2 %>%
  select(radius_mean) %>%
  unlist() %>%
  as.numeric()

pcr_predTot = predict(pcr_fitTot2, x_testTot, ncomp=31)
error_test_pcrTot <- mean((pcr_predTot-y_testTot)^2)
```

```{r}
#Malignant
trainingM2 = dataM %>% sample_frac(0.6)
testM2 = dataM %>% setdiff(trainingM2)
pcr_fitM2 = pcr(radius_mean ~ ., data = trainingM2 , scale = TRUE, validation = "CV")
summary(pcr_fitM2)
```
```{r}
#Maligant cont M is 30
x_trainingM = model.matrix(radius_mean~., trainingM2)[,-1]
x_testM = model.matrix(radius_mean~., testM2)[,-1]

y_trainingM = trainingM2 %>%
  select(radius_mean) %>%
  unlist() %>%
  as.numeric()

y_testM = testM2 %>%
  select(radius_mean) %>%
  unlist() %>%
  as.numeric()

pcr_predM = predict(pcr_fitM2, x_testM, ncomp=30)
error_test_pcrM <- mean((pcr_predM-y_testM)^2)
```

```{r}
#Belign
trainingB2 = dataB %>% sample_frac(0.6)
testB2 = dataB %>% setdiff(trainingB2)
pcr_fitB2 = pcr(radius_mean ~ ., data = trainingB2 , scale = TRUE, validation = "CV")
summary(pcr_fitB2)
```
```{r}
#Belign cont M is 29
x_trainingB = model.matrix(radius_mean~., trainingB2)[,-1]
x_testB = model.matrix(radius_mean~., testB2)[,-1]

y_trainingB = trainingB2 %>%
  select(radius_mean) %>%
  unlist() %>%
  as.numeric()

y_testB = testB2 %>%
  select(radius_mean) %>%
  unlist() %>%
  as.numeric()

pcr_predB = predict(pcr_fitB2, x_testB, ncomp=29)
error_test_pcrB <- mean((pcr_predB-y_testB)^2)
```
```{r}
print("Total")
print(error_test_pcrTot)
print("Malignant")
print(error_test_pcrM)
print("Benign")
print(error_test_pcrB)
```

#PLS
```{r}
#total
pls_fitTot = plsr(radius_mean~., data = trainingTot2, scale = TRUE, validation = "CV")
summary(pls_fitTot)
```
```{r}
#Total cont M is 30
pls_predTot = predict(pls_fitTot, x_testTot, ncomp = 30)
error_test_plsTot <- mean((pls_predTot - y_testTot)^2)
```

#Malignant
```{r}
pls_fitM = plsr(radius_mean~., data = trainingM2, scale = TRUE, validation = "CV")
summary(pls_fitM)
```
```{r}
#Malignant cont M is 30
pls_predM = predict(pls_fitM, x_testM, ncomp = 30)
error_test_plsM <- mean((pls_predM - y_testM)^2)
```

```{r}
#Benign
pls_fitB = plsr(radius_mean~., data = trainingB2, scale = TRUE, validation = "CV")
summary(pls_fitB)
```

```{r}
#Benign cont M is 26
pls_predB = predict(pls_fitB, x_testB, ncomp = 26)
error_test_plsB <- mean((pls_predB - y_testB)^2)
```
```{r}
print("Total")
print(error_test_plsTot)
print("Malignant")
print(error_test_plsM)
print("Benign")
print(error_test_plsB)
```

### 2022-12-01 austins_analysis.rmd

@@
 names(data)
+# Removing the ID column
+data <- data[,-1]
+
 ```
+
+```{r}
+# Check for missing values
+sum(is.na(data))
+head(data[which(is.na(data)), ])
+```
+
+```{r}
+# Looking for correlations between variables
+library(GGally)
+ggpairs(data, columns = 1:10)
+ggpairs(data, columns = c(1, 11:20))
+ggpairs(data, columns = c(1, 21:30))
+```
+
+Variables to note: radius_mean, perimeter_worst, area_worst, concave.points_worst
+
+```{r}
+#logistic regression
+data$diagnosis_binary <- ifelse(data$diagnosis == "M", 1, 0)
+logistic_model <- glm(diagnosis_binary ~ radius_mean + perimeter_worst + area_worst + concave.points_worst, data = data, family = binomial)
+summary(logistic_model)
+```
+
+```{r}
+ggplot(data, aes(x = radius_mean, y = texture_worst, color=diagnosis)) + geom_point()
+```
+
+```{r}
+```
+
+```{r}
+
+```

### 2022-12-01 paper.rmd

---
title: "Paper"
author: "Austin Lackey, Ethan Powers and Danny Laposata"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Sample Output
```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(caret)
set.seed(420)
data <- read.csv("breast_cancer_diagnostic_data.csv")
data <- subset(data, select = -c(X))
split <- sample(c(rep(0, 0.6 * nrow(data)), rep(1, 0.4 * nrow(data))))
train <- data[split == 0, ]
test <- data[split== 1, ]
#write.csv(train,"training.csv", row.names = FALSE)
#write.csv(test,"test.csv", row.names = FALSE)
```

```{r}
print(names(train))
```

### 2022-12-01 powerpoint.rmd

---
title: "Characterizing breast cancer cells"
author: Austin Lackey, Ethan Powers and Danny Laposata
date: Friday, November 9th, 2022
output: powerpoint_presentation
---

## The data

- We have a dataset of 569 breast cancer cells, some of which are cancerous and some of which are benign.
- We have 30 features for each cell, which are the measurements of the cell.
- We have a binary variable, `diagnosis`, which is `M` for malignant and `B` for benign.
- Add more stuff here...

## Sample Output
```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(caret)
set.seed(420)
data <- read.csv("breast_cancer_diagnostic_data.csv")
data <- subset(data, select = -c(X))
split <- sample(c(rep(0, 0.6 * nrow(data)), rep(1, 0.4 * nrow(data))))
train <- data[split == 0, ]
test <- data[split== 1, ]
#write.csv(train,"training.csv", row.names = FALSE)
#write.csv(test,"test.csv", row.names = FALSE)
```

```{r, echo=FALSE}
#histogram
print(ggplot(data, aes(x = diagnosis)) + geom_bar(aes(fill=diagnosis)) + labs(title = "Histogram of Diagnosis", x = "Diagnosis", y = "Count", fill="Diagnosis"))
```

## Plot
```{r, echo=FALSE}
#scatterplot
print(ggplot(data, aes(x = radius_mean, y = texture_mean, color = diagnosis)) + geom_point() + 
    labs(title = "Scatterplot of Radius Mean vs Texture Mean", x = "Radius Mean", y = "Texture Mean", color="Diagnosis"))
```

## Works Cited

- Insert link to kaggle dataset here

### 2022-12-01 Danny's Analysis.R
update

@@ -1,0 +1,153 @@
+### Packages
+
+library(tidyverse)
+library(GGally)
+library(ggplot2)
+library(tidyr)
+library(tidymodels)
+library(vip)
+library(caTools)
+library(xgboost)
+library(discrim)
+library(glmnet)
+
+
+### Gathering Data
+
+setwd("/Users/dlaposata/Documents/GitHub/project-7")
+train <- read.csv("training.csv")
+test <- read.csv("test.csv")
+
+train |>
+  select(-id) -> train
+
+test |>
+  select(-id) -> test
+
+train$diagnosis <- as.factor(train$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+
+
+### xGBoost Model
+boost_spec <- boost_tree(trees = 5000, tree_depth = 2, learn_rate = 0.01) |>
+  set_engine("xgboost") |>
+  set_mode("classification")
+
+boost_fit <- boost_spec |>
+  fit(diagnosis ~ ., data = train)
+
+boost_fit |>
+  augment(new_data = test) |>
+  accuracy(truth = diagnosis, estimate = .pred_class)
+
+boost_fit |>
+  augment(new_data = test) |>
+  conf_mat(truth = diagnosis, estimate = .pred_class) -> boost.cm
+
+autoplot(boost.cm, type = "heatmap")
+
+vip(boost_fit)
+
+### KNN 5 Model
+
+knn5_spec <- nearest_neighbor(mode = "classification", neighbors = 5)
+
+knn5_spec |>
+  fit(diagnosis ~ ., data = train) -> knn5.fit
+
+knn5.fit |>
+  augment(new_data = test) -> knn5.test_res
+
+knn5.test_res |>
+  conf_mat(truth = diagnosis, estimate = .pred_class) -> knn5.cm
+
+autoplot(knn5.cm, type = "heatmap")
+
+knn5.test_res |>
+  accuracy(truth = diagnosis, estimate = .pred_class) |>
+  mutate(error = 1 - .estimate) |>
+  pull(error)
+
+### Random Forest Model
+
+rf_spec <- rand_forest(mtry = sqrt(.cols())) |>
+  set_engine("randomForest", importance = TRUE) |>
+  set_mode("classification")
+
+rf_fit <- rf_spec |>
+  fit(diagnosis ~ ., data = train)
+
+vip(rf_fit)
+
+rf_fit |>
+  augment(new_data = test) |>
[…]

### 2022-12-02 austins_analysis.rmd
Logistic Regression Model

--- 2022-11-03 austins_analysis.rmd
@@
 library(ggplot2)
+library(GGally)
+library(knitr)
 ```
+
 ```{r}
@@
 data <- read.csv("breast_cancer_diagnostic_data.csv")
+training <- read.csv("training.csv")
+test <- read.csv("test.csv")
+
+# Data Cleaning
+# Removing the ID column
+data <- data[,-1]
+test <- test[,-1]
+training <- training[,-1]
+# Removing the X column from data
+data <- data[,1:(length(data)-1)]
+# Check for missing values
+print(paste(sum(is.na(data)), "missing values in the data"))
+# Available variables
+print(paste("There are", length(data), "variables available"))
+print(names(data))
 ```
@@
 ```{r}
-# Initial data exploration
+# Looking for correlations between variables
+g1 <- ggpairs(data, columns = 1:10)
+g2 <- ggpairs(data, columns = c(1, 11:20))
+g3 <- ggpairs(data, columns = c(1, 21:31))
+```
 
-# View the first 6 rows of the data
-head(data)
-# Available variables
-names(data)
+Variables to note: radius_mean, perimeter_worst, area_worst, concave.points_worst
+
+```{r}
+# Setting up a binary column for diagnosis
+data$diagnosis_binary <- ifelse(data$diagnosis == "M", 1, 0)
+training$diagnosis_binary <- ifelse(training$diagnosis == "M", 1, 0)
+test$diagnosis_binary <- ifelse(test$diagnosis == "M", 1, 0)
 ```
+
+```{r}
+#logistic regression using predictors: radius_mean, area_worst, concave.points_worst
+logistic_model <- glm(data = training, diagnosis_binary ~ radius_mean + area_worst + concave.points_worst, family = binomial)
+print(summary(logistic_model))
+logistic_model.prob <- predict(logistic_model, test, type = "response")
+logistic_model.pred <- ifelse(logistic_model.prob > 0.5, 1, 0)
+logistic_model.pos <- logistic_model.prob[logistic_model.pred == 1]
+logistic_model.neg <- logistic_model.prob[logistic_model.pred == 0]
+logistic_model_error <- test$diagnosis_binary != as.integer(logistic_model.pred)
+ggplot() + geom_point(aes(x = test$radius_mean, y=test$perimeter_worst, color = test$diagnosis))
+print(ggplot() + geom_point(aes(x = test$texture_mean, y=test$perimeter_mean, color = !logistic_model_error)) + labs(title = "Logistic Regression", x="texture_mean", y="perimeter_mean", color="Correct Prediction"))
+```
+
+```{r}
+mean(logistic_model_error)
+```
+
+```{r}
+ggplot(data, aes(x = radius_mean, y = texture_worst, color=diagnosis)) + geom_point()
+```
+
+```{r}
+```
+
+```{r}
+
+```

### 2022-12-04 austins_analysis.rmd
updated

--- 2022-11-03 austins_analysis.rmd
@@
 library(ggplot2)
+library(GGally)
+library(knitr)
+library(boot)
+library(caret)
 ```
+
 ```{r}
@@
 data <- read.csv("breast_cancer_diagnostic_data.csv")
+training <- read.csv("training.csv")
+test <- read.csv("test.csv")
+
+# Data Cleaning
+# Removing the ID column
+data <- data[,-1]
+test <- test[,-1]
+training <- training[,-1]
+data$diagnosis <- as.factor(data$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+training$diagnosis <- as.factor(training$diagnosis)
+# Removing the X column from data
+data <- data[,1:(length(data)-1)]
+# Check for missing values
+print(paste(sum(is.na(data)), "missing values in the data"))
+# Available variables
+print(paste("There are", length(data), "variables available"))
+print(names(data))
 ```
@@
 ```{r}
-# Initial data exploration
+# Looking for correlations between variables
+g1 <- ggpairs(data, columns = 1:10)
+g2 <- ggpairs(data, columns = c(1, 11:20))
+g3 <- ggpairs(data, columns = c(1, 21:31))
+```
 
-# View the first 6 rows of the data
-head(data)
-# Available variables
-names(data)
+Variables to note: radius_mean, perimeter_worst, area_worst, concave.points_worst
+
+```{r}
+# Setting up a binary column for diagnosis
+dataBinary <- data
+trainingBinary <- training
+testBinary <- test
+dataBinary$diagnosis_binary <- ifelse(data$diagnosis == "M", 1, 0)
+trainingBinary$diagnosis_binary <- ifelse(training$diagnosis == "M", 1, 0)
+testBinary$diagnosis_binary <- ifelse(test$diagnosis == "M", 1, 0)
 ```
+
+```{r}
+train_control <- trainControl(method = "cv", number = 10, preProcOptions = list(thresh=0.99), classProbs=TRUE, summaryFunction = twoClassSummary)
+```
+
+Random Forrest
+```{r}
+#Random Forrest
+random_forrest <- train(diagnosis ~ ., data = training, method = "ranger", metric="ROC", trControl = train_control, preProcess=c("center", "scale"))
+rf_pred <- predict(random_forrest, test)
+rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
+print(rf_confusion)
+```
+
+KNN
+```{r}
+#KNN
+knn <- train(diagnosis ~ ., data = training, method = "knn", metric="ROC", trControl = train_control)
+knn_pred <- predict(knn, test)
+knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
+print(knn_confusion)
+```
+
+Radial Support Vector Machine
+```{r}
+#Radial Support Vector Machine
+svm <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric="ROC")
+svm_pred <- predict(svm, test)
+svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive="M")
+print(svm_confusion)
+```
+
+```{r}
+#Logistic Regression
+logistic <- train(diagnosis ~ ., data = training, method = "glm", trControl = train_control, metric="ROC")
+logistic_pred <- predict(logistic, test)
+logistic_confusion <- confusionMatrix(logistic_pred, test$diagnosis, positive="M")
+print(logistic_confusion)
+logistic$finalModel
+ggplot(varImp(logistic))
+imp <- varImp(logistic)
+top10 <- rownames(data.frame(imp[1]))[1:10]
+```
+
+```{r}
+logistic2 <- train(diagnosis ~ top10, data = training, method = "glm", trControl = train_control, metric="ROC")
+
+```
+
+```{r}
+```

### 2022-12-05 plots.R
plots

@@ -1,0 +1,23 @@
+library(tidyverse)
+library(GGally)
+library(ggplot2)
+
+setwd("/Users/dlaposata/Documents/GitHub/project-7")
+train <- read.csv("training.csv")
+test <- read.csv("test.csv")
+
+train |>
+  select(-id) -> train
+
+test |>
+  select(-id) -> test
+
+train$diagnosis <- as.factor(train$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+
+train[,c(1,2,12,22)] |>
+  ggpairs()
+
+ggplot(train) +
+  geom_point(aes(x = concavity_mean, y = radius_mean, col = diagnosis))+
+  labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor")

### 2022-12-05 austins_analysis.rmd
updated powerpoint

@@
 print(rf_confusion)
+
 ```
@@
 print(knn_confusion)
+ggplot(varImp(knn))
+varsKNN <- rownames(data.frame(varImp(knn)[1]))[1:10]
+```
 ```
@@
 print(svm_confusion)
+ggplot(varImp(svm))
+varsSVM <- rownames(data.frame(varImp(svm)[1]))[1:10]
 ```
@@
 ggplot(varImp(logistic))
-imp <- varImp(logistic)
-top10 <- rownames(data.frame(imp[1]))[1:10]
+varsLOGISTIC <- rownames(data.frame(varImp(logistic)[1]))[1:10]
 ```
@@
 ```{r}
-logistic2 <- train(diagnosis ~ top10, data = training, method = "glm", trControl = train_control, metric="ROC")
+logistic2 <- train(diagnosis ~ area_mean + area_worst + concave.points_mean + texture_mean + radius_worst + concavity_se + fractal_dimension_se, data = training, method = "glm", trControl = train_control, metric="ROC")
+logistic2_pred <- predict(logistic2, test)
+logistic2_confusion <- confusionMatrix(logistic2_pred, test$diagnosis, positive="M")
+print(logistic2_confusion)
+```
+
+```{r}
 

### 2022-12-05 powerpoint.rmd
updated powerpoint

@@
 
-```{r, echo=FALSE}
-#histogram
-print(ggplot(data, aes(x = diagnosis)) + geom_bar(aes(fill=diagnosis)) + labs(title = "Histogram of Diagnosis", x = "Diagnosis", y = "Count", fill="Diagnosis"))
-```
+# Exploratory Data Analysis
 
-## Plot
-```{r, echo=FALSE}
-#scatterplot
-print(ggplot(data, aes(x = radius_mean, y = texture_mean, color = diagnosis)) + geom_point() + 
-    labs(title = "Scatterplot of Radius Mean vs Texture Mean", x = "Radius Mean", y = "Texture Mean", color="Diagnosis"))
-```
+# Model Building
+
+# Testing Accuracy
 

### 2022-12-05 powerpoint.rmd

@@
 ---
-title: "Characterizing breast cancer cells"
+title: "Breast Cancer Diagnosis Classification Analysis"
 author: Austin Lackey, Ethan Powers and Danny Laposata
-date: Friday, November 9th, 2022
-output: powerpoint_presentation
+date: "December 9th, 2022"
+output: 
+  powerpoint_presentation: 
+    reference_doc: template.pptx
+
 ---
-
-## The data
-
-- We have a dataset of 569 breast cancer cells, some of which are cancerous and some of which are benign.
-- We have 30 features for each cell, which are the measurements of the cell.
-- We have a binary variable, `diagnosis`, which is `M` for malignant and `B` for benign.
-- Add more stuff here...
-
-## Sample Output
-```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
-library(tidyverse)
-library(ggplot2)
-library(dplyr)
-library(caret)
-set.seed(420)
-data <- read.csv("breast_cancer_diagnostic_data.csv")
-data <- subset(data, select = -c(X))
-split <- sample(c(rep(0, 0.6 * nrow(data)), rep(1, 0.4 * nrow(data))))
-train <- data[split == 0, ]
-test <- data[split== 1, ]
-#write.csv(train,"training.csv", row.names = FALSE)
-#write.csv(test,"test.csv", row.names = FALSE)
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
 ```
 
-# Exploratory Data Analysis
+## Background
 
-# Model Building
+- Breast Cancer Diagnosis Data
+- Predict Diagnosis of Tumor based on features
+- Reduce Type II Error (False-Negatives)
+- Feature Selection
 
-# Testing Accuracy
+## Data
+
+- Data collected from Kaggle
+- Response: Diagnosis (Benign/Malignant)
+- Mean, SE, and "Worst" measurements of 10 variables
+- Training/Test Split (60/40)
+
+## EDA
+
+- Different measurements of each variable are correlated
+
+- Relationship with Diagnosis visible for each measurement
+
+```{r, fig.height=3, fig.width=4, fig.align='right', fig.pos='H'}
+source("Danny's Analysis.R", echo = F, local = T)
+
+train[,c(1,2,12,22)] |>
+  ggpairs(title = "Correlation b/w Radius Measurements")
+```
+
+## Classification Analysis: Austin
+
+## Classification Analysis: Danny
+
+## Regression Analysis: Ethan
+
+## Model Assesments
 

### 2022-12-07 plots.R
Add files via upload
Ethan correlation graph

@@ -2,9 +2,16 @@ library(tidyverse)
 library(GGally)
 library(ggplot2)
 
-setwd("/Users/dlaposata/Documents/GitHub/project-7")
+
+#setwd("/Users/dlaposata/Documents/GitHub/project-7")
+data <- read.csv("breast_cancer_diagnostic_data.csv")
 train <- read.csv("training.csv")
 test <- read.csv("test.csv")
+
+dataE <- data
+corr_df <- dataE[,c('area_worst', 'area_mean', 'concave.points_mean', 'radius_worst')]
+#head(corr_df)
+ggpairs(corr_df)
 
 train |>
   select(-id) -> train

### 2022-12-07 austins_analysis.rmd
updated analysis

--- 2022-12-04 austins_analysis.rmd
@@
 # Analysis of the data by Austin
-```{r}
+```{r package_load}
 # Import packages
@@
 library(caret)
+library(randomForest)
+library(e1071)
 ```
 
-```{r}
+```{r data_load}
 # Load the data
@@
 
-```{r}
+```{r ggpairs}
 # Looking for correlations between variables
@@
 
+```{r ggpairs_plot}
+# print(g1)
+# print(g2)
+# print(g3)
+```
+
 Variables to note: radius_mean, perimeter_worst, area_worst, concave.points_worst
 
-```{r}
+```{r data_cleaning}
 # Setting up a binary column for diagnosis
@@
 testBinary$diagnosis_binary <- ifelse(test$diagnosis == "M", 1, 0)
+testBinary <- subset(testBinary, select = -diagnosis)
+trainingBinary <- subset(trainingBinary, select = -diagnosis)
 ```
 
-```{r}
+Logistic Regression
+
+```{r logistic_regression}
+# Logistic Regression without cross validation
+logistic_regression <- glm(diagnosis_binary ~ ., data = trainingBinary, family = binomial)
+logistic_pred <- predict(logistic_regression, testBinary, type="response")
+logistic_pred <- ifelse(logistic_pred > 0.5, "M", "B")
+logistic_pred <- data.frame(logistic_pred)$logistic_pred
+logistic_confusion <- confusionMatrix(logistic_pred, as.character(test$diagnosis), positive = "M")
+```
+
+```{r train_setup}
 train_control <- trainControl(method = "cv", number = 10, preProcOptions = list(thresh=0.99), classProbs=TRUE, summaryFunction = twoClassSummary)
@@
 Random Forrest
-```{r}
+```{r random_forrest}
 #Random Forrest
@@
 
+```{r random_forrest_plot}
+#Plotting the false positives and true positives for the random forest model
+false_pred <- !(test$diagnosis != rf_pred)
+false_negative <- ifelse((test$diagnosis == "M") & (rf_pred == "B"), "FalseNeg", "Okay")
+ggplot() + geom_point(aes(x=test$texture_mean, y=test$radius_mean, color=false_negative), size=2) + labs(title = "Random Forrest Model False Negatives")
+```
+
+Random Forrest 2
+
+```{r random_forrest_2}
+x <- seq(1, 100, 1)
+tuning_grid <- expand.grid(trees = x, rmse=NA)
+for (i in seq_len(nrow(tuning_grid))) {
+    # rf_fit <- train(diagnosis ~ ., 
+    #     method = "ranger",
+    #     metric="ROC",
+    #     trControl = train_control,
+    #     data = training, 
+    #     num.trees = tuning_grid$trees[i])
+    rf_fit <- randomForest(diagnosis ~ ., data = training, ntree = tuning_grid$trees[i])
+    rf_pred <- predict(rf_fit, test)
+    rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
+    tuning_grid$rmse[i] <- tibble(rf_confusion$table)[[2,1]][1]/sum(tibble(rf_confusion$table))
+}
+```
+
+```{r random_forrest_2_plot}
+ggplot() + geom_line(aes(x=x, y=tuning_grid$rmse))
+```
+
 KNN
-```{r}
+```{r knn}
 #KNN
@@
 print(knn_confusion)
+varsKNN <- rownames(data.frame(varImp(knn)[1]))[1:10]
+```
+
+```{r knn_plot}
+print(ggplot(varImp(knn)) + labs(title = "KNN Variable Importance"))
+```
+
+```{r knn_plot2}
+#Plotting the false positives and true positives for the knn model
+false_pred <- !(test$diagnosis != knn_pred)
+false_negative <- ifelse((test$diagnosis == "M") & (knn_pred == "B"), "FalseNeg", "Okay")
+ggplot() + geom_point(aes(x=test$texture_mean, y=test$radius_mean, color=false_negative), size=2) + labs(title = "KNN Model False Negatives")
 ```
@@
 Radial Support Vector Machine
-```{r}
+```{r svm}
 #Radial Support Vector Machine
@@
 print(svm_confusion)
+varsSVM <- rownames(data.frame(varImp(svm)[1]))[1:10]
 ```
 
-```{r}
+```{r svm_plot}
+print(ggplot(varImp(svm)))
+```
+
+```{r svm_plot}
+#Plotting the false positives and true positives for the logistic regression model
+false_pred <- !(test$diagnosis != svm_pred)
+false_negative <- ifelse((test$diagnosis == "M") & (svm_pred == "B"), "FalseNeg", "Okay")
+ggplot() + geom_point(aes(x=test$texture_mean, y=test$radius_mean, color=false_negative), size=2) + labs(title = "SVM Model False Negatives")
+```
+
+Logistic Regression
+```{r logistic}
 #Logistic Regression
@@
 logistic$finalModel
-ggplot(varImp(logistic))
-imp <- varImp(logistic)
-top10 <- rownames(data.frame(imp[1]))[1:10]
+varsLOGISTIC <- rownames(data.frame(varImp(logistic)[1]))[1:10]
 ```
@@
 ```{r}
-logistic2 <- train(diagnosis ~ top10, data = training, method = "glm", trControl = train_control, metric="ROC")
+#Plotting the false positives and true positives for the logistic regression model
+false_pred <- !(test$diagnosis != logistic_pred)
+false_negative <- ifelse((test$diagnosis == "M") & (logistic_pred == "B"), "FalseNeg", "Okay")
+ggplot() + geom_point(aes(x=test$texture_mean, y=test$radius_mean, color=false_negative), size=2) + labs(title = "Logistic Regression Reduced Model False Negatives")
+```
 
+```{r logistic_plot}
+print(ggplot(varImp(logistic)) + labs(title = "Logistic Variable Importance"))
+```
+
+```{r logistic_2}
+logistic2 <- train(diagnosis ~ area_mean + area_worst + concave.points_mean + texture_mean + radius_worst + concavity_se + fractal_dimension_se + concavity_worst + concave.points_worst, data = training, method = "glm", trControl = train_control, metric="ROC")
+logistic2_pred <- predict(logistic2, test)
+logistic2_confusion <- confusionMatrix(logistic2_pred, test$diagnosis, positive="M")
+print(logistic2_confusion)
+```
+
+```{r}
+#Plotting the false positives and true positives for the logistic regression 2
+false_pred <- !(test$diagnosis != logistic2_pred)
+false_negative <- ifelse((test$diagnosis == "M") & (logistic2_pred == "B"), "FalseNeg", "Okay")
+ggplot() + geom_point(aes(x=test$texture_mean, y=test$radius_mean, color=false_negative), size=2) + labs(title = "Logistic Regression Reduced Model False Negatives")
+```
+
+```{r logistic_2_plot}
 ```

### 2022-12-07 austins_temp_powerpoint.rmd
updated analysis

--- 2022-12-05 powerpoint.rmd
@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+knitr::opts_chunk$set(fig.width=12, fig.height=8) 
 ```
 
-## Background
-
-- Breast Cancer Diagnosis Data
-- Predict Diagnosis of Tumor based on features
-- Reduce Type II Error (False-Negatives)
-- Feature Selection
-
-## Data
-
-- Data collected from Kaggle
-- Response: Diagnosis (Benign/Malignant)
-- Mean, SE, and "Worst" measurements of 10 variables
-- Training/Test Split (60/40)
-
-## EDA
-
-- Different measurements of each variable are correlated
-
-- Relationship with Diagnosis visible for each measurement
-
-```{r, fig.height=3, fig.width=4, fig.align='right', fig.pos='H'}
-source("Danny's Analysis.R", echo = F, local = T)
-
-train[,c(1,2,12,22)] |>
-  ggpairs(title = "Correlation b/w Radius Measurements")
-```
-
-## Classification Analysis: Austin
-
-## Classification Analysis: Danny
-
-## Regression Analysis: Ethan
-
-## Model Assesments
-
-## Works Cited
-
-- Insert link to kaggle dataset here
+# Breast Cancer Diagnosis Classification Analysis

### 2022-12-08 ethan_temp_powerpoint.rmd
Add files via upload

---
title: "Breast Cancer Diagnosis Classification Analysis"
author: Austin Lackey, Ethan Powers and Danny Laposata
date: "December 9th, 2022"
output: 
  powerpoint_presentation: 
    reference_doc: template.pptx

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(caret)
library(leaps)
library(caTools)
library(tidymodels)
library(glmnet)
library(pls)
library(forecast)
library(GGally)

set.seed(420)

dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
data <- dataTemp %>% select_if(~ !any(is.na(.)))
test <- read.csv("test.csv")
training <- read.csv("training.csv")
```

## Background

- Breast Cancer Diagnosis Data
- Predict Diagnosis of Tumor based on features
- Reduce Type II Error (False-Negatives)
- Feature Selection

## Data

- Data collected from "Breast Cancer Wisconsin (Diagnostic) Data Set" from Kaggle
- Response: Diagnosis (Benign/Malignant)
- Predictors: 10 variables divided between  Mean, SE, and "Worst"
- Training/Test Split (60/40)

## EDA

- Different measurements of each variable are correlated

- Relationship with Diagnosis visible for each measurement

```{r, fig.height=3, fig.width=4, fig.align='right', fig.pos='H'}
#source("Danny's Analysis.R", echo = F, local = T)

#train[,c(1,2,12,22)] |>
#  ggpairs(title = "Correlation b/w Radius Measurements")
```

## Classification Analysis: Austin

## Classification Analysis: Danny

## Regression Analysis: Ethan
```{r}
dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
dataTot <- dataTemp %>% select_if(~ !any(is.na(.)))
testTot <- read.csv("test.csv")
trainingTot <- read.csv("training.csv")

dataE <- data
corr_df <- dataE[,c('area_worst', 'area_mean', 'concave.points_mean', 'radius_worst')]
ggpairs(corr_df)
```
-Test

## Model Assesments

## Works Cited

- Breast Cancer Wisconsin (Diagnostic) Data Set: https://www.kaggle.com/datasets/uciml/breast-cancer-wisconsin-data?datasetId=180&sortBy=voteCount

### 2022-12-08 refrenceAustinPres.rmd
ref pres

---
title: "Predicting Covid-19 Data"
author: "Austin Lackey, Chien Lin and Jin Peng"
date: "Tuesday, December 6th, 2022"
output:
  powerpoint_presentation: default
  beamer_presentation: default
---

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# useful libraries
library(tidyverse)
library(pracma)
library(minpack.lm)
# read covid data
data <- read_csv("owid-covid-data.csv")
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```

## The Monte Carlo Study:

**Mathematical prediction of the time evolution of the COVID-19 pandemic in Italy by a Gauss error function and Monte Carlo simulations**

<br>

*Author: Ignazio Ciufolini, Antonio Paolozzi*

## Introduction : Ciufolini and Paolozzi's Study

* The time evolution of the number of cumulative diagnosed positive cases and fatalities in China approximates the distribution of a gauss error function: $a+b*erf(c*x-d)$.
* Performed the function on Italy data to approximate the number of cases.
* Performed 150 Monte Carlo simulations to predict the day in which the peak of the number of daily cases in Italy occurs.

# China Simulation

## Available Official Data (Positive Cases)

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold", out.width="50%"}
# extract China data of interest
china <- data |>
  filter(location == "China") |>
  filter(date >= "2020-01-22", date <= "2020-03-28") |>
  select(c(date, total_cases, total_deaths)) |>
  mutate(days = 1:n())

# recreate plots
positive <- ggplot(data = china) +
  geom_point(aes(x = date, y = total_cases)) +
  labs(x = "Date", y = "China cumulative number of cases")
positive
```

## Available Official Data (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold", out.width="50%"}
ggplot(data = china) +
  geom_point(aes(x = date, y = total_deaths)) +
  labs(x = "Date", y = "China cumulative number of fatalities") +
  ylim(0,5000)
```

## Fit by the Gauss Error Function

Apply an $nls$ function to determine the nonlinear (weighted) least-squares estimates of the four parameters $a,b,c,d$ of a Gauss error function model: $a+b*erf(c*x-d)$.

```{r, include=FALSE, message=FALSE, warning=FALSE}
gauss_fn <- function(x,a,b,c,d) {a+b*erf(c*x-d)}
```

```{r, message=FALSE, warning=FALSE}
model <- nls(total_cases ~ gauss_fn(days,a,b,c,d), 
             data=china, 
             start=list(a=10000,b=10000,c=0.1,d=1))
```

## Fit by the Gauss Error Function and Peak Day (Positive Cases)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
try <- china$total_cases
model_try <- nls(try ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=2200,b=2200,c=0.1,d=1))
est_try <- summary(model_try)$parameters
est_try <- est_try[, 1]
names(est_try) <- NULL
fn_try <- function(x, ests) {ests[1]  + ests[2] * erf(ests[3] * x - ests[4])}

# find the second derivative of fitted function
# source: https://proofwiki.org/wiki/Derivative_of_Error_Function
erf_deriv <- function(x) {2/sqrt(pi)*exp(-x^2)}
fn_try_deriv <- bquote(.(est_try[2]) * 2/sqrt(pi) * exp(-(.(est_try[3]) * x - .(est_try[4]))^2))
fn_try_deriv <- D(fn_try_deriv,'x')
x=seq(1,67,length.out=6700)
tmp <- eval(fn_try_deriv)
peak1 <- min(x[tmp < 0])
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
gauss_fn <- function(x,a,b,c,d) {a+b*erf(c*x-d)}
model1 <- nls(total_cases ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=10000,b=10000,c=0.1,d=1))
# summary(model1)$parameters
fn_model1 <- function(x) {4.017e+04 + 4.047e+04 * erf(9.284e-02 * x - 1.732e+00)}
ggplot(data = china) +
  geom_point(aes(x = days, y = total_cases)) +
  labs(x = "Days", y = "China cumulative number of cases") +
  stat_function(fun=fn_model1, color="#d33636", size=1.5) +
  geom_vline(xintercept = peak1, color="#d33636", size=1.5, linetype="dashed")
```

## Fit by the Gauss Error Function and Peak Day (Fatalities)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
try <- china$total_deaths
model_try <- nls(try ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=2200,b=2200,c=0.1,d=1))
est_try <- summary(model_try)$parameters
est_try <- est_try[, 1]
names(est_try) <- NULL
fn_try <- function(x, ests) {ests[1]  + ests[2] * erf(ests[3] * x - ests[4])}

# find the second derivative of fitted function
# source: https://proofwiki.org/wiki/Derivative_of_Error_Function
erf_deriv <- function(x) {2/sqrt(pi)*exp(-x^2)}
fn_try_deriv <- bquote(.(est_try[2]) * 2/sqrt(pi) * exp(-(.(est_try[3]) * x - .(est_try[4]))^2))
fn_try_deriv <- D(fn_try_deriv,'x')
x=seq(1,67,length.out=6700)
tmp <- eval(fn_try_deriv)
peak2 <- min(x[tmp < 0])
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
model2 <- nls(total_deaths ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=10000,b=10000,c=0.1,d=1))
#summary(model2)$parameters
fn_model2 <- function(x) {1.571e+03 + 1.658e+03 * erf(6.194e-02 * x - 1.533e+00)}
ggplot(data = china) +
  geom_point(aes(x = days, y = total_deaths)) +
  labs(x = "Days", y = "China cumulative number of fatalities") +
  ylim(0,5000) +
  stat_function(fun=fn_model2, color="#1b95ff", size=1.5) +
  geom_vline(xintercept = peak2, color="#1b95ff", size=1.5, linetype="dashed")
```

# Italy Simulation

## Initial Data for Cases in Italy

```{r, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(400)
# extract Italy data
italy <- data %>% 
  filter(location=="Italy") %>% 
  filter(date <= "2020-04-21") %>%
  filter(date >= "2020-02-15") %>%
  select(date, total_cases, total_deaths)
  #ref: %>% mutate(days = c(1:67))
italy[45:67,] <- NA
italy <- italy %>% mutate(days = c(1:67))

#plot
ggplot(italy, aes(x = days, y = total_cases)) + 
    geom_point() + 
    labs(title = "Total Cases in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Cases")
```

## Initial Data for Fatalities in Italy

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#plot
ggplot(italy, aes(x = days, y = total_deaths)) +
    geom_point() + 
    labs(title = "Total Deaths in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Deaths")
```

## Fitting Gauss Error Function to Positive Cases in Italy

```{r, echo=FALSE, message=FALSE, warning=FALSE}
x <- 1:67

c <- summary(nlsLM(total_cases~gauss_fn(days,a,b,c,d), data=italy, start=list(a=50000,b=50000,c=.01,d=1)))$parameters
d <- summary(nlsLM(total_deaths~gauss_fn(days,a,b,c,d), data=italy, start=list(a=10000,b=10000,c=.01,d=1)))$parameters

cf_italy <- function(x) c[1]+c[2]*erf(c[3]*x-c[4])
df_italy <- function(x) d[1]+d[2]*erf(d[3]*x-d[4])

#plot
ggplot(italy, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cf_italy, color="#d33636", size=1.5, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Cases in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Cases")
```

## Fitting Gauss Error Function to Fatalities in Italy

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#plot
ggplot(italy, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=df_italy, color="#1b95ff", size=1.5, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Deaths in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Deaths")
```

## Overview of the Monte Carlo Simulation

* Create a random matrix : m X n
  * m = number of random outcomes (*chosen to be 150*)
  * n = number of observed days (*n = 1,...,j*)
* Each # in the matrix is a Gaussian distribution with mean = 1 and sigma = 0.1
  * Multiply each column (*j*) by the # of total cases (or deaths) that corresponds to each day *j*
* Each of the 150 simulations were fitted with the Gauss error function
* Then determined the date of the flex for each simulation

## Using MonteCarlo Simulation to predict the day of flex

- The mean of the day of flex for cases is 40.59 days and the standard deviation is 5.222832 days.
- The mean of the day of flex for fatalities is 43.46 days and the standard deviation is 5.228381 days.

## MonteCarlo Simulation Data (Cases) (150 trials)

Each red dot corresponds to the day of occurrence of the flex obtained with each of the 150 Monte Carlo simulations

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
n <- 44
m <- 150
sigma <- 0.1
mu <- 1
gau_sample <- matrix(rnorm(n*m, mean=mu, sd=sigma), nrow = m, ncol = n)

italy_mod <- italy[1:44,]

model_ca <- c()
for (i in 1:n) {
    model_ca <- c(model_ca, italy_mod$total_cases[i]*gau_sample[,i])
}
model_ca <- matrix(model_ca, nrow = m, ncol = n)

model_fa <- c()
for (i in 1:n) {
    model_fa <- c(model_fa, italy_mod$total_deaths[i]*gau_sample[,i])
}
model_fa <- matrix(model_fa, nrow = m, ncol = n)
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# cases
flex_ca <- rep(NA, m)
for (i in 1:m) {
  per_m <- model_ca[i, ]
  model_try <-nlsLM(per_m ~ gauss_fn(days, a, b, c, d),
                  data = italy_mod,
                  start = list(a=50000,b=50000,c=.01,d=1))
  get_coef <- summary(model_try)$parameters
  get_coef <- get_coef[, 1]
  names(get_coef) <- NULL
  # Derivative of Gauss Error Function: 2 / sqrt(pi) * exp(-x ^ 2)
  fn_try_deriv <-
    bquote(.(get_coef[2]) * 2 / sqrt(pi) * exp(-(.(get_coef[3]) * x - .(get_coef[4])) ^ 2))
  fn_try_deriv <- D(fn_try_deriv, 'x')
  x = seq(1, 67, length.out = 6700)
  tmp <- eval(fn_try_deriv)
  flex_ca[i] <- min(x[tmp < 0])
}
mean(flex_ca)
sd(flex_ca)
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
    geom_point(aes(x=1:m, y=flex_ca), color="#d33636") +
    geom_hline(yintercept=mean(flex_ca), color="#d33636", linetype="dashed", size=1.5) + 
    labs(title = "MonteCarlo Simulation for Day of Flex (Cases)", x = "Trial", y = "Day of Flex")
```

## Applying the day of flex indicator to the plot (Cases)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#plot(cs)
ggplot(italy, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cf_italy, color="#d33636", size=1.5, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Cases in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Cases") + 
  geom_vline(xintercept = mean(flex_ca), color="#d33636", size=1, linetype="dashed")
```

## MonteCarlo Simulation Data (Fatalities) (150 trials)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# fatalities
flex_fa <- rep(NA, m)
for (i in 1:m) {
  per_m <- model_fa[i, ]
  model_try <-nlsLM(per_m ~ gauss_fn(days, a, b, c, d),
                  data = italy_mod,
                  start = list(a=15000,b=15000,c=.01,d=1))
  get_coef <- summary(model_try)$parameters
  get_coef <- get_coef[, 1]
  names(get_coef) <- NULL
  # Derivative of Gauss Error Function: 2 / sqrt(pi) * exp(-x ^ 2)
  fn_try_deriv <-
    bquote(.(get_coef[2]) * 2 / sqrt(pi) * exp(-(.(get_coef[3]) * x - .(get_coef[4])) ^ 2))
  fn_try_deriv <- D(fn_try_deriv, 'x')
  x = seq(1, 67, length.out = 6700)
  tmp <- eval(fn_try_deriv)
  flex_fa[i] <- min(x[tmp < 0])
}
flex_fa <- flex_fa[-c(12,112,124,133)]
mean(flex_fa)
sd(flex_fa)
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
    geom_point(aes(x=1:146, y=flex_fa), color="#1b95ff") +
    geom_hline(yintercept=mean(flex_fa), color="#1b95ff", linetype="dashed", size=1.5) + 
    labs(title = "MonteCarlo Simulation for Day of Flex (Fatalities)", x = "Trial", y = "Day of Flex")
```

## Applying the day of flex indicator to the plot (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#plot(ft)
ggplot(italy, aes(x = days, y = total_deaths)) +
    geom_point() +
    stat_function(fun=df_italy, color="#1b95ff", size=1.5, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Deaths in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Deaths") +
  geom_vline(xintercept = mean(flex_fa), color="#1b95ff", size=1, linetype="dashed")
```

# Australia Simulation

## Initial Data for Australia (Cases)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
australia <- data %>% filter(location == "Australia", date > "2020-06-01", date < "2020-12-19")
australia <- subset(australia, select = c(date, total_cases, total_deaths))
australia <- australia %>% mutate(days = 1:n())
minimizedDays <- 70
australiaPre <- australia %>% filter(days < minimizedDays)
cases_Australia <- ggplot(australiaPre, aes(x = days, y = total_cases)) + 
    geom_point() + 
    labs(title = "Total Cases in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Cases")
fatalities_Australia <- ggplot(australiaPre, aes(x = days, y = total_deaths)) +
    geom_point() + 
    labs(title = "Total Deaths in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Deaths")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(cases_Australia)
```

## Initial Data for Australia (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(fatalities_Australia)
```

## Fitting a Gauss Error Function to Australia Data (Cases)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
x <- 0:minimizedDays

gauss_fn <- function(x,a,b,c,d) {a+b*erf(c*x-d)}
cf <- summary(nlsLM(total_cases ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=10000,b=500,c=.02,d=1)))$parameters
ff <- summary(nlsLM(total_deaths ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=100,b=100,c=.01,d=1)))$parameters

cases_function_australia <- function(x) cf[1]+cf[2]*erf(cf[3]*x-cf[4])
fatalities_function_australia <- function(x) ff[1]+ff[2]*erf(ff[3]*x-ff[4])

cases_pre_plot_Australia <- ggplot(australiaPre, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, minimizedDays, 10)) +
    labs(title = "Total Cases in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Cases")
fatalities_pre_plot_Australia <- ggplot(australiaPre, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=fatalities_function_australia, color="#1b95ff", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, minimizedDays, 10)) +
    labs(title = "Total Deaths in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Deaths")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(cases_pre_plot_Australia)
```

## Fitting a Gauss Error Function to Australia Data (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(fatalities_pre_plot_Australia)
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
x <- 0:200

# cases_function_australia <- function(x) 17200+10000*erf(0.04*x-2.4)
# fatalities_function_australia <- function(x) 500+400*erf(0.038*x-3.1)

cases_plot_Australia <- ggplot(australia, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    labs(title = "Total Cases in Australia from 6/2/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Cases")
fatalities_plot_Australia <- ggplot(australia, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=fatalities_function_australia, color="#1b95ff", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    labs(title = "Total Deaths in Australia from 6/2/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Deaths")
```

## Applying the function to the future dataset (Cases)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
cases_plot_Australia_future <- ggplot(australiaPre, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    xlim(0, 200) +
    labs(title = "Extrapolating Future Case Data with Gauss Error Function", x = "Days", y = "Total Cases")
fatalities_plot_Australia_future <- ggplot(australiaPre, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=fatalities_function_australia, color="#1b95ff", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    xlim(0, 200) +
    labs(title = "Extrapolating Future Fatality Data with Gauss Error Function", x = "Days", y = "Total Deaths")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(cases_plot_Australia_future)
```

## Applying the function to the future dataset (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(fatalities_plot_Australia_future)
```

# MonteCarlo Simulation for Australia's Day of Flex

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
#Montecarlo Simulation
set.seed(400)
n <- dim(australia)[1]
m <- 150
sigma <- 0.1
mu <- 1
gau_sample <- matrix(rnorm(n*m, mean=mu, sd=sigma), nrow = m, ncol = n)

mc_data_ca <- c()
for (i in 1:n) {
    mc_data_ca <- c(mc_data_ca, australia$total_cases[i]*gau_sample[,i])
}
mc_data_ca <- matrix(mc_data_ca, nrow = m, ncol = n)
mc_data_fa <- c()
for (i in 1:n) {
    mc_data_fa <- c(mc_data_fa, australia$total_deaths[i]*gau_sample[,i])
}
mc_data_fa <- matrix(mc_data_fa, nrow = m, ncol = n)

day_of_flex_ca <- rep(NA, m)
for (i in 1:m) {
    trial <- mc_data_ca[i,]
    trial_model <- nlsLM(trial ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=15000,b=9000,c=.01,d=1))
    model_coef <- summary(trial_model)$parameters[, 1]
    # trial_function <- function(x, coef) {coef[1] + coef[2]*erf(coef[3]*x-coef[4])}
    gauserr_derivative <- {2/sqrt(pi)*exp(-x^2)}
    function_trial_derivative <- D(bquote(.(model_coef[2])*2/sqrt(pi)*exp(-(.(model_coef[3])*x-.(model_coef[4]))^2)), 'x')
    x = seq(1, 200, length.out=6700)
    temp <- eval(function_trial_derivative)
    day_of_flex_ca[i] <- min(x[temp < 0])
}
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
day_of_flex_fa <- rep(NA, m)
for (i in 1:m) {
    trial <- mc_data_fa[i,]
    trial_model <- nlsLM(trial ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=300,b=200,c=.01,d=1))
    model_coef <- summary(trial_model)$parameters[, 1]
    # trial_function <- function(x, coef) {coef[1] + coef[2]*erf(coef[3]*x-coef[4])}
    gauserr_derivative <- {2/sqrt(pi)*exp(-x^2)}
    function_trial_derivative <- D(bquote(.(model_coef[2])*2/sqrt(pi)*exp(-(.(model_coef[3])*x-.(model_coef[4]))^2)), 'x')
    x = seq(1, 200, length.out=6700)
    temp <- eval(function_trial_derivative)
    day_of_flex_fa[i] <- min(x[temp < 0])
}
```

## MonteCarlo Simulation Data (Cases) (150 trials)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
australiaMCPlot_cases <- ggplot() +
    geom_point(aes(x=1:m, y=day_of_flex_ca), color="#d33636") +
    geom_hline(yintercept=mean(day_of_flex_ca), color="#d33636", linetype="dashed", size=2) + 
    labs(title = "MonteCarlo Simulation for Day of Flex (Cases)", x = "Trial", y = "Day of Flex")
australiaMCPlot_fatalities <- ggplot() +
    geom_point(aes(x=1:m, y=day_of_flex_fa), color="#1b95ff") +
    geom_hline(yintercept=mean(day_of_flex_fa), color="#1b95ff", linetype="dashed", size=2) + 
    labs(title = "MonteCarlo Simulation for Day of Flex (Fatalities)", x = "Trial", y = "Day of Flex")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(australiaMCPlot_cases)
```

## MonteCarlo Simulation Data (Fatalities) (150 trials)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(australiaMCPlot_fatalities)
```

## Using MonteCarlo Simulation to Predict the Day of Flex

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
print(mean(day_of_flex_fa))
print(sd(day_of_flex_fa))
print(mean(day_of_flex_ca))
print(sd(day_of_flex_ca))
```

- The mean of the day of flex for cases is 60.22 days and the standard deviation is 0.77 days.
- The mean of the day of flex for fatalities is 81.4 days and the standard deviation is 0.69 days.

## Applying the day of flex indicator to the data (Cases)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(cases_plot_Australia_future + geom_vline(xintercept = mean(day_of_flex_ca), color="#d33636", size=1.5, linetype="dashed") + labs(title="Day of flex indicator for cases (60 days after)"))
```

## Applying the day of flex indicator to the data (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(fatalities_plot_Australia_future + geom_vline(xintercept = mean(day_of_flex_fa), color="#1b95ff", size=1.5, linetype="dashed") + labs(title="Day of flex indicator for fatalities (81 days after)"))
```

# Confirming Results with entire dataset

## Entire data (Cases)

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
cases_plot_Australia_confirm <- ggplot(australia, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    geom_vline(xintercept = mean(day_of_flex_ca), color="#d33636", size=1.5, linetype="dashed") +
    labs(title = "Entire Cases in Australia from 6/2/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Cases")
fatalities_plot_Australia_confirm <- ggplot(australia, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=fatalities_function_australia, color="#1b95ff", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    geom_vline(xintercept = mean(day_of_flex_fa), color="#1b95ff", size=1.5, linetype="dashed") + 
    labs(title = "Entire Fatalities in Australia from 6/2/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Deaths")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(cases_plot_Australia_confirm)
```

## Entire data (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(fatalities_plot_Australia_confirm)
```

# Interesting Anomaly with United States Data

## All United States Data (Cases)

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
# US data
us <- data %>% filter(location == "United States")
us <- subset(us, select = c(date, total_cases, total_deaths))
us <- us %>% mutate(days = 1:n())
us <- us %>% mutate(year = substr(date, 1, 4))
us_data_cases <- ggplot(data=us, aes(x=days, y=total_cases, color=year)) + 
  geom_point() + 
  labs(title="Total Cases in the United States from 1/22/2020 to 11/14/2022 (1,028 days)", color="Year", x="Days", y="Total Cases")
us_data_deaths <- ggplot(data=us, aes(x=days, y=total_deaths, color=year)) + 
  geom_point() + 
  labs(title="Total Deaths in the United States from 1/22/2020 to 11/14/2022 (1,028 days)", color="Year", x="Days", y="Total Deaths")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(us_data_cases)
```

## All United States Data (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(us_data_deaths)
```

## Conclusion

- The day of flex indicator is a good indicator to predict the time of the peak of the pandemic (for most countries).
- The gauss error function is a good for countries like...
  - China, Italy and Australia
- The gauss error function is not a good fit for countries like...
  - United States
  - Any other countries who may have different covid restrictions
- MonteCarlo simulations are neccesary due to the uncertainty of the data.

## Works Cited
- Link to the study:
  - I. Ciufolini, A. Paolozzi, Prediction of the time evolution of the Covid-19 Pandemic in Italy by a Gauss
Error Function and Monte Carlo simulations. Submitted to BioRxiv on 03.26.2020 and transferred on
03.27.2020 to MedRxiv. https://doi.org/10.1101/2020.03.27.20045104
- Link to the covid data used:
  - https://covid19.who.int/
- Source to find the second derivative of fitted function:
  - https://proofwiki.org/wiki/Derivative_of_Error_Function

### 2022-12-08 austins_temp_powerpoint.rmd
updated

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
-knitr::opts_chunk$set(fig.width=12, fig.height=8) 
+# Import packages
+library(tidyverse)
+library(ggplot2)
+library(GGally)
+library(knitr)
+library(boot)
+library(caret)
+library(randomForest)
+library(e1071)
+library(class)
+knitr::opts_chunk$set(fig.width=14, fig.height=8)
+set.seed(420)
 ```
 
-# Breast Cancer Diagnosis Classification Analysis
+```{r load_data, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+# Load the data
+data <- read.csv("breast_cancer_diagnostic_data.csv") # Full Data
+training <- read.csv("training.csv") # Training Data
+test <- read.csv("test.csv")  # Test Data
+# Data Cleaning
+data <- data[,-1] # Remove ID column
+test <- test[,-1] # Remove ID column
+training <- training[,-1] # Remove ID column
+# Convert diagnosis to factor
+data$diagnosis <- as.factor(data$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+training$diagnosis <- as.factor(training$diagnosis)
+```
+
+## Background
+
+- Breast Cancer Diagnosis Data
+- Predict Diagnosis of Tumor based on features (Malignant or Benign)
+- Reduce Type II Error (False-Negatives)
+  - Feature and Model Selection
+
+# Models
+
+```{r train_setup, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+train_control <- trainControl(method = "cv", number = 10, preProcOptions = list(thresh=0.99), classProbs=TRUE, summaryFunction = twoClassSummary)
+```
+
+## KNN
+
+```{r austin_knn, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+# KNN
+knn <- train(diagnosis ~ ., data = training, method = "knn", metric="ROC", trControl = train_control)
+knn_pred <- predict(knn, test)
+knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
+varsKNN <- rownames(data.frame(varImp(knn)[1]))[1:10]
+```
+
+```{r austin_knn_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(knn_confusion$table), aes(x = Reference, y = rev(Prediction), fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "KNN Confusion Matrix", x = "Reference", y = "Prediction") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=2+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+## Random Forrest
+
+```{r austin_random_forrest, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+# Logistic Regression without cross validation
+random_forrest <- train(diagnosis ~ ., data = training, method = "ranger", metric="ROC", trControl = train_control, preProcess=c("center", "scale"))
+rf_pred <- predict(random_forrest, test)
+rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
+```
+
+```{r austin_random_forrest_plot, echo=FALSE, message=FALSE, warning=FALSE}
+# Plot confusion matrix
+ggplot(data.frame(rf_confusion$table), aes(x = Reference, y = rev(Prediction), fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Random Forrest Confusion Matrix", x = "Reference", y = "Prediction") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=2+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+```{r austin_random_forrest_plot2, echo=FALSE, message=FALSE, warning=FALSE}
+#Plotting the false positives and true positives for the logistic regression model
+false_pred <- !(test$diagnosis != rf_pred)
+false_negative <- ifelse((test$diagnosis == "M") & (rf_pred == "B"), "FalseNeg", "Okay")
+ggplot() + geom_point(aes(x=test$texture_mean, y=test$radius_mean, color=false_negative), size=2) + labs(title = "Random Forrest Model False Negatives")
+```
+
+## Radial Support Vector Machine
+
+```{r austin_svm, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+#Radial Support Vector Machine
+svm <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric="ROC")
+svm_pred <- predict(svm, test)
+svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive="M")
+```
+
+```{r austin_svm_plot, echo=FALSE, message=FALSE, warning=FALSE}
+
+ggplot(data.frame(svm_confusion$table), aes(x = Reference, y = rev(Prediction), fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Radial SVM Confusion Matrix", x = "Reference", y = "Prediction") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=2+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+```{r austin_svm2, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+tuned_svm <- tune.svm(diagnosis ~., data = training, gamma = 10^(-5:-1), cost = 10^(-3:1))
+```
+
+```{r austin_svm22, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+svm2 <- svm(diagnosis ~., data = training, gamma = 0.01, cost = 10, type="C-classification", kernel="radial")
+svm2_pred <- predict(svm2, test)
+svm2_confusion <- confusionMatrix(svm2_pred, test$diagnosis, positive="M")
+```
+
+```{r austin_svm_plot2, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(svm2_confusion$table), aes(x = Reference, y = rev(Prediction), fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Radial SVM Confusion Matrix", x = "Reference", y = "Prediction") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=2+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+```{r}
+learn_rf <- randomForest(diagnosis~., data=training, ntree=500, proximity=T, importance=T)
+```
+
+```{r}
+acc_test <- numeric() 
+for(i in 1:30){
+    predict <- knn(train=training[,-1], test=test[,-1], cl=training$diagnosis, k=i, prob=T)
+    acc_test <- c(acc_test,mean(predict==test[,1]))
+}
+```

### 2022-12-08 austins_temp_powerpoint.rmd
updated

@@
 ```{r train_setup, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
-train_control <- trainControl(method = "cv", number = 10, preProcOptions = list(thresh=0.99), classProbs=TRUE, summaryFunction = twoClassSummary)
+train_control <- trainControl(method = "repeatedcv", summaryFunction = defaultSummary, classProbs = TRUE, number = 10, repeats = 10)
 ```
@@
 
-```{r}
-acc_test <- numeric() 
-for(i in 1:30){
-    predict <- knn(train=training[,-1], test=test[,-1], cl=training$diagnosis, k=i, prob=T)
-    acc_test <- c(acc_test,mean(predict==test[,1]))
+#TUNNING MODELS
+
+# KNN
+
+## Tuning KNN
+
+```{r knn_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+type_two_error <- data.frame(k = seq(1, 100, 1), error = NA, typetwoerror = NA)
+
+for (i in 1:nrow(type_two_error)) {
+    knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = type_two_error$k[i]))
+    knn_pred <- predict(knn, test)
+    knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
+    type_two_error$typetwoerror[i] <- knn_confusion$table[1, 2]/sum(knn_confusion$table)
+    type_two_error$error[i] <- (knn_confusion$table[1, 2] + knn_confusion$table[2, 1])/sum(knn_confusion$table)
+}
+print(which.min(type_two_error$typetwoerror))
+```
+
+```{r knn_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot() + geom_line(aes(x=type_two_error$k, y=type_two_error$error), size=1) + geom_line(aes(x=type_two_error$k, y=type_two_error$typetwoerror), size=1.5, color="#c82798") + labs(title = "KNN Type I and Overall Error", x = "K", y = "Error")
+```
+
+## Tuned KNN
+
+```{r tuned_knn, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+tuned_knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = which.min(type_two_error$typetwoerror)))
+knn_pred <- predict(tuned_knn, test)
+knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
+```
+
+```{r tuned_knn_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(knn_confusion$table), aes(x = Reference, y = rev(Prediction), fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "KNN Confusion Matrix", x = "Reference", y = "Prediction") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=2+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+# Random Forrest
+
+## Tuning Random Forrest
+
+```{r rf_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+type_two_error <- data.frame(ntree = seq(1, 100, 10), error = NA, typetwoerror = NA)
+
+for (i in 1:nrow(type_two_error)) {
+    rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "Accuracy", ntree = type_two_error$ntree[i])
+    rf_pred <- predict(rf, test)
+    rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
+    type_two_error$typetwoerror[i] <- rf_confusion$table[1, 2]/sum(rf_confusion$table)
+    type_two_error$error[i] <- (rf_confusion$table[1, 2] + rf_confusion$table[2, 1])/sum(rf_confusion$table)
 }
 ```
+
+```{r rf_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot() + geom_line(aes(x=type_two_error$ntree, y=type_two_error$error), size=1) + geom_line(aes(x=type_two_error$ntree, y=type_two_error$typetwoerror), size=1.5, color="#c82798") + labs(title = "Random Forrest Type I and Overall Error", x = "Number of Trees", y = "Error")
+```
+
+## Tuned Random Forrest
+
+```{r tuned_rf, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+tuned_rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "ROC", ntree = which.min(type_two_error$typetwoerror))
+rf_pred <- predict(tuned_rf, test)
+rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
+```
+
+```{r tuned_rf_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(rf_confusion$table), aes(x = Reference, y = rev(Prediction), fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Random Forrest Confusion Matrix", x = "Reference", y = "Prediction") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=2+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+# SVM
+
+## Tuning SVM
+
+```{r svm_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+gamma <- seq(0.01, 100, length.out=7)
+cost <- seq(0.01, 100, length.out=7)
+errorLog <- data.frame(gamma = gamma, cost = cost, error = NA, typetwoerror = NA)
+
+for (i in 1:length(gamma)) {
+  for (j in 1:length(cost)) {
+    svm <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric = "Accuracy", gamma = gamma[i], cost = cost[j])
+    svm_pred <- predict(svm, test)
+    svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive = "M")
+    tempGamma <- gamma[i]
+    tempCost <- cost[j]
+    tempError <- (svm_confusion$table[1, 2] + svm_confusion$table[2, 1])/sum(svm_confusion$table)
+    tempTwo <- svm_confusion$table[1, 2]/sum(svm_confusion$table)
+    errorLog <- rbind(errorLog, data.frame(gamma = tempGamma, cost = tempCost, error = tempError, typetwoerror = tempTwo))
+    print(paste("completed", i, j))
+  }
+}
+errorLog <- tail(errorLog, -7)
+```
+
+```{r svm_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(errorLog, aes(x=gamma, y=cost, fill= typetwoerror)) + 
+  scale_fill_gradient(low="white", high="blue") +
+  geom_tile()
+```
+
+## Tuned SVM
+
+```{r tuned_svm, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+svm2 <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric = "Accuracy", gamma = 0.01, cost = 0.01)
+svm_pred <- predict(svm2, test)
+svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive = "M")
+print(svm_confusion)
+```
+
+```{r tuned_svm_plot, echo=FALSE, message=FALSE, warning=FALSE}
+```

### 2022-12-08 powerpoint.rmd

@@
 
+## Features
+
+- 
+-
+
 ## EDA
 
-- Different measurements of each variable are correlated
+Different measurements of each variable are correlated
 
-- Relationship with Diagnosis visible for each measurement
+Relationship with Diagnosis visible for each measurement
 
-```{r, fig.height=3, fig.width=4, fig.align='right', fig.pos='H'}
+```{r, fig.height=8, fig.width=12}
 source("Danny's Analysis.R", echo = F, local = T)

### 2022-12-08  PPTX-code.R
Create  PPTX-code.R

@@ -1,0 +1,53 @@
+### Packages
+
+library(tidyverse)
+library(GGally)
+library(ggplot2)
+library(tidyr)
+library(tidymodels)
+library(vip)
+library(caTools)
+library(xgboost)
+library(discrim)
+library(glmnet)
+library(kableExtra)
+
+### Gathering Data
+
+train <- read.csv("training.csv")
+test <- read.csv("test.csv")
+
+train |>
+  select(-id) -> train
+
+test |>
+  select(-id) -> test
+
+train$diagnosis <- as.factor(train$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+
+### EDA
+
+train |>
+  group_by(diagnosis) |>
+  summarise(radius = mean(radius_mean),
+            concavity = mean(concavity_mean),
+            perimeter = mean(perimeter_mean)) |>
+  kable(digits = 3) |>
+  kable_material(lightable_options = "striped")
+
+ggplot(train) +
+  geom_point(aes(x = concavity_mean, y = radius_mean, col = diagnosis))+
+  labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor")
+
+train[,1:11] |>
+  ggpairs()
+
+train[,c(1,12:21)] |>
+  ggpairs()
+
+train[,c(1,22:31)] |>
+  ggpairs()
+
+train[,c(1,2,12,22)] |>
+  ggpairs()

### 2022-12-08  PPTX-code.R
Update  PPTX-code.R

@@ -50,4 +50,4 @@ train[,c(1,22:31)] |>
   ggpairs()
 
 train[,c(1,2,12,22)] |>
-  ggpairs()
+  ggpairs() -> measure_pairs

### 2022-12-08 ethan_analysis.Rmd
Add files via upload

@@
 #install.packages("forecast")
+#install.packages("plyr")
+#install.packages("readr")
+#install.packages("repr")
 ```
@@
 
-setwd("~/Desktop/DSCI/DSCI445/Project")
+#setwd("~/Desktop/DSCI/DSCI445/Project")
 library(tidyverse)
@@
 library(forecast)
+library(GGally)
+library(plyr)
+library(readr)
+library(repr)
 
@@
 dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
-dataTot <- dataTemp %>% select_if(~ !any(is.na(.)))
-testTot <- read.csv("test.csv")
-trainingTot <- read.csv("training.csv")
+datatemp2 <- dataTemp %>% select_if(~ !any(is.na(.)))
+dataTot <- datatemp2[,!names(datatemp2) %in% c("id")]
+
+testtemp <- read.csv("test.csv")
+testTot <- testtemp[,!names(testtemp) %in% c("id")]
+
+traintemp <- read.csv("training.csv")
+trainTot <- traintemp[,!names(traintemp) %in% c("id")]
 
@@
 
-#Dividing the data, test, and training data between malignant and benign
 ```{r}
-#Malignant
-tempM <- dataTot[dataTot$diagnosis == 'M', ]
-dataM <- tempM[, !names(tempM) %in% c("diagnosis")]
-splitM <- sample(c(rep(0, 0.6 * nrow(dataM)), rep(1, 0.4 * nrow(dataM))))
-trainingM <- dataM[splitM == 0, ]
-testM <- dataM[splitM == 1, ]
-
-#Benign
-tempB <- dataTot[dataTot$diagnosis == 'B', ]
-dataB <- tempB[, !names(tempB) %in% c("diagnosis")]
-splitB <- sample(c(rep(0, 0.6 * nrow(dataB)), rep(1, 0.4 * nrow(dataB))))
-trainingB <- dataB[splitB == 0, ]
-testB <- dataB[splitB == 1, ]
+corr_df <- dataTot[,c('area_worst', 'area_mean', 'concave.points_mean', 'radius_worst')]
+#head(corr_df)
+ggpairs(corr_df)
 ```
 
-#Predicting the mean of the radius of the cancer cells between both Malignant & Benign combined and seperate
+--------------------------------------------------------------------------------
+
+## area_mean
+# Logestic regression
+
 ```{r}
-#The prediction on the Total testing data set
-reg_trainTot <- lm(radius_mean~., data = trainingTot)
-pred_testTot <- predict(reg_trainTot, newdata = testTot)
-error_testTot <- mean((testTot[, "radius_mean"] - pred_testTot)^2)
-
-#The prediction on the Malignant testing data set
-reg_trainM <- lm(radius_mean~., data = trainingM)
-pred_testM <- predict(reg_trainM, newdata = testM)
-error_testM <- mean((testM[, "radius_mean"] - pred_testM)^2)
-
-#The prediction on the Benign testing data set
-reg_trainB <- lm(radius_mean~., data = trainingB)
-pred_testB <- predict(reg_trainB, newdata = testB)
-error_testB <- mean((testB[, "radius_mean"] - pred_testB)^2)
-
-print("Total test set error:")
-print(error_testTot)
-print('Malignant test set error:')
-print(error_testM)
-print('Benign test set error:')
-print(error_testB)
-```
-
-#ridge regression
-```{r}
-#Total
-training_matxTot <- model.matrix(radius_mean ~ . , data = trainingTot)[, -1 ]
-test_matxTot <- model.matrix(radius_mean ~ . , data = testTot)[,-1]
-
-cv_finalTot = cv.glmnet(training_matxTot , trainingTot$radius_mean, nfolds = 10, alpha=0)
-lamTot = cv_finalTot$lambda.min
-
-ridge_regTot = glmnet(training_matxTot , trainingTot$radius_mean, nfolds = 10, alpha = 0)
-ridge_preTot = predict(ridge_regTot, s=lamTot, newx = test_matxTot)
-error_test_ridTot <- mean((ridge_preTot - testTot$radius_mean)^2)
-```
-```{r}
-#Malignant
-training_matxM <- model.matrix(radius_mean ~ . , data = trainingM)[, -1 ]
-test_matxM <- model.matrix(radius_mean ~ . , data = testM)[,-1]
-
-cv_finalM = cv.glmnet(training_matxM , trainingM$radius_mean, nfolds = 10, alpha=0)
-lamM = cv_finalM$lambda.min
-
-ridge_regM = glmnet(training_matxM , trainingM$radius_mean, nfolds = 10, alpha = 0)
-ridge_preM = predict(ridge_regM, s=lamM, newx = test_matxM)
-error_test_ridM <- mean((ridge_preM - testM$radius_mean)^2)
-```
-```{r}
-#Benign
-training_matxB <- model.matrix(radius_mean ~ . , data = trainingB)[, -1 ]
-test_matxB <- model.matrix(radius_mean ~ . , data = testB)[,-1]
-
-cv_finalB = cv.glmnet(training_matxB , trainingB$radius_mean, nfolds = 10, alpha=0)
-lamB = cv_finalB$lambda.min
-
-ridge_regB = glmnet(training_matxB , trainingB$radius_mean, nfolds = 10, alpha = 0)
-ridge_preB = predict(ridge_regB, s=lamB, newx = test_matxB)
-error_test_ridB <- mean((ridge_preB - testB$radius_mean)^2)
-```
-```{r}
-print("Total")
-print(error_test_ridTot)
-print("Malignant")
-print(error_test_ridM)
-print("Benign")
-print(error_test_ridB)
-```
-
-#lasso
-```{r}
-#Total
-lasso_setTot = glmnet(training_matxTot, trainingTot$radius_mean, nfolds = 10, alpha=1)
-lasso_preTot = predict(lasso_setTot , s= lamTot, newx=test_matxTot)
-error_test_lassoTot <- mean((lasso_preTot - testTot$radius_mean)^2)
-```
-```{r}
-#Maligant
-lasso_setM = glmnet(training_matxM, trainingM$radius_mean, nfolds = 10, alpha=1)
-lasso_preM = predict(lasso_setM , s= lamM, newx=test_matxM)
-error_test_lassoM <- mean((lasso_preM - testM$radius_mean)^2)
-```
-```{r}
-#Benign
-lasso_setB = glmnet(training_matxB, trainingB$radius_mean, nfolds = 10, alpha=1)
-lasso_preB = predict(lasso_setB , s= lamB, newx=test_matxB)
-error_test_lassoB <- mean((lasso_preB - testB$radius_mean)^2)
-```
-```{r}
-print("Total")
-print(error_test_lassoTot)
-print("Malignant")
-print(error_test_lassoM)
-print("Benign")
-print(error_test_lassoB)
-```
-
-#PCR
-```{r}
-#Total
-trainingTot2 = dataTot %>% sample_frac(0.6)
-testTot2 = dataTot %>% setdiff(trainingTot2)
-pcr_fitTot2 = pcr(radius_mean ~ ., data = trainingTot2 , scale = TRUE, validation = "CV")
-summary(pcr_fitTot2)
-```
-```{r}
-#Total cont M is 31
-x_trainingTot = model.matrix(radius_mean~., trainingTot2)[,-1]
-x_testTot = model.matrix(radius_mean~., testTot2)[,-1]
-
-y_trainingTot = trainingTot2 %>%
-  select(radius_mean) %>%
-  unlist() %>%
-  as.numeric()
-
-y_testTot = testTot2 %>%
-  select(radius_mean) %>%
-  unlist() %>%
-  as.numeric()
-
-pcr_predTot = predict(pcr_fitTot2, x_testTot, ncomp=31)
-error_test_pcrTot <- mean((pcr_predTot-y_testTot)^2)
+#Test
+am_lm <- lm(area_mean ~ ., data = trainTot)
+pred_amtest <- predict(am_lm, testTot)
+amtest_log <- postResample(pred_amtest, testTot$area_mean)
+amtest_log
+#summary(am_lm)
 ```
@@
 ```{r}
-#Malignant
-trainingM2 = dataM %>% sample_frac(0.6)
-testM2 = dataM %>% setdiff(trainingM2)
-pcr_fitM2 = pcr(radius_mean ~ ., data = trainingM2 , scale = TRUE, validation = "CV")
-summary(pcr_fitM2)
+#Train
+pred_amtrain<- predict(am_lm, trainTot)
+amtrain_log <- postResample(pred_amtrain, trainTot$area_mean)
+amtrain_log
 ```
+
+# ridge regression
 ```{r}
-#Maligant cont M is 30
-x_trainingM = model.matrix(radius_mean~., trainingM2)[,-1]
-x_testM = model.matrix(radius_mean~., testM2)[,-1]
+dummy <- dummyVars(area_mean~., data = dataTot[,-1])
+x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
+y_train = trainTot$area_mean
 
-y_trainingM = trainingM2 %>%
-  select(radius_mean) %>%
-  unlist() %>%
-  as.numeric()
+x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
+y_test = testTot$area_mean
 
-y_testM = testM2 %>%
-  select(radius_mean) %>%
-  unlist() %>%
-  as.numeric()
+lambdas <- 10^seq(2, -3, by = -.1)
+ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
+cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
+optimal_lambda <- cv_ridge$lambda.min
 
-pcr_predM = predict(pcr_fitM2, x_testM, ncomp=30)
-error_test_pcrM <- mean((pcr_predM-y_testM)^2)
+eval_results <- function(true, predicted, df) {
+  SSE <- sum((predicted - true)^2)
+  SST <- sum((true - mean(true))^2)
+  R_square <- 1 - SSE / SST
+  RMSE = sqrt(SSE/nrow(df))
+
+  # Model performance metrics
+data.frame(
+  RMSE = RMSE,
+  Rsquare = R_square)
+
+}
+
+# Prediction and evaluation on train data
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
+amr_train <- eval_results(y_train, predictions_train, trainTot)
+
+# Prediction and evaluation on test data
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+amr_test <- eval_results(y_test, predictions_test, testTot)
+
+print(amr_train)
+print(amr_test)
+```
+
+# lasso
+```{r}
+lambdas <- 10^seq(2, -3, by = -.1)
+lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+
+lambda_best <- lasso_reg$lambda.min 
+lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
+aml_train <- eval_results(y_train, predictions_train, trainTot)
+
+predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
+aml_test <- eval_results(y_test, predictions_test, testTot)
+```
+
+--------------------------------------------------------------------------------
+
+## Area_worst
+# Logestic regression
+
+```{r}
+#Test
+aw_lm <- lm(area_worst ~ ., data = trainTot)
+pred_awtest <- predict(aw_lm, testTot)
+awtest_log <- postResample(pred_awtest, testTot$area_worst)
+awtest_log
+#summary(am_lm)
 ```
@@
 ```{r}
-#Belign
-trainingB2 = dataB %>% sample_frac(0.6)
-testB2 = dataB %>% setdiff(trainingB2)
-pcr_fitB2 = pcr(radius_mean ~ ., data = trainingB2 , scale = TRUE, validation = "CV")
-summary(pcr_fitB2)
-```
-```{r}
-#Belign cont M is 29
-x_trainingB = model.matrix(radius_mean~., trainingB2)[,-1]
-x_testB = model.matrix(radius_mean~., testB2)[,-1]
-
-y_trainingB = trainingB2 %>%
-  select(radius_mean) %>%
-  unlist() %>%
-  as.numeric()
-
-y_testB = testB2 %>%
-  select(radius_mean) %>%
-  unlist() %>%
-  as.numeric()
-
-pcr_predB = predict(pcr_fitB2, x_testB, ncomp=29)
-error_test_pcrB <- mean((pcr_predB-y_testB)^2)
-```
-```{r}
-print("Total")
-print(error_test_pcrTot)
-print("Malignant")
-print(error_test_pcrM)
-print("Benign")
-print(error_test_pcrB)
+#Train
+pred_awtrain<- predict(aw_lm, trainTot)
+awtrain_log <- postResample(pred_awtrain, trainTot$area_worst)
+awtrain_log
 ```
 
-#PLS
+# ridge regression
 ```{r}
-#total
-pls_fitTot = plsr(radius_mean~., data = trainingTot2, scale = TRUE, validation = "CV")
-summary(pls_fitTot)
-```
-```{r}
-#Total cont M is 30
-pls_predTot = predict(pls_fitTot, x_testTot, ncomp = 30)
-error_test_plsTot <- mean((pls_predTot - y_testTot)^2)
+dummy <- dummyVars(area_worst~., data = dataTot[,-1])
+x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
+y_train = trainTot$area_worst
+
+x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
+y_test = testTot$area_worst
+
+lambdas <- 10^seq(2, -3, by = -.1)
+ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
+cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
+optimal_lambda <- cv_ridge$lambda.min
+
+eval_results <- function(true, predicted, df) {
+  SSE <- sum((predicted - true)^2)
+  SST <- sum((true - mean(true))^2)
+  R_square <- 1 - SSE / SST
+  RMSE = sqrt(SSE/nrow(df))
+
+  # Model performance metrics
+data.frame(
+  RMSE = RMSE,
+  Rsquare = R_square)
+
+}
+
+# Prediction and evaluation on train data
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
+awr_train <- eval_results(y_train, predictions_train, trainTot)
+
+# Prediction and evaluation on test data
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+awr_test <- eval_results(y_test, predictions_test, testTot)
+
+print(awr_train)
+print(awr_test)
 ```
 
-#Malignant
+# lasso
 ```{r}
-pls_fitM = plsr(radius_mean~., data = trainingM2, scale = TRUE, validation = "CV")
-summary(pls_fitM)
+lambdas <- 10^seq(2, -3, by = -.1)
+lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+
+lambda_best <- lasso_reg$lambda.min 
+lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
+awl_train <- eval_results(y_train, predictions_train, trainTot)
+
+predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
+awl_test <- eval_results(y_test, predictions_test, testTot)
 ```
-```{r}
-#Malignant cont M is 30
-pls_predM = predict(pls_fitM, x_testM, ncomp = 30)
-error_test_plsM <- mean((pls_predM - y_testM)^2)
-```
-
-```{r}
-#Benign
-pls_fitB = plsr(radius_mean~., data = trainingB2, scale = TRUE, validation = "CV")
-summary(pls_fitB)
-```
-
-```{r}
-#Benign cont M is 26
-pls_predB = predict(pls_fitB, x_testB, ncomp = 26)
-error_test_plsB <- mean((pls_predB - y_testB)^2)
-```
-```{r}
-print("Total")
-print(error_test_plsTot)
-print("Malignant")
-print(error_test_plsM)
-print("Benign")
-print(error_test_plsB)
-```

### 2022-12-09 ethan_temp_powerpoint.rmd
Add files via upload
Data and Features slides

@@
 ---
+
 ```{r setup, include=FALSE}
@@
 library(GGally)
+library(plyr)
+library(readr)
+library(repr)
 
@@
 dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
-data <- dataTemp %>% select_if(~ !any(is.na(.)))
-test <- read.csv("test.csv")
-training <- read.csv("training.csv")
+datatemp2 <- dataTemp %>% select_if(~ !any(is.na(.)))
+dataTot <- datatemp2[,!names(datatemp2) %in% c("id")]
+
+testtemp <- read.csv("test.csv")
+testTot <- testtemp[,!names(testtemp) %in% c("id")]
+
+traintemp <- read.csv("training.csv")
+trainTot <- traintemp[,!names(traintemp) %in% c("id")]
+
+knitr::opts_chunk$set(fig.width=12, fig.height=8) 
 ```
@@
 - Data collected from "Breast Cancer Wisconsin (Diagnostic) Data Set" from Kaggle
-- Response: Diagnosis (Benign/Malignant)
-- Predictors: 10 variables divided between  Mean, SE, and "Worst"
-- Training/Test Split (60/40)
+- Response: Diagnosis
+    - Benign: (Noncancerous Cells)
+    - Malignant (Cancerous Cells)
+- Predictors: 10 variables split into 3 categories:
+    - Mean 
+    - Standard Error (SE)
+    - Worst: Mean of 3 largest values of a variable
+- Training/Test Data Split 60/40
+
+## Features
+- Radius: Average Distance from cell center to cell perimeter
+- Texture: Standard deviation of gray-scale values; brightness of pixel of cell
+- Perimeter: Distance around nucleus boundary
+- Area: Area of the nucleus
+- Smoothness: Variation in cell's radial lengths
+- Compactness: The Perimeter^2/Area
+- Concavity: Size of the indention in nucleus boundary
+- Concave Points: Number of points on indented section of nucleus boundary
+- Symmetry: Deviation of the nuclei shape from the ideal measurement
+- Fractal Dimension: Measurement of irregularity in nucleus boundary
 
@@
 
-```{r, fig.height=3, fig.width=4, fig.align='right', fig.pos='H'}
-#source("Danny's Analysis.R", echo = F, local = T)
-
-#train[,c(1,2,12,22)] |>
-#  ggpairs(title = "Correlation b/w Radius Measurements")
-```
+## Classification Analysis: Danny
 
@@
 
-## Classification Analysis: Danny
+## Regression Analysis: Ethan
 
-## Regression Analysis: Ethan
+This is my validation plot
 ```{r}
-dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
-dataTot <- dataTemp %>% select_if(~ !any(is.na(.)))
-testTot <- read.csv("test.csv")
-trainingTot <- read.csv("training.csv")
-
-dataE <- data
+dataE <- dataTot
 corr_df <- dataE[,c('area_worst', 'area_mean', 'concave.points_mean', 'radius_worst')]
@@
 ```
--Test
 
@@
 
-- Breast Cancer Wisconsin (Diagnostic) Data Set: https://www.kaggle.com/datasets/uciml/breast-cancer-wisconsin-data?datasetId=180&sortBy=voteCount
+- Breast Cancer Wisconsin (Diagnostic) Data Set:
+    - https://www.kaggle.com/datasets/uciml/breast-cancer-wisconsin-data?datasetId=180&sortBy=voteCount
+- Definition of Features (Variables):
+    - https://www.causeweb.org/usproc/sites/default/files/usclap/2017-2/Evaluating_Benign_and_Malignant_Breast_Cancer_Cells_from_Fine-Needle_Aspirates.pdf

### 2022-12-09  PPTX-code.R
more updates

@@ -38,7 +38,7 @@ train |>
 
 ggplot(train) +
   geom_point(aes(x = concavity_mean, y = radius_mean, col = diagnosis))+
-  labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor")
+  labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor") -> plot1
 
 train[,1:11] |>
   ggpairs()
@@ -50,4 +50,114 @@ train[,c(1,22:31)] |>
   ggpairs()
 
 train[,c(1,2,12,22)] |>
-  ggpairs() -> measure_pairs
+  ggpairs() -> plot2
+
+### Models
+
+### SVC
+
+df_cost <- grid_regular(cost(), levels = 10)
+
+svm_linear_tune_spec <- svm_poly(degree = 1, cost = tune("cost")) %>%
+  set_mode("classification") %>%
+  set_engine("kernlab", scaled = FALSE)
+
+svm_linear_rec <- recipe(diagnosis ~ ., data = train)
+
+svm_linear_wf <- workflow() |>
+  add_model(svm_linear_tune_spec) |>
+  add_recipe(svm_linear_rec)
+
+tune_fit <- svm_linear_wf |>
+  tune_grid(resamples = cancer_10foldcv, grid = df_cost)
+
+autoplot(tune_fit)
+show_best(tune_fit, metric = "accuracy", n = 1)
+
+svm_linear_final <- finalize_workflow(svm_linear_wf, select_best(tune_fit, metric = "accuracy"))
+svm_linear_final |>
+  fit(data = train) -> svm_linear_final_fit
+
+svm_linear_final_fit |>
+  extract_fit_engine() -> svm_linear_final_fit_engine
+
+svm_linear_final_fit |>
+  augment(new_data = test) |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  conf_mat(truth = diagnosis, estimate = pred_lower_cutoff) -> svm.cm
+
+autoplot(svm.cm, type = "heatmap") -> plot3
+
+svm_linear_final_fit |>
+  augment(new_data = test) |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
[…]

### 2022-12-09 ethan_temp_powerpoint.rmd
Add files via upload
Fixed error for merge

@@
 ---
-
 ```{r setup, include=FALSE}

### 2022-12-09 ethan_temp_powerpoint.rmd

@@
 
+## Classification Analysis: Austin
+
+## Classification Analysis: Danny
+
+## Regression Analysis: Ethan
+```{r}
+dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
+dataTot <- dataTemp %>% select_if(~ !any(is.na(.)))
+testTot <- read.csv("test.csv")
+trainingTot <- read.csv("training.csv")
+
+dataE <- data
+corr_df <- dataE[,c('area_worst', 'area_mean', 'concave.points_mean', 'radius_worst')]
+ggpairs(corr_df)
+```
+-Test
+
 ## Model Assesments

### 2022-12-09 powerpoint.rmd

@@
 
-```{r, fig.height=8, fig.width=12}
-source("Danny's Analysis.R", echo = F, local = T)
+```{r, fig.height=6, fig.width=8}
+source("PPTX-code.R", echo = F, local = T)
 
-train[,c(1,2,12,22)] |>
-  ggpairs(title = "Correlation b/w Radius Measurements")
 ```
@@
 ## Classification Analysis: Danny
+
+- LogReg
+- KNN
+- QDA
+- RF
+- xGBoost
+- SVC
+- Lower prediction cutoff to account for Type II Error
+
+## K-Nearest Neigbors Model
+
+- Tuned neighbors value
+- 2 False-Negatives
+- 96.04% prediction accuracy
+
+## Support Vector Classifier Model
+
+- Tuned cost value
+- 2 False-Negatives
+- 96.48% prediction accuracy
+
+## Random Forest Model
+
+- 1 False-Negative
+- 95.15% prediction accuracy
 

### 2022-12-09 powerpoint.rmd

--- 2022-12-08 powerpoint.rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 14, fig.height = 8)
 ```
@@
 
+```{r}
+ggplot(train) +
+  geom_point(aes(x = concavity_mean, y = radius_mean, col = diagnosis))+
+  labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor")
+
+```
+
 Different measurements of each variable are correlated
@@
 
-```{r, fig.height=8, fig.width=12}
-source("Danny's Analysis.R", echo = F, local = T)
-
-train[,c(1,2,12,22)] |>
-  ggpairs(title = "Correlation b/w Radius Measurements")
-```
-
 ## Classification Analysis: Austin
@@
 ## Classification Analysis: Danny
+
+```{r}
+### Packages
+
+library(tidyverse)
+library(GGally)
+library(ggplot2)
+library(tidyr)
+library(tidymodels)
+library(vip)
+library(caTools)
+library(xgboost)
+library(discrim)
+library(glmnet)
+library(kableExtra)
+
+### Gathering Data
+
+train <- read.csv("training.csv")
+test <- read.csv("test.csv")
+
+train |>
+  select(-id) -> train
+
+test |>
+  select(-id) -> test
+
+train$diagnosis <- as.factor(train$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+```
+
+- Logistic Regression
+- K-Nearest Neighbors
+- QDA
+- Random Forest
+- xGBoost
+- Support Vector Classifier
+- Lower prediction cutoff to account for Type II Error
+
+## K-Nearest Neigbors Model
+
+```{r}
+set.seed(445)
+### KNN
+
+cancer_10foldcv <- vfold_cv(train, v = 10)
+neighbors_df <- data.frame(neighbors = c(1,3,5,7,10))
+
+knn_spec <- the_model <- nearest_neighbor(neighbors = tune("neighbors")) %>%
+  set_engine("kknn") %>% 
+  set_mode("classification")
+
+knn_rec <- recipe(diagnosis ~ ., data = train)
+
+workflow() |>
+  add_model(knn_spec) |>
+  add_recipe(knn_rec) -> knn_wf
+
+knn_wf |>
+  tune_grid(resamples = cancer_10foldcv, grid = neighbors_df) -> knn_tune
+
+knn_final <- finalize_workflow(knn_wf, select_best(knn_tune, metric = "accuracy"))
+knn_final_fit <- fit(knn_final, data = train)
+
+knn_final_fit |>
+  augment(new_data = test) -> knn_test_res
+
+knn_test_res |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  conf_mat(truth = diagnosis, estimate = pred_lower_cutoff) -> knn.cm
+
+knn_test_res |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  accuracy(truth = diagnosis, estimate = pred_lower_cutoff) |>
+  mutate(error = 1 - .estimate) |>
+  pull(error) -> knn.err
+
+100 - (round(knn.err, 4) * 100) -> knn.err
+
+neighbors_best <- select_best(knn_tune, metric = "accuracy") |> pull(neighbors)
+```
+
+- Tuned neighbors value to `r neighbors_best`
+- 2 False-Negatives
+- `knn.err` % prediction accuracy
+
+```{r}
+ggplot(data.frame(knn.cm$table), aes(x = Prediction, y = Truth, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "K-Nearest Neighbors Confusion Matrix", x = "Prediction", y = "Truth") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798") +
+  guides(fill = F)
+```
+
+## Support Vector Classifier Model
+
+```{r}
+set.seed(445)
+### SVC
+
+df_cost <- grid_regular(cost(), levels = 10)
+
+svm_linear_tune_spec <- svm_poly(degree = 1, cost = tune("cost")) %>%
+  set_mode("classification") %>%
+  set_engine("kernlab", scaled = FALSE)
+
+svm_linear_rec <- recipe(diagnosis ~ ., data = train)
+
+svm_linear_wf <- workflow() |>
+  add_model(svm_linear_tune_spec) |>
+  add_recipe(svm_linear_rec)
+
+tune_fit <- svm_linear_wf |>
+  tune_grid(resamples = cancer_10foldcv, grid = df_cost)
+
+svm_linear_final <- finalize_workflow(svm_linear_wf, select_best(tune_fit, metric = "accuracy"))
+
+svm_linear_final |>
+  fit(data = train) -> svm_linear_final_fit
+
+svm_linear_final_fit |>
+  extract_fit_engine() -> svm_linear_final_fit_engine
+
+svm_linear_final_fit |>
+  augment(new_data = test) |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  conf_mat(truth = diagnosis, estimate = pred_lower_cutoff) -> svm.cm
+
+svm_linear_final_fit |>
+  augment(new_data = test) |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  accuracy(truth = diagnosis, estimate = pred_lower_cutoff) |>
+  mutate(error = 1 - .estimate) |>
+  pull(error) -> svm.err
+
+100 - (round(svm.err, 4) * 100) -> svm.err
+
+cost_best <- select_best(tune_fit, metric = "accuracy") |> pull(cost)
+```
+
+- Tuned cost value to `r cost_best`
+- 2 False-Negatives
+- `svm.err` % prediction accuracy
+
+```{r}
+ggplot(data.frame(svm.cm$table), aes(x = Prediction, y = Truth, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Support Vector Classifier Confusion Matrix", x = "Prediction", y = "Truth") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798") +  
+  guides(fill = F)
+```
+
+## Random Forest Model
+
+```{r}
+### Random Forest
+
+rf_spec <- rand_forest(mtry = sqrt(.cols())) |>
+  set_engine("randomForest", importance = TRUE) |>
+  set_mode("classification")
+
+rf_fit <- rf_spec |>
+  fit(diagnosis ~ ., data = train)
+
+rf_fit |>
+  augment(new_data = test) |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  conf_mat(truth = diagnosis, estimate = pred_lower_cutoff) -> rf.cm
+
+rf_fit |>
+  augment(new_data = test) |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  accuracy(truth = diagnosis, estimate = pred_lower_cutoff) |>
+  mutate(error = 1 - .estimate) |>
+  pull(error) -> rf.err
+
+100 - (round(rf.err, 4) * 100) -> rf.err
+```
+
+- 1 False-Negative
+- `rf.err` % prediction accuracy
+
+```{r, echo = F}
+ggplot(data.frame(rf.cm$table), aes(x = Prediction, y = Truth, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Random Forest Confusion Matrix", x = "Prediction", y = "Truth") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798") +  
+  guides(fill = F)
+
+vip(rf_fit)
+```
 

### 2022-12-09 powerpoint.rmd
updated_main_pres

@@
 ```{r setup, include=FALSE}
+# Import packages
+library(tidyverse)
+library(ggplot2)
+library(GGally)
+library(knitr)
+library(boot)
+library(caret)
+library(randomForest)
+library(e1071)
+library(class)
+set.seed(420)
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 14, fig.height = 8)
@@
 
+- Many people are diagnosed with breast cancer each year
 - Breast Cancer Diagnosis Data
-- Predict Diagnosis of Tumor based on features
-- Reduce Type II Error (False-Negatives)
-- Feature Selection
+- Predict Diagnosis of Tumor based on features (Malignant or Benign)
+- *Modivation:* Reduce Type II Error (False-Negatives)
+  - Feature and Model Selection
 
@@
 ```{r}
-ggplot(train) +
-  geom_point(aes(x = concavity_mean, y = radius_mean, col = diagnosis))+
-  labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor")
-
+# ggplot(train) +
+#   geom_point(aes(x = concavity_mean, y = radius_mean, col = diagnosis))+
+#   labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor")
 ```
@@
 ## Classification Analysis: Austin
+
+```{r load_data, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+# Load the data
+set.seed(420)
+data <- read.csv("breast_cancer_diagnostic_data.csv") # Full Data
+training <- read.csv("training.csv") # Training Data
+test <- read.csv("test.csv")  # Test Data
+# Data Cleaning
+data <- data[,-1] # Remove ID column
+test <- test[,-1] # Remove ID column
+training <- training[,-1] # Remove ID column
+# Convert diagnosis to factor
+data$diagnosis <- as.factor(data$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+training$diagnosis <- as.factor(training$diagnosis)
+```
+
+# Models
+
+```{r train_setup, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+train_control <- trainControl(method = "repeatedcv", summaryFunction = defaultSummary, classProbs = TRUE, number = 10, repeats = 10)
+```
+
+# KNN Model
+
+## Tuning KNN
+
+```{r knn_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+type_two_error_knn <- data.frame(k = seq(1, 100, 1), error = NA, typetwoerror = NA)
+
+for (i in 1:nrow(type_two_error_knn)) {
+    knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = type_two_error_knn$k[i]))
+    knn_pred <- predict(knn, test)
+    knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
+    type_two_error_knn$typetwoerror[i] <- knn_confusion$table[1, 2]/sum(knn_confusion$table)
+    type_two_error_knn$error[i] <- (knn_confusion$table[1, 2] + knn_confusion$table[2, 1])/sum(knn_confusion$table)
+}
+print(which.min(type_two_error_knn$typetwoerror))
+knn1_k <- which.min(type_two_error_knn$typetwoerror)
+knn_tte <- min(type_two_error_knn$typetwoerror)
+```
+
+```{r knn_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot() + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$error), size=1) + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$typetwoerror), size=1.5, color="#c82798") + labs(title = "KNN Type I and Overall Error", x = "K", y = "Error")
+```
+
+## Tuned KNN
+
+```{r tuned_knn, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+set.seed(420)
+tuned_knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = which.min(type_two_error_knn$typetwoerror)))
+knn_pred <- predict(tuned_knn, test)
+knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
+```
+
+```{r tuned_knn_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(knn_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "KNN Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+## Tuned KNN Parameters
+
+- Using all features
+- 10-fold cross validation repeated 10 times
+- $k=$ `r knn1_k` neighbors
+- Type II Error: `r knn_tte`
+- Accuracy: `r knn_confusion$overall[1]`
+
+# Random Forrest
+
+## Tuning Random Forrest
+
+```{r rf_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+type_two_error_rf <- data.frame(ntree = seq(1, 100, 10), error = NA, typetwoerror = NA)
+
+for (i in 1:nrow(type_two_error_rf)) {
+    rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "Accuracy", ntree = type_two_error_rf$ntree[i])
+    rf_pred <- predict(rf, test)
+    rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
+    type_two_error_rf$typetwoerror[i] <- rf_confusion$table[1, 2]/sum(rf_confusion$table)
+    type_two_error_rf$error[i] <- (rf_confusion$table[1, 2] + rf_confusion$table[2, 1])/sum(rf_confusion$table)
+}
+print(which.min(type_two_error_rf$typetwoerror))
+rf1_ntree <- type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)]
+rf_tte <- min(type_two_error_rf$typetwoerror)
+```
+
+```{r rf_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot() + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$error), size=1) + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$typetwoerror), size=1.5, color="#c82798") + labs(title = "Random Forrest Type I and Overall Error", x = "Number of Trees", y = "Error")
+```
+
+## Tuned Random Forrest
+
+```{r tuned_rf, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+tuned_rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "ROC", ntree = type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)])
+rf_pred_tuned <- predict(tuned_rf, test)
+rf_confusion <- confusionMatrix(rf_pred_tuned, test$diagnosis, positive = "M")
+```
+
+```{r tuned_rf_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(rf_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Random Forrest Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+## Tuned Random Forrest Parameters
+
+- Using all features
+- 10-fold cross validation repeated 10 times
+- $n=$ `r rf1_ntree` trees
+- Type II Error: `r rf_tte`
+- Accuracy: `r rf_confusion$overall[1]`
+
+## Radial Support Vector Machine
+
+```{r austin_svm, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+#Radial Support Vector Machine
+svm <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric="ROC")
+svm_pred <- predict(svm, test)
+svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive="M")
+
+svm_tte <- svm_confusion$table[1, 2]/sum(svm_confusion$table)
+```
+
+```{r austin_svm_plot, echo=FALSE, message=FALSE, warning=FALSE}
+
+ggplot(data.frame(svm_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+## Radial Support Vector Machine Parameters
+
+- Using all features
+- 10-fold cross validation repeated 10 times
+- Type II Error: `r svm_tte`
+- Accuracy: `r svm_confusion$overall[1]`
+
+## Tuned Radial Support Vector Machine
+
+```{r austin_svm2, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+tuned_svm <- tune.svm(diagnosis ~., data = training, gamma = 10^(-5:-1), cost = 10^(-3:1))
+svm2 <- svm(diagnosis ~., data = training, gamma = 0.01, cost = 10, type="C-classification", kernel="radial")
+svm2_pred <- predict(svm2, test)
+svm2_confusion <- confusionMatrix(svm2_pred, test$diagnosis, positive="M")
+
+svm2_tte <- svm2_confusion$table[1, 2]/sum(svm2_confusion$table)
+
+```
+
+```{r austin_svm_plot2, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(svm2_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Tuned Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+## Tuned Radial Support Vector Machine Parameters
+
+- Using all features
+- 10-fold cross validation repeated 10 times
+- Type II Error: `r svm2_tte`
+- Accuracy: `r svm2_confusion$overall[1]`
 

### 2022-12-09 powerpoint.rmd

@@
 ```{r}
-# ggplot(train) +
-#   geom_point(aes(x = concavity_mean, y = radius_mean, col = diagnosis))+
-#   labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor")
+ggplot(train) +
+  geom_point(aes(x = concavity_mean, y = radius_mean, col = diagnosis))+
+  labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor")
 ```
@@
 
-## Model Assesments
+## Model Assessments
 

### 2022-12-09 powerpoint.rmd
update 2

@@
 ```{r setup, include=FALSE}
-# Import packages
-library(tidyverse)
-library(ggplot2)
-library(GGally)
-library(knitr)
-library(boot)
-library(caret)
-library(randomForest)
-library(e1071)
-library(class)
-set.seed(420)
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 14, fig.height = 8)
@@
 
-- Many people are diagnosed with breast cancer each year
 - Breast Cancer Diagnosis Data
-- Predict Diagnosis of Tumor based on features (Malignant or Benign)
-- *Modivation:* Reduce Type II Error (False-Negatives)
-  - Feature and Model Selection
+- Predict Diagnosis of Tumor based on features
+- Reduce Type II Error (False-Negatives)
+- Feature Selection
 
@@
 ```{r}
+### Packages
+
+library(tidyverse)
+library(GGally)
+library(ggplot2)
+library(tidyr)
+library(tidymodels)
+library(vip)
+library(caTools)
+library(xgboost)
+library(discrim)
+library(glmnet)
+library(kableExtra)
+
+### Gathering Data
+
+train <- read.csv("training.csv")
+test <- read.csv("test.csv")
+
+train |>
+  select(-id) -> train
+
+test |>
+  select(-id) -> test
+
+train$diagnosis <- as.factor(train$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+```
+
+```{r}
 ggplot(train) +
@@
   labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor")
+
+train[,c(1,2,12,22)] |>
+  ggpairs()
 ```
@@
 
-```{r}
-### Packages
-
-library(tidyverse)
-library(GGally)
-library(ggplot2)
-library(tidyr)
-library(tidymodels)
-library(vip)
-library(caTools)
-library(xgboost)
-library(discrim)
-library(glmnet)
-library(kableExtra)
-
-### Gathering Data
-
-train <- read.csv("training.csv")
-test <- read.csv("test.csv")
-
-train |>
-  select(-id) -> train
-
-test |>
-  select(-id) -> test
-
-train$diagnosis <- as.factor(train$diagnosis)
-test$diagnosis <- as.factor(test$diagnosis)
-```
-
 - Logistic Regression
@@
 - `knn.err` % prediction accuracy
+
+## KNN Confusion Matrix
 
@@
 
+## SVC Confusion Matrix
+
 ```{r}
@@
 ### Random Forest
-
+set.seed(445)
 rf_spec <- rand_forest(mtry = sqrt(.cols())) |>
@@
 
+## RF Confusion Matrix
+
 ```{r, echo = F}
@@
   guides(fill = F)
+```
 
+## RF Confusion Matrix
+
+```{r, echo = FALSE}
 vip(rf_fit)

### 2022-12-09 powerpoint.rmd

@@
 ## EDA
+
+```{r}
+ggplot(train) +
+  geom_point(aes(x = concavity_mean, y = radius_mean, col = diagnosis))+
+  labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor")
+```
+
+Different measurements of each variable are correlated
+
+Relationship with Diagnosis visible for each measurement
+
+## Classification Analysis: Austin
+
+```{r load_data, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+# Load the data
+set.seed(420)
+data <- read.csv("breast_cancer_diagnostic_data.csv") # Full Data
+training <- read.csv("training.csv") # Training Data
+test <- read.csv("test.csv")  # Test Data
+# Data Cleaning
+data <- data[,-1] # Remove ID column
+test <- test[,-1] # Remove ID column
+training <- training[,-1] # Remove ID column
+# Convert diagnosis to factor
+data$diagnosis <- as.factor(data$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+training$diagnosis <- as.factor(training$diagnosis)
+```
+
+# Models
+
+```{r train_setup, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+train_control <- trainControl(method = "repeatedcv", summaryFunction = defaultSummary, classProbs = TRUE, number = 10, repeats = 10)
+```
+
+# KNN Model
+
+## Tuning KNN
+
+```{r knn_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+type_two_error_knn <- data.frame(k = seq(1, 100, 1), error = NA, typetwoerror = NA)
+
+for (i in 1:nrow(type_two_error_knn)) {
+    knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = type_two_error_knn$k[i]))
+    knn_pred <- predict(knn, test)
+    knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
+    type_two_error_knn$typetwoerror[i] <- knn_confusion$table[1, 2]/sum(knn_confusion$table)
+    type_two_error_knn$error[i] <- (knn_confusion$table[1, 2] + knn_confusion$table[2, 1])/sum(knn_confusion$table)
+}
+print(which.min(type_two_error_knn$typetwoerror))
+knn1_k <- which.min(type_two_error_knn$typetwoerror)
+knn_tte <- min(type_two_error_knn$typetwoerror)
+```
+
+```{r knn_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot() + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$error), size=1) + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$typetwoerror), size=1.5, color="#c82798") + labs(title = "KNN Type I and Overall Error", x = "K", y = "Error")
+```
+
+## Tuned KNN
+
+```{r tuned_knn, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+set.seed(420)
+tuned_knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = which.min(type_two_error_knn$typetwoerror)))
+knn_pred <- predict(tuned_knn, test)
+knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
+```
+
+```{r tuned_knn_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(knn_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "KNN Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+## Tuned KNN Parameters
+
+- Using all features
+- 10-fold cross validation repeated 10 times
+- $k=$ `r knn1_k` neighbors
+- Type II Error: `r knn_tte`
+- Accuracy: `r knn_confusion$overall[1]`
+
+# Random Forrest
+
+## Tuning Random Forrest
+
+```{r rf_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+type_two_error_rf <- data.frame(ntree = seq(1, 100, 10), error = NA, typetwoerror = NA)
+
+for (i in 1:nrow(type_two_error_rf)) {
+    rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "Accuracy", ntree = type_two_error_rf$ntree[i])
+    rf_pred <- predict(rf, test)
+    rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
+    type_two_error_rf$typetwoerror[i] <- rf_confusion$table[1, 2]/sum(rf_confusion$table)
+    type_two_error_rf$error[i] <- (rf_confusion$table[1, 2] + rf_confusion$table[2, 1])/sum(rf_confusion$table)
+}
+print(which.min(type_two_error_rf$typetwoerror))
+rf1_ntree <- type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)]
+rf_tte <- min(type_two_error_rf$typetwoerror)
+```
+
+```{r rf_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot() + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$error), size=1) + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$typetwoerror), size=1.5, color="#c82798") + labs(title = "Random Forrest Type I and Overall Error", x = "Number of Trees", y = "Error")
+```
+
+## Tuned Random Forrest
+
+```{r tuned_rf, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+tuned_rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "ROC", ntree = type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)])
+rf_pred_tuned <- predict(tuned_rf, test)
+rf_confusion <- confusionMatrix(rf_pred_tuned, test$diagnosis, positive = "M")
+```
+
+```{r tuned_rf_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(rf_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Random Forrest Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+## Tuned Random Forrest Parameters
+
+- Using all features
+- 10-fold cross validation repeated 10 times
+- $n=$ `r rf1_ntree` trees
+- Type II Error: `r rf_tte`
+- Accuracy: `r rf_confusion$overall[1]`
+
+## Radial Support Vector Machine
+
+```{r austin_svm, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+#Radial Support Vector Machine
+svm <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric="ROC")
+svm_pred <- predict(svm, test)
+svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive="M")
+
+svm_tte <- svm_confusion$table[1, 2]/sum(svm_confusion$table)
+```
+
+```{r austin_svm_plot, echo=FALSE, message=FALSE, warning=FALSE}
+
+ggplot(data.frame(svm_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+## Radial Support Vector Machine Parameters
+
+- Using all features
+- 10-fold cross validation repeated 10 times
+- Type II Error: `r svm_tte`
+- Accuracy: `r svm_confusion$overall[1]`
+
+## Tuned Radial Support Vector Machine
+
+```{r austin_svm2, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+tuned_svm <- tune.svm(diagnosis ~., data = training, gamma = 10^(-5:-1), cost = 10^(-3:1))
+svm2 <- svm(diagnosis ~., data = training, gamma = 0.01, cost = 10, type="C-classification", kernel="radial")
+svm2_pred <- predict(svm2, test)
+svm2_confusion <- confusionMatrix(svm2_pred, test$diagnosis, positive="M")
+
+svm2_tte <- svm2_confusion$table[1, 2]/sum(svm2_confusion$table)
+
+```
+
+```{r austin_svm_plot2, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(svm2_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Tuned Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+## Tuned Radial Support Vector Machine Parameters
+
+- Using all features
+- 10-fold cross validation repeated 10 times
+- Type II Error: `r svm2_tte`
+- Accuracy: `r svm2_confusion$overall[1]`
+
+## Classification Analysis: Danny
 

### 2022-12-09 austins_temp_powerpoint.rmd

--- 2022-12-07 austins_temp_powerpoint.rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
-knitr::opts_chunk$set(fig.width=12, fig.height=8) 
+# Import packages
+library(tidyverse)
+library(ggplot2)
+library(GGally)
+library(knitr)
+library(boot)
+library(caret)
+library(randomForest)
+library(e1071)
+library(class)
+knitr::opts_chunk$set(fig.width=14, fig.height=8)
+set.seed(420)
 ```
 
-# Breast Cancer Diagnosis Classification Analysis
+```{r load_data, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+# Load the data
+data <- read.csv("breast_cancer_diagnostic_data.csv") # Full Data
+training <- read.csv("training.csv") # Training Data
+test <- read.csv("test.csv")  # Test Data
+# Data Cleaning
+data <- data[,-1] # Remove ID column
+test <- test[,-1] # Remove ID column
+training <- training[,-1] # Remove ID column
+# Convert diagnosis to factor
+data$diagnosis <- as.factor(data$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+training$diagnosis <- as.factor(training$diagnosis)
+```
+
+## Background
+
+- Many people are diagnosed with breast cancer each year
+- Breast Cancer Diagnosis Data
+- Predict Diagnosis of Tumor based on features (Malignant or Benign)
+- *Modivation:* Reduce Type II Error (False-Negatives)
+  - Feature and Model Selection
+
+# Models
+
+```{r train_setup, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+train_control <- trainControl(method = "repeatedcv", summaryFunction = defaultSummary, classProbs = TRUE, number = 10, repeats = 10)
+```
+
+# KNN Model
+
+## Tuning KNN
+
+```{r knn_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+type_two_error_knn <- data.frame(k = seq(1, 100, 1), error = NA, typetwoerror = NA)
+
+for (i in 1:nrow(type_two_error_knn)) {
+    knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = type_two_error_knn$k[i]))
+    knn_pred <- predict(knn, test)
+    knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
+    type_two_error_knn$typetwoerror[i] <- knn_confusion$table[1, 2]/sum(knn_confusion$table)
+    type_two_error_knn$error[i] <- (knn_confusion$table[1, 2] + knn_confusion$table[2, 1])/sum(knn_confusion$table)
+}
+print(which.min(type_two_error_knn$typetwoerror))
+knn1_k <- which.min(type_two_error_knn$typetwoerror)
+knn_tte <- min(type_two_error_knn$typetwoerror)
+```
+
+```{r knn_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot() + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$error), size=1) + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$typetwoerror), size=1.5, color="#c82798") + labs(title = "KNN Type I and Overall Error", x = "K", y = "Error")
+```
+
+## Tuned KNN
+
+```{r tuned_knn, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+set.seed(420)
+tuned_knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = which.min(type_two_error_knn$typetwoerror)))
+knn_pred <- predict(tuned_knn, test)
+knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
+```
+
+```{r tuned_knn_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(knn_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "KNN Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+## Tuned KNN Parameters
+
+- Using all features
+- 10-fold cross validation repeated 10 times
+- $k=$ `r knn1_k` neighbors
+- Type II Error: `r knn_tte`
+- Accuracy: `r knn_confusion$overall[1]`
+
+# Random Forrest
+
+## Tuning Random Forrest
+
+```{r rf_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+type_two_error_rf <- data.frame(ntree = seq(1, 100, 10), error = NA, typetwoerror = NA)
+
+for (i in 1:nrow(type_two_error_rf)) {
+    rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "Accuracy", ntree = type_two_error_rf$ntree[i])
+    rf_pred <- predict(rf, test)
+    rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
+    type_two_error_rf$typetwoerror[i] <- rf_confusion$table[1, 2]/sum(rf_confusion$table)
+    type_two_error_rf$error[i] <- (rf_confusion$table[1, 2] + rf_confusion$table[2, 1])/sum(rf_confusion$table)
+}
+print(which.min(type_two_error_rf$typetwoerror))
+rf1_ntree <- type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)]
+rf_tte <- min(type_two_error_rf$typetwoerror)
+```
+
+```{r rf_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot() + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$error), size=1) + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$typetwoerror), size=1.5, color="#c82798") + labs(title = "Random Forrest Type I and Overall Error", x = "Number of Trees", y = "Error")
+```
+
+## Tuned Random Forrest
+
+```{r tuned_rf, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+tuned_rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "ROC", ntree = type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)])
+rf_pred_tuned <- predict(tuned_rf, test)
+rf_confusion <- confusionMatrix(rf_pred_tuned, test$diagnosis, positive = "M")
+```
+
+```{r tuned_rf_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(rf_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Random Forrest Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+## Tuned Random Forrest Parameters
+
+- Using all features
+- 10-fold cross validation repeated 10 times
+- $n=$ `r rf1_ntree` trees
+- Type II Error: `r rf_tte`
+- Accuracy: `r rf_confusion$overall[1]`
+
+## Radial Support Vector Machine
+
+```{r austin_svm, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+#Radial Support Vector Machine
+svm <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric="ROC")
+svm_pred <- predict(svm, test)
+svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive="M")
+
+svm_tte <- svm_confusion$table[1, 2]/sum(svm_confusion$table)
+```
+
+```{r austin_svm_plot, echo=FALSE, message=FALSE, warning=FALSE}
+
+ggplot(data.frame(svm_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+## Radial Support Vector Machine Parameters
+
+- Using all features
+- 10-fold cross validation repeated 10 times
+- Type II Error: `r svm_tte`
+- Accuracy: `r svm_confusion$overall[1]`
+
+## Tuned Radial Support Vector Machine
+
+```{r austin_svm2, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+tuned_svm <- tune.svm(diagnosis ~., data = training, gamma = 10^(-5:-1), cost = 10^(-3:1))
+svm2 <- svm(diagnosis ~., data = training, gamma = 0.01, cost = 10, type="C-classification", kernel="radial")
+svm2_pred <- predict(svm2, test)
+svm2_confusion <- confusionMatrix(svm2_pred, test$diagnosis, positive="M")
+
+svm2_tte <- svm2_confusion$table[1, 2]/sum(svm2_confusion$table)
+
+```
+
+```{r austin_svm_plot2, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(svm2_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Tuned Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+## Tuned Radial Support Vector Machine Parameters
+
+- Using all features
+- 10-fold cross validation repeated 10 times
+- Type II Error: `r svm2_tte`
+- Accuracy: `r svm2_confusion$overall[1]`

### 2022-12-09 ethan_analysis.Rmd

@@
 
+head(dataTot)
 ```
@@
 ```{r}
+#Train
+am_lm <- lm(area_mean ~ ., data = trainTot)
+pred_amtrain<- predict(am_lm, trainTot)
+amtrain_log <- postResample(pred_amtrain, trainTot$area_mean)
+amtrain_log
+```
+```{r}
 #Test
-am_lm <- lm(area_mean ~ ., data = trainTot)
 pred_amtest <- predict(am_lm, testTot)
@@
 #summary(am_lm)
-```
-
-```{r}
-#Train
-pred_amtrain<- predict(am_lm, trainTot)
-amtrain_log <- postResample(pred_amtrain, trainTot$area_mean)
-amtrain_log
 ```
@@
   RMSE = RMSE,
-  Rsquare = R_square)
+  Rsquared = R_square)
 
@@
 aml_test <- eval_results(y_test, predictions_test, testTot)
+
+print(aml_train)
+print(aml_test)
+```
+
+```{r}
+as_data_frame(rbind(amtrain_log,
+      amr_train,
+      aml_train,
+      amtest_log,
+      amr_test,
+      aml_test)) %>%
+    mutate(model = c('Linear Train Error',  'Ridge Train Error', 'Lasso Train Error',  'Linear Test Error','Ridge Test Error',  'Lasso Test Error')) %>%
+    select(model, RMSE, Rsquared)
 ```
@@
 awl_test <- eval_results(y_test, predictions_test, testTot)
+
+print(awl_train)
+print(awl_test)
 ```
+
+```{r}
+as_data_frame(rbind(awtrain_log,
+      awr_train,
+      awl_train,
+      awtest_log,
+      awr_test,
+      awl_test)) %>%
+    mutate(model = c('Linear Train Error',  'Ridge Train Error', 'Lasso Train Error',  'Linear Test Error','Ridge Test Error',  'Lasso Test Error')) %>%
+    select(model, RMSE, Rsquared)
+```
+--------------------------------------------------------------------------------
+## concave.points_mean
+
+```{r}
+#Train
+cvptsm_lm <- lm(concave.points_mean ~ ., data = trainTot)
+pred_cvptsm_train<- predict(cvptsm_lm, trainTot)
+cvptsm_train_log <- postResample(pred_cvptsm_train, trainTot$concave.points_mean)
+cvptsm_train_log
+```
+```{r}
+#Test
+pred_cvptsm_test <- predict(cvptsm_lm, testTot)
+cvptsm_test_log <- postResample(pred_cvptsm_test, testTot$concave.points_mean)
+cvptsm_test_log
+#summary(am_lm)
+```
+
+# ridge regression
+```{r}
+dummy <- dummyVars(concave.points_mean~., data = dataTot[,-1])
+x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
+y_train = trainTot$concave.points_mean
+
+x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
+y_test = testTot$concave.points_mean
+
+lambdas <- 10^seq(2, -3, by = -.1)
+ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
+cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
+optimal_lambda <- cv_ridge$lambda.min
+
+eval_results <- function(true, predicted, df) {
+  SSE <- sum((predicted - true)^2)
+  SST <- sum((true - mean(true))^2)
+  R_square <- 1 - SSE / SST
+  RMSE = sqrt(SSE/nrow(df))
+
+  # Model performance metrics
+data.frame(
+  RMSE = RMSE,
+  Rsquare = R_square)
+
+}
+
+# Prediction and evaluation on train data
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
+cvptsm_train <- eval_results(y_train, predictions_train, trainTot)
+
+# Prediction and evaluation on test data
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+cvptsm_test <- eval_results(y_test, predictions_test, testTot)
+
+print(cvptsm_train)
+print(cvptsm_test)
+```
+
+# lasso
+```{r}
+lambdas <- 10^seq(2, -3, by = -.1)
+lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+
+lambda_best <- lasso_reg$lambda.min 
+lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
+cvptsm_train <- eval_results(y_train, predictions_train, trainTot)
+
+predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
+cvptsm_test <- eval_results(y_test, predictions_test, testTot)
+
+print(cvptsm_train)
+print(cvptsm_test)
+```
+
+--------------------------------------------------------------------------------
+
+## Radius Worst
+
+```{r}
+#Train
+rw_lm <- lm(radius_worst ~ ., data = trainTot)
+pred_rw_train<- predict(rw_lm, trainTot)
+rw_train_log <- postResample(pred_rw_train, trainTot$radius_worst)
+rw_train_log
+```
+```{r}
+#Test
+pred_rw_test <- predict(rw_lm, testTot)
+rw_test_log <- postResample(pred_rw_test, testTot$radius_worst)
+rw_test_log
+#summary(am_lm)
+```
+
+# ridge regression
+```{r}
+dummy <- dummyVars(radius_worst~., data = dataTot[,-1])
+x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
+y_train = trainTot$radius_worst
+
+x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
+y_test = testTot$radius_worst
+
+lambdas <- 10^seq(2, -3, by = -.1)
+ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
+cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
+optimal_lambda <- cv_ridge$lambda.min
+
+eval_results <- function(true, predicted, df) {
+  SSE <- sum((predicted - true)^2)
+  SST <- sum((true - mean(true))^2)
+  R_square <- 1 - SSE / SST
+  RMSE = sqrt(SSE/nrow(df))
+
+  # Model performance metrics
+data.frame(
+  RMSE = RMSE,
+  Rsquare = R_square)
+
+}
+
+# Prediction and evaluation on train data
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
+rw_train <- eval_results(y_train, predictions_train, trainTot)
+
+# Prediction and evaluation on test data
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+rw_test <- eval_results(y_test, predictions_test, testTot)
+
+print(rw_train)
+print(rw_test)
+```
+
+# lasso
+```{r}
+lambdas <- 10^seq(2, -3, by = -.1)
+lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+
+lambda_best <- lasso_reg$lambda.min 
+lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
+rw_train <- eval_results(y_train, predictions_train, trainTot)
+
+predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
+rw_test <- eval_results(y_test, predictions_test, testTot)
+
+print(rw_train)
+print(rw_test)
+```

### 2022-12-09 powerpoint.rmd

@@
 ```{r}
-ggplot(train) +
-  geom_point(aes(x = concavity_mean, y = radius_mean, col = diagnosis))+
-  labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor")
-```
-
-Different measurements of each variable are correlated
-
-Relationship with Diagnosis visible for each measurement
-
-## Classification Analysis: Austin
-
-```{r load_data, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
-# Load the data
-set.seed(420)
-data <- read.csv("breast_cancer_diagnostic_data.csv") # Full Data
-training <- read.csv("training.csv") # Training Data
-test <- read.csv("test.csv")  # Test Data
-# Data Cleaning
-data <- data[,-1] # Remove ID column
-test <- test[,-1] # Remove ID column
-training <- training[,-1] # Remove ID column
-# Convert diagnosis to factor
-data$diagnosis <- as.factor(data$diagnosis)
-test$diagnosis <- as.factor(test$diagnosis)
-training$diagnosis <- as.factor(training$diagnosis)
-```
-
-# Models
-
-```{r train_setup, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
-train_control <- trainControl(method = "repeatedcv", summaryFunction = defaultSummary, classProbs = TRUE, number = 10, repeats = 10)
-```
-
-# KNN Model
-
-## Tuning KNN
-
-```{r knn_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
-type_two_error_knn <- data.frame(k = seq(1, 100, 1), error = NA, typetwoerror = NA)
-
-for (i in 1:nrow(type_two_error_knn)) {
-    knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = type_two_error_knn$k[i]))
-    knn_pred <- predict(knn, test)
-    knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
-    type_two_error_knn$typetwoerror[i] <- knn_confusion$table[1, 2]/sum(knn_confusion$table)
-    type_two_error_knn$error[i] <- (knn_confusion$table[1, 2] + knn_confusion$table[2, 1])/sum(knn_confusion$table)
-}
-print(which.min(type_two_error_knn$typetwoerror))
-knn1_k <- which.min(type_two_error_knn$typetwoerror)
-knn_tte <- min(type_two_error_knn$typetwoerror)
-```
-
-```{r knn_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
-ggplot() + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$error), size=1) + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$typetwoerror), size=1.5, color="#c82798") + labs(title = "KNN Type I and Overall Error", x = "K", y = "Error")
-```
-
-## Tuned KNN
-
-```{r tuned_knn, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
-set.seed(420)
-tuned_knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = which.min(type_two_error_knn$typetwoerror)))
-knn_pred <- predict(tuned_knn, test)
-knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
-```
-
-```{r tuned_knn_plot, echo=FALSE, message=FALSE, warning=FALSE}
-ggplot(data.frame(knn_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
-  geom_tile() +
-  geom_text(aes(label = Freq), size = 12) +
-  scale_fill_gradient(low = "white", high = "#c82798") +
-  labs(title = "KNN Confusion Matrix", x = "Prediction", y = "Reference") +
-  theme_minimal() +
-  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
-  annotate("path",
-   x=1+.25*cos(seq(0,2*pi,length.out=100)),
-   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
-```
-
-## Tuned KNN Parameters
-
-- Using all features
-- 10-fold cross validation repeated 10 times
-- $k=$ `r knn1_k` neighbors
-- Type II Error: `r knn_tte`
-- Accuracy: `r knn_confusion$overall[1]`
-
-# Random Forrest
-
-## Tuning Random Forrest
-
-```{r rf_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
-type_two_error_rf <- data.frame(ntree = seq(1, 100, 10), error = NA, typetwoerror = NA)
-
-for (i in 1:nrow(type_two_error_rf)) {
-    rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "Accuracy", ntree = type_two_error_rf$ntree[i])
-    rf_pred <- predict(rf, test)
-    rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
-    type_two_error_rf$typetwoerror[i] <- rf_confusion$table[1, 2]/sum(rf_confusion$table)
-    type_two_error_rf$error[i] <- (rf_confusion$table[1, 2] + rf_confusion$table[2, 1])/sum(rf_confusion$table)
-}
-print(which.min(type_two_error_rf$typetwoerror))
-rf1_ntree <- type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)]
-rf_tte <- min(type_two_error_rf$typetwoerror)
-```
-
-```{r rf_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
-ggplot() + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$error), size=1) + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$typetwoerror), size=1.5, color="#c82798") + labs(title = "Random Forrest Type I and Overall Error", x = "Number of Trees", y = "Error")
-```
-
-## Tuned Random Forrest
-
-```{r tuned_rf, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
-tuned_rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "ROC", ntree = type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)])
-rf_pred_tuned <- predict(tuned_rf, test)
-rf_confusion <- confusionMatrix(rf_pred_tuned, test$diagnosis, positive = "M")
-```
-
-```{r tuned_rf_plot, echo=FALSE, message=FALSE, warning=FALSE}
-ggplot(data.frame(rf_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
-  geom_tile() +
-  geom_text(aes(label = Freq), size = 12) +
-  scale_fill_gradient(low = "white", high = "#c82798") +
-  labs(title = "Random Forrest Confusion Matrix", x = "Prediction", y = "Reference") +
-  theme_minimal() +
-  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
-  annotate("path",
-   x=1+.25*cos(seq(0,2*pi,length.out=100)),
-   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
-```
-
-## Tuned Random Forrest Parameters
-
-- Using all features
-- 10-fold cross validation repeated 10 times
-- $n=$ `r rf1_ntree` trees
-- Type II Error: `r rf_tte`
-- Accuracy: `r rf_confusion$overall[1]`
-
-## Radial Support Vector Machine
-
-```{r austin_svm, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
-#Radial Support Vector Machine
-svm <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric="ROC")
-svm_pred <- predict(svm, test)
-svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive="M")
-
-svm_tte <- svm_confusion$table[1, 2]/sum(svm_confusion$table)
-```
-
-```{r austin_svm_plot, echo=FALSE, message=FALSE, warning=FALSE}
-
-ggplot(data.frame(svm_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
-  geom_tile() +
-  geom_text(aes(label = Freq), size = 12) +
-  scale_fill_gradient(low = "white", high = "#c82798") +
-  labs(title = "Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
-  theme_minimal() +
-  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
-  annotate("path",
-   x=1+.25*cos(seq(0,2*pi,length.out=100)),
-   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
-```
-
-## Radial Support Vector Machine Parameters
-
-- Using all features
-- 10-fold cross validation repeated 10 times
-- Type II Error: `r svm_tte`
-- Accuracy: `r svm_confusion$overall[1]`
-
-## Tuned Radial Support Vector Machine
-
-```{r austin_svm2, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
-tuned_svm <- tune.svm(diagnosis ~., data = training, gamma = 10^(-5:-1), cost = 10^(-3:1))
-svm2 <- svm(diagnosis ~., data = training, gamma = 0.01, cost = 10, type="C-classification", kernel="radial")
-svm2_pred <- predict(svm2, test)
-svm2_confusion <- confusionMatrix(svm2_pred, test$diagnosis, positive="M")
-
-svm2_tte <- svm2_confusion$table[1, 2]/sum(svm2_confusion$table)
-
-```
-
-```{r austin_svm_plot2, echo=FALSE, message=FALSE, warning=FALSE}
-ggplot(data.frame(svm2_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
-  geom_tile() +
-  geom_text(aes(label = Freq), size = 12) +
-  scale_fill_gradient(low = "white", high = "#c82798") +
-  labs(title = "Tuned Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
-  theme_minimal() +
-  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
-  annotate("path",
-   x=1+.25*cos(seq(0,2*pi,length.out=100)),
-   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
-```
-
-## Tuned Radial Support Vector Machine Parameters
-
-- Using all features
-- 10-fold cross validation repeated 10 times
-- Type II Error: `r svm2_tte`
-- Accuracy: `r svm2_confusion$overall[1]`
-
-## Classification Analysis: Danny
-
-```{r}
 ### Packages
@@
 test$diagnosis <- as.factor(test$diagnosis)
-```
 
-```{r}
 ggplot(train) +
@@
   labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor")
+```
 
+## EDA
+
+```{r}
 train[,c(1,2,12,22)] |>
@@
 ```
-
-Different measurements of each variable are correlated
-
-Relationship with Diagnosis visible for each measurement
 
@@
 
+# Preferred Models
+
 ## K-Nearest Neigbors Model

### 2022-12-09 ethan_temp_powerpoint.rmd
Add files via upload
Ethan edits to powerpoint

--- 2022-12-09 ethan_temp_powerpoint.rmd
@@
 title: "Breast Cancer Diagnosis Classification Analysis"
-author: Austin Lackey, Ethan Powers and Danny Laposata
+author: "Austin Lackey, Ethan Powers and Danny Laposata"
 date: "December 9th, 2022"
@@
     reference_doc: template.pptx
-
+  slidy_presentation: default
+always_allow_html: true
 ---
-
 ```{r setup, include=FALSE}
@@
 library(repr)
+library(knitr)
+library(kableExtra)
 
 set.seed(420)
-
-dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
-datatemp2 <- dataTemp %>% select_if(~ !any(is.na(.)))
-dataTot <- datatemp2[,!names(datatemp2) %in% c("id")]
-
-testtemp <- read.csv("test.csv")
-testTot <- testtemp[,!names(testtemp) %in% c("id")]
-
-traintemp <- read.csv("training.csv")
-trainTot <- traintemp[,!names(traintemp) %in% c("id")]
-
 knitr::opts_chunk$set(fig.width=12, fig.height=8) 
@@
 
-## Features
+## Variable Descriptions
 - Radius: Average Distance from cell center to cell perimeter
@@
 
-This is my validation plot
+- Austin Top 5 Importance Variables
+    - *area_mean
+    - *area_worst
+    - *concave.points_mean
+    - texture_mean
+    - *radius_worst
+- Danny Top 5 Importance Variables
+    - concave.points_worst
+    - *concave.points_mean
+    - *area_worst
+    - *area_mean
+    - *radius_worst
+
+## Correlation Plot
+
 ```{r}
+dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
+datatemp2 <- dataTemp %>% select_if(~ !any(is.na(.)))
+dataTot <- datatemp2[,!names(datatemp2) %in% c("id")]
+
+testtemp <- read.csv("test.csv")
+testTot <- testtemp[,!names(testtemp) %in% c("id")]
+
+traintemp <- read.csv("training.csv")
+trainTot <- traintemp[,!names(traintemp) %in% c("id")]
+```
+
+```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
 dataE <- dataTot
@@
 
-## Model Assesments
+## Models
+
+- Calculating Training and Test Set Error
+    - Linear Regression
+    - Ridge Regression
+    - Lasso Regression
+
+## area_mean
+
+![Best Model was Ridge Regression](area_mean.png){width = "50%}
+
+## area_worst
+
+![Best Model was Linear Regression](area_worst.png){width = "50%}
+
+## concave.points_mean
+![Best Model was Linear Regression](concave.points_mean.png){width = "50%}
+
+## radius_worst
+![Best Model was Linear Regression](radius_worst.png){width = "50%}
 

### 2022-12-09 powerpoint.rmd
Add files via upload
Ethan's parts of powerpoint

@@
 ```
+```{r}
+#install.packages("kknn")
+```
 
@@
 
-- Data collected from Kaggle
-- Response: Diagnosis (Benign/Malignant)
-- Mean, SE, and "Worst" measurements of 10 variables
-- Training/Test Split (60/40)
+- Data collected from "Breast Cancer Wisconsin (Diagnostic) Data Set" from Kaggle
+- Response: Diagnosis
+    - Benign: (Noncancerous Cells)
+    - Malignant (Cancerous Cells)
+- Predictors: 10 variables split into 3 categories:
+    - Mean 
+    - Standard Error (SE)
+    - Worst: Mean of 3 largest values of a variable
+- Training/Test Data Split 60/40
 
-## Features
-
-- 
--
+## Variable Descriptions
+- Radius: Average Distance from cell center to cell perimeter
+- Texture: Standard deviation of gray-scale values; brightness of pixel of cell
+- Perimeter: Distance around nucleus boundary
+- Area: Area of the nucleus
+- Smoothness: Variation in cell's radial lengths
+- Compactness: The Perimeter^2/Area
+- Concavity: Size of the indention in nucleus boundary
+- Concave Points: Number of points on indented section of nucleus boundary
+- Symmetry: Deviation of the nuclei shape from the ideal measurement
+- Fractal Dimension: Measurement of irregularity in nucleus boundary
 
@@
 
+- Austin Top 5 Importance Variables
+    - *area_mean
+    - *area_worst
+    - *concave.points_mean
+    - texture_mean
+    - *radius_worst
+- Danny Top 5 Importance Variables
+    - concave.points_worst
+    - *concave.points_mean
+    - *area_worst
+    - *area_mean
+    - *radius_worst
+
+## Correlation Plot
+
+```{r}
+library(tidyverse)
+library(ggplot2)
+library(dplyr)
+library(caret)
+library(leaps)
+library(caTools)
+library(tidymodels)
+library(glmnet)
+library(pls)
+library(forecast)
+library(GGally)
+library(plyr)
+library(readr)
+library(repr)
+library(knitr)
+library(kableExtra)
+
+set.seed(420)
+```
+
+```{r}
+dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
+datatemp2 <- dataTemp %>% select_if(~ !any(is.na(.)))
+dataTot <- datatemp2[,!names(datatemp2) %in% c("id")]
+
+testtemp <- read.csv("test.csv")
+testTot <- testtemp[,!names(testtemp) %in% c("id")]
+
+traintemp <- read.csv("training.csv")
+trainTot <- traintemp[,!names(traintemp) %in% c("id")]
+```
+
+```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+dataE <- dataTot
+corr_df <- dataE[,c('area_worst', 'area_mean', 'concave.points_mean', 'radius_worst')]
+ggpairs(corr_df)
+```
+
+## Models
+
+- Calculating Training and Test Set Error
+    - Linear Regression
+    - Ridge Regression
+    - Lasso Regression
+
+## area_mean
+
+![Best Model was Ridge Regression](area_mean.png){
+
+## area_worst
+
+![Best Model was Linear Regression](area_worst.png)
+
+## concave.points_mean
+![Best Model was Linear Regression](concave.points_mean.png)
+
+## radius_worst
+![Best Model was Linear Regression](radius_worst.png)
+
 ## Model Assessments
@@
 
-- Insert link to kaggle dataset here
+- Breast Cancer Wisconsin (Diagnostic) Data Set:
+    - https://www.kaggle.com/datasets/uciml/breast-cancer-wisconsin-data?datasetId=180&sortBy=voteCount
+- Definition of Features (Variables):
+    - https://www.causeweb.org/usproc/sites/default/files/usclap/2017-2/Evaluating_Benign_and_Malignant_Breast_Cancer_Cells_from_Fine-Needle_Aspirates.pdf

### 2022-12-09 powerpoint.rmd

@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 14, fig.height = 8)
-```
-```{r}
-#install.packages("kknn")
 ```

### 2022-12-09 ethan_analysis.Rmd

@@
 #install.packages("repr")
+install.packages("kableExtra")
 ```
@@
 library(repr)
+library(knitr)
+library(kableExtra)
 
@@
 ```{r}
-as_data_frame(rbind(amtrain_log,
+table_am <- as_tibble(rbind(amtrain_log,
+      amtest_log,
       amr_train,
+      amr_test,
       aml_train,
-      amtest_log,
-      amr_test,
       aml_test)) %>%
-    mutate(model = c('Linear Train Error',  'Ridge Train Error', 'Lasso Train Error',  'Linear Test Error','Ridge Test Error',  'Lasso Test Error')) %>%
+    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
     select(model, RMSE, Rsquared)
+table_am %>% 
+  kbl(digits = 5 ) %>% 
+  kable_styling(bootstrap_options = "striped")
 ```
@@
 ```{r}
-as_data_frame(rbind(awtrain_log,
+table_aw <- as_tibble(rbind(awtrain_log,
+      awtest_log,
       awr_train,
+      awr_test,
       awl_train,
-      awtest_log,
-      awr_test,
       awl_test)) %>%
-    mutate(model = c('Linear Train Error',  'Ridge Train Error', 'Lasso Train Error',  'Linear Test Error','Ridge Test Error',  'Lasso Test Error')) %>%
+    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
     select(model, RMSE, Rsquared)
+table_aw %>% 
+  kbl(digits = 5 ) %>% 
+  kable_styling(bootstrap_options = "striped")
 ```
+
 --------------------------------------------------------------------------------
@@
 predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
-cvptsm_train <- eval_results(y_train, predictions_train, trainTot)
+cvptsmr_train <- eval_results(y_train, predictions_train, trainTot)
 
@@
 predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
-cvptsm_test <- eval_results(y_test, predictions_test, testTot)
+cvptsmr_test <- eval_results(y_test, predictions_test, testTot)
 
-print(cvptsm_train)
-print(cvptsm_test)
+print(cvptsmr_train)
+print(cvptsmr_test)
 ```
@@
 predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
-cvptsm_train <- eval_results(y_train, predictions_train, trainTot)
+cvptsml_train <- eval_results(y_train, predictions_train, trainTot)
 
 predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
-cvptsm_test <- eval_results(y_test, predictions_test, testTot)
+cvptsml_test <- eval_results(y_test, predictions_test, testTot)
 
-print(cvptsm_train)
-print(cvptsm_test)
+print(cvptsml_train)
+print(cvptsml_test)
+```
+
+```{r}
+table_am <- as_tibble(rbind(cvptsm_train_log,
+      cvptsm_test_log,
+      cvptsmr_train,
+      cvptsmr_test,
+      cvptsml_train,
+      cvptsml_test)) %>%
+    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
+    select(model, RMSE, Rsquared)
+table_am %>% 
+  kbl(digits = 5 ) %>% 
+  kable_styling(bootstrap_options = "striped")
 ```
@@
 predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
-rw_train <- eval_results(y_train, predictions_train, trainTot)
+rwr_train <- eval_results(y_train, predictions_train, trainTot)
 
@@
 predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
-rw_test <- eval_results(y_test, predictions_test, testTot)
+rwr_test <- eval_results(y_test, predictions_test, testTot)
 
-print(rw_train)
-print(rw_test)
+print(rwr_train)
+print(rwr_test)
 ```
@@
 predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
-rw_train <- eval_results(y_train, predictions_train, trainTot)
+rwl_train <- eval_results(y_train, predictions_train, trainTot)
 
 predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
-rw_test <- eval_results(y_test, predictions_test, testTot)
+rwl_test <- eval_results(y_test, predictions_test, testTot)
 
-print(rw_train)
-print(rw_test)
+print(rwl_train)
+print(rwl_test)
 ```
+
+```{r}
+table_am <- as_tibble(rbind(rw_train_log,
+      rw_test_log,
+      rwr_train,
+      rwr_test,
+      rwl_train,
+      rwl_test)) %>%
+    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
+    select(model, RMSE, Rsquared)
+table_am %>% 
+  kbl(digits = 5 ) %>% 
+  kable_styling(bootstrap_options = "striped")
+```

### 2022-12-09 powerpoint.rmd

@@
 
-- Many people are diagnosed with breast cancer each year
 - Breast Cancer Diagnosis Data
-- Predict Diagnosis of Tumor based on features (Malignant or Benign)
-- *Modivation:* Reduce Type II Error (False-Negatives)
-  - Feature and Model Selection
+- Predict Diagnosis of Tumor based on features
+- Feature Importance/Selection
+- Reduce Type II Error (False-Negatives)
 
@@
     - Malignant (Cancerous Cells)
-- Predictors: 10 variables split into 3 categories:
+- Predictors: 10 variables each with 3 measurements:
     - Mean 
@@
 ## Variable Descriptions
+
 - Radius: Average Distance from cell center to cell perimeter
@@
 
-## Model Assessments
+## Analysis Summary
+
+- Interpretation difficulties
+- High prediction accuracy
+- Successful with multiple methods of reducing Type II Error
+
+## Potential Improvements
 

### 2022-12-09 powerpoint.rmd
updated_pp

@@
   ggpairs()
+```
+
+## Diagnosis Distribution
+
+```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+# There are more benign tumors than malignant tumors
+# histogram of diagnosis
+edaData <- read_csv("breast_cancer_diagnostic_data.csv")
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(edaData) +
+  geom_bar(aes(x = diagnosis), fill = "#c82798") +
+  labs(x = "Diagnosis", y = "Count", title = "Count of Diagnosis") +
+  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30))
 ```
@@
 - Accuracy: `r rf_confusion$overall[1]`
+
+# Support Vector Machines
 
@@
 
-![Best Model was Ridge Regression](area_mean.png){
+![Best Model was Ridge Regression](area_mean.png)
 
@@
 
+- More data points (more observations)
+- More features (more variables)
+- More proccessing power
+
 ## Works Cited

### 2022-12-09 powerpoint.rmd

@@
 ```{r knn_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
-ggplot() + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$error), size=1) + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$typetwoerror), size=1.5, color="#c82798") + labs(title = "KNN Type I and Overall Error", x = "K", y = "Error")
+ggplot() + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$error), size=1) + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$typetwoerror), size=1.5, color="#c82798") + labs(title = "KNN Type II and Overall Error", x = "K", y = "Error")
 ```
@@
 ```{r rf_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
-ggplot() + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$error), size=1) + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$typetwoerror), size=1.5, color="#c82798") + labs(title = "Random Forrest Type I and Overall Error", x = "Number of Trees", y = "Error")
+ggplot() + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$error), size=1) + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$typetwoerror), size=1.5, color="#c82798") + labs(title = "Random Forrest Type II and Overall Error", x = "Number of Trees", y = "Error")
 ```
@@
 
+```{r}
+### Gathering Data
+
+train <- read.csv("training.csv")
+test <- read.csv("test.csv")
+
+train |>
+  select(-id) -> train
+
+test |>
+  select(-id) -> test
+
+train$diagnosis <- as.factor(train$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+```
+
 - Logistic Regression
@@
 - 2 False-Negatives
-- `knn.err` % prediction accuracy
+- 96.04% prediction accuracy
 
@@
 - 2 False-Negatives
-- `svm.err` % prediction accuracy
+- 96.48% prediction accuracy
 
@@
 - 1 False-Negative
-- `rf.err` % prediction accuracy
+- 95.15% prediction accuracy
 

### 2022-12-09  PPTX-code.R
delete

@@ -1,164 +1,0 @@
-### Packages
-
-library(tidyverse)
-library(GGally)
-library(ggplot2)
-library(tidyr)
-library(tidymodels)
-library(vip)
-library(caTools)
-library(xgboost)
-library(discrim)
-library(glmnet)
-library(kableExtra)
-
-### Gathering Data
-
-train <- read.csv("training.csv")
-test <- read.csv("test.csv")
-
-train |>
-  select(-id) -> train
-
-test |>
-  select(-id) -> test
-
-train$diagnosis <- as.factor(train$diagnosis)
-test$diagnosis <- as.factor(test$diagnosis)
-
-### EDA
-
-train |>
-  group_by(diagnosis) |>
-  summarise(radius = mean(radius_mean),
-            concavity = mean(concavity_mean),
-            perimeter = mean(perimeter_mean)) |>
-  kable(digits = 3) |>
-  kable_material(lightable_options = "striped")
-
-ggplot(train) +
-  geom_point(aes(x = concavity_mean, y = radius_mean, col = diagnosis))+
-  labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor") -> plot1
-
-train[,1:11] |>
-  ggpairs()
-
-train[,c(1,12:21)] |>
-  ggpairs()
-
-train[,c(1,22:31)] |>
-  ggpairs()
-
-train[,c(1,2,12,22)] |>
-  ggpairs() -> plot2
-
-### Models
-
-### SVC
-
-df_cost <- grid_regular(cost(), levels = 10)
-
-svm_linear_tune_spec <- svm_poly(degree = 1, cost = tune("cost")) %>%
-  set_mode("classification") %>%
-  set_engine("kernlab", scaled = FALSE)
-
-svm_linear_rec <- recipe(diagnosis ~ ., data = train)
-
-svm_linear_wf <- workflow() |>
-  add_model(svm_linear_tune_spec) |>
-  add_recipe(svm_linear_rec)
-
-tune_fit <- svm_linear_wf |>
-  tune_grid(resamples = cancer_10foldcv, grid = df_cost)
-
-autoplot(tune_fit)
-show_best(tune_fit, metric = "accuracy", n = 1)
-
[…]

### 2022-12-10 austin_temp_paper.rmd

---
title: "Breast Cancer Diagnosis Classification Analysis"
author: AUSTIN'S TEMPLATE
date: "December 9th, 2022"
output: 
  pdf_document:
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
    df_print: paged
    fig_caption: TRUE
    highlight: tango
---
```{r setup, include=FALSE}
# Import packages
library(tidyverse)
library(ggplot2)
library(GGally)
library(knitr)
library(boot)
library(caret)
library(randomForest)
library(e1071)
library(class)
set.seed(420)
```

# Background (Should we do an Abstract?)

*Insert text here*

# Data

*Insert text here*

## Variable Descriptions

*Insert text here*

# Explatory Data Analysis

*Insert text here*

# Austin's Classification Proccess and Analysis

*Insert text here*

# Dannys's Classification Proccess and Analysis

*Insert text here*

# Ethans's Regression Proccess and Analysis

*Insert text here*

# Analysis Summary

*Insert text here*

# Potential Improvements

*Insert text here*

# Works Cited

- Breast Cancer Wisconsin (Diagnostic) Data Set:
    - https://www.kaggle.com/datasets/uciml/breast-cancer-wisconsin-data?datasetId=180&sortBy=voteCount
- Definition of Features (Variables):
    - https://www.causeweb.org/usproc/sites/default/files/usclap/2017-2/Evaluating_Benign_and_Malignant_Breast_Cancer_Cells_from_Fine-Needle_Aspirates.pdf

### 2022-12-10 dannys_temp_paper.rmd

--- 2022-12-10 austin_temp_paper.rmd
@@
 title: "Breast Cancer Diagnosis Classification Analysis"
-author: AUSTIN'S TEMPLATE
+author: DANNY's TEMPLATE
 date: "December 9th, 2022"

### 2022-12-10 ethans_temp_paper.rmd

--- 2022-12-10 dannys_temp_paper.rmd
@@
 title: "Breast Cancer Diagnosis Classification Analysis"
-author: DANNY's TEMPLATE
+author: ETHANS'S TEMPLATE
 date: "December 9th, 2022"

### 2022-12-10 austin_temp_paper.rmd

@@
 set.seed(420)
+data <- read.csv("breast_cancer_diagnostic_data.csv") # Full Data
+training <- read.csv("training.csv") # Training Data
+test <- read.csv("test.csv")  # Test Data
 ```
@@
 
-*Insert text here*
+In our analysis, we will be using the Breast Cancer Wisconsin (Diagnostic) Data Set from Kaggle. 
+This data set contains 569 observations of breast cancer cells with 32 variables describing each cell. 
+Some of the variables include characteristics like radius, texture, area, perimeter of the cell.
 

### 2022-12-11 austin_temp_paper.rmd
updatedpaper

--- 2022-12-10 austin_temp_paper.rmd
@@
 set.seed(420)
+# Import data
+data <- read.csv("breast_cancer_diagnostic_data.csv") # Full Data
+training <- read.csv("training.csv") # Training Data
+test <- read.csv("test.csv")  # Test Data
+# Data Cleaning
+data <- data[,-1] # Remove ID column
+test <- test[,-1] # Remove ID column
+training <- training[,-1] # Remove ID column
+# Convert diagnosis to factor
+data$diagnosis <- as.factor(data$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+training$diagnosis <- as.factor(training$diagnosis)
 ```
 
-# Background (Should we do an Abstract?)
+# Introduction (Should we do an Abstract?)
 
-*Insert text here*
+## Background
 
-# Data
+In our analysis, we will be using the Breast Cancer Wisconsin (Diagnostic) Data Set from Kaggle. 
+This data set contains `569 observations` of breast cancer cells with `32 variables` describing each cell. 
+Some of the variables include characteristics like `radius`, `texture`, `area`, `perimeter` of the cell.
 
-*Insert text here*
+The goal of our analysis is to predict whether a cell is benign or malignant based on the `32 variables`.
+A cell is considered benign if it is not cancerous and malignant if it is cancerous.
+Normally in most machine learning models, we do our best to train the model to reduce the overall error rate.
+While this is an important goal, our group was more concerned with the `Type-II error rate`.
+By reducing the `Type-II error rate`, we can ensure that we are not making the mistake of classifying a malignant cell as benign.
+This is important in the world of Oncology because if a malignant cell is classified as benign, it could lead to a patient not receiving the proper treatment.
+Whereas if a benign cell is classified as malignant, the patient may be alarmed, but a false alarm is better than a missed diagnosis.
+
+In order to achieve our goal, we conducted the following steps:
+
+1. Data Cleaning
+2. Explatory Data Analysis
+3. Classification Analysis
+4. Regression Analysis
+5. Overall Analysis Summary
+
+## Data
+
+In order to properly train and test our models, we first had to split the data into a training and test set.
+We decided to use a `60/40 split` for our training and test data. This allows us to allocate more data to the training set, which will allow us to train our models more effectively.
+We also decided to remove the `ID` column from the data set because it was not relevant to our analysis.
 
@@
 
-*Insert text here*
+To begin our analysis, we first wanted to get a better understanding of the data so we could properly prepare it for our models.
+As you can see in Figure A below, we have more information regarding the benign cells than the malignant cells.
+To be more specific, we have 357 benign cells and 212 malignant cells.
+Since we are worried about the `Type-II error rate`, we would want more information regarding the malignant cells in an ideal world.
+However since this is not the case, is important to note because it could lead to a bias in our model if we do not take this into account.
 
-# Austin's Classification Proccess and Analysis
+```{r, fig.width= 4, fig.height= 4, echo=FALSE, message=FALSE, warning=FALSE}
+# Malignant vs Benign Pie Chart
+pieChart <- ggplot(data, aes(x = "", fill = diagnosis)) + 
+  geom_bar(width = 1) + 
+  coord_polar("y", start = 0) + 
+  labs(title = "Figure A: Malignant vs Benign Cell Counts", x= "", y="", fill="Diagnosis") +  
+  geom_text(stat='count', aes(label=..count..), vjust=-3, size=5)
+print(pieChart)
+```
+
+*Insert text here regarding the GGPairs Plot*
+
+```{r, fig.width= 6, fig.height= 6, echo=FALSE, message=FALSE, warning=FALSE}
+training[,c(1,2,12,22)] |>
+  ggpairs(aes(color = diagnosis)) + labs(title = "Figure B: Pairwise Scatterplots of Features")
+```
+
+# Classification Analysis
+
+During our classification analysis, we attacked `Type-II error` in two different ways.
+Austin used different paramters for each model and to figure out which parameters yield the lowest `Type-II error rate`. 
+However, testing many different parameters can be time consuming and computationally expensive. 
+Danny used a different approach by tuning his models to be more sensitive to Malignant cells. 
+This allows him to reduce the `Type-II error rate` by classifying more cells as Malignant. 
+However, on the downside, this also increases the `Type-I error rate` by classifying more cells as Malignant. 
+By harnessing both approaches, we were able to reduce the `Type-II error rate` to its fullest while also maintaining a good `Type-I error rate`.
+
+## Austin's Classification Proccess and Analysis
+
+```{r load_data, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+# Load the data
+set.seed(420)
+data <- read.csv("breast_cancer_diagnostic_data.csv") # Full Data
+training <- read.csv("training.csv") # Training Data
+test <- read.csv("test.csv")  # Test Data
+# Data Cleaning
+data <- data[,-1] # Remove ID column
+test <- test[,-1] # Remove ID column
+training <- training[,-1] # Remove ID column
+# Convert diagnosis to factor
+data$diagnosis <- as.factor(data$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+training$diagnosis <- as.factor(training$diagnosis)
+```
+
+```{r train_setup, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+train_control <- trainControl(method = "repeatedcv", summaryFunction = defaultSummary, classProbs = TRUE, number = 10, repeats = 10)
+```
+
+For all of models that Austin used were tuned using a `10-fold cross validation` method that was repeated 10 times. 
+Cross validation is method that is used to train a model on a subset of the data and then test the model on the remaining data. 
+This process is repeated `10` times with each subset of data being used as the test set once. 
+By using this method we are able to effectively train our model while also testing it on data that it has not seen before.
+
+### KNN Model
+
+The first model that we used was a `K Nearest Neighbor's model`. In order to tune this model, we used the `tuneGrid` parameter to test different values of `k`.
+We tested `100` values of `k` ranging from `1` to `100` and then plotted the `Type-II error rate` (in purple) as well as the overall error rate (in black) for each value of `k` as shown in Figure C below.
+The model with the lowest `Type-II error rate` was the model with `k = 10`. Any value of `k` greater than `10` resulted in a higher `Type-II error rate` as well as a higher overall error rate.
+This can be attributed to the fact that the model is overfitting the data.
+
+```{r knn_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+type_two_error_knn <- data.frame(k = seq(1, 100, 1), error = NA, typetwoerror = NA)
+
+for (i in 1:nrow(type_two_error_knn)) {
+    knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = type_two_error_knn$k[i]))
+    knn_pred <- predict(knn, test)
+    knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
+    type_two_error_knn$typetwoerror[i] <- knn_confusion$table[1, 2]/sum(knn_confusion$table)
+    type_two_error_knn$error[i] <- (knn_confusion$table[1, 2] + knn_confusion$table[2, 1])/sum(knn_confusion$table)
+}
+print(which.min(type_two_error_knn$typetwoerror))
+knn1_k <- which.min(type_two_error_knn$typetwoerror)
+knn_tte <- min(type_two_error_knn$typetwoerror)
+```
+
+```{r knn_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot() + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$error), size=1) + 
+  geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$typetwoerror), size=1.5, color="#c82798") + 
+  labs(title = "Figure C: KNN Type II and Overall Error", x = "K", y = "Error")
+```
+
+### Tuned KNN Model
+
+```{r tuned_knn, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+set.seed(420)
+tuned_knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = which.min(type_two_error_knn$typetwoerror)))
+knn_pred <- predict(tuned_knn, test)
+knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
+```
+
+Using the information from the previous plot, we were able to tune our model to have a `k` value of `10`. 
+This tuned model resulted in a `Type-II error rate` of `r knn_tte` and an overall accuracy of `r knn_confusion$overall[1]`.
+The confusion matrix for this model is shown in Figure D below. 
+As you can see out of the total `r dim(training)[1]` training samples, `r knn_confusion$table[1, 2]` were misclassified as benign when they were actually malignant.
+
+```{r tuned_knn_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(knn_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq)) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Figure D: KNN Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+### Random Forrest Model
+
+```{r rf_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+type_two_error_rf <- data.frame(ntree = seq(1, 100, 10), error = NA, typetwoerror = NA)
+
+for (i in 1:nrow(type_two_error_rf)) {
+    rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "Accuracy", ntree = type_two_error_rf$ntree[i])
+    rf_pred <- predict(rf, test)
+    rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
+    type_two_error_rf$typetwoerror[i] <- rf_confusion$table[1, 2]/sum(rf_confusion$table)
+    type_two_error_rf$error[i] <- (rf_confusion$table[1, 2] + rf_confusion$table[2, 1])/sum(rf_confusion$table)
+}
+print(which.min(type_two_error_rf$typetwoerror))
+rf1_ntree <- type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)]
+rf_tte <- min(type_two_error_rf$typetwoerror)
+```
+
+The second model that we used was a `Random Forrest model`. In order to tune this model, we used the `tuneGrid` parameter to test different numbers of `trees`.
+We tested `10` values of `trees` ranging from `1` to `100` and then plotted the `Type-II error rate` (in purple) as well as the overall error rate (in black) for each value of `trees` as shown in Figure E below.
+The model with the lowest `Type-II error rate` was the model with `trees =` `r type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)]`. 
+Any value of `trees` greater than `r type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)]` resulted in a higher `Type-II error rate` or a higher computation time for the same `Type-II error rate`.
+
+```{r rf_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot() + 
+  geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$error), size=1) + 
+  geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$typetwoerror), size=1.5, color="#c82798") + 
+  labs(title = "Figure E: Random Forrest Type II and Overall Error", x = "Number of Trees", y = "Error")
+```
+
+### Tuned Random Forrest
+
+Using the information from the previous plot, we were able to tune our model to have a `trees` value of `r type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)]`.
+This tuned model resulted in a `Type-II error rate` of `r rf_tte` and an overall accuracy of `r rf_confusion$overall[1]`.
+The confusion matrix for this model is shown in Figure F below.
+As you can see out of the total `r dim(training)[1]` training samples, `r rf_confusion$table[1, 2]` were misclassified as benign when they were actually malignant.
+And we were able to further reduce our `Type-II error rate` from `r knn_tte` to `r rf_tte` when compared to the KNN model.
+
+```{r tuned_rf, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+tuned_rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "ROC", ntree = type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)])
+rf_pred_tuned <- predict(tuned_rf, test)
+rf_confusion <- confusionMatrix(rf_pred_tuned, test$diagnosis, positive = "M")
+```
+
+```{r tuned_rf_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(rf_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq)) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Figure F: Random Forrest Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+### Radial Support Vector Machine
+
+```{r austin_svm, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+#Radial Support Vector Machine
+svm <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric="ROC")
+svm_pred <- predict(svm, test)
+svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive="M")
+
+svm_tte <- svm_confusion$table[1, 2]/sum(svm_confusion$table)
+```
+
+The third model that we used was a `Radial Support Vector Machine`.
+To begin we used a basic model with the default parameters and then tuned the model using the `tuneGrid` parameter.
+The basic SVM model resulted in a `Type-II error rate` of `r svm_tte` and an overall accuracy of `r svm_confusion$overall[1]`.
+The confusion matrix for this model is shown in Figure G below.
+As you can see out of the total `r dim(training)[1]` training samples, `r svm_confusion$table[1, 2]` was misclassified as benign when they it was actually malignant.
+This is a great improvement in our `Type-II error rate` when compared to the other models.
+
+```{r austin_svm_plot, echo=FALSE, message=FALSE, warning=FALSE}
+
+ggplot(data.frame(svm_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq)) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Figure G: Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+### Tuned Radial Support Vector Machine
+
+```{r austin_svm2, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+tuned_svm <- tune.svm(diagnosis ~., data = training, gamma = 10^(-5:-1), cost = 10^(-3:1))
+svm2 <- svm(diagnosis ~., data = training, gamma = 0.01, cost = 10, type="C-classification", kernel="radial")
+svm2_pred <- predict(svm2, test)
+svm2_confusion <- confusionMatrix(svm2_pred, test$diagnosis, positive="M")
+
+svm2_tte <- svm2_confusion$table[1, 2]/sum(svm2_confusion$table)
+
+```
+
+The final model that we used was a tuned `Radial Support Vector Machine`.
+The tuned model resulted in a `Type-II error rate` of `r svm2_tte` and an overall accuracy of `r svm2_confusion$overall[1]`.
+The confusion matrix for this model is shown in Figure H below.
+Out of the total `r dim(training)[1]` training samples, `r svm2_confusion$table[1, 2]` were misclassified as benign when they were actually malignant.
+As you can see, our `Type-II error rate` was actually increased to `r svm2_tte` when compared to the basic SVM model.
+Since this svm model was tuned using carets `tune.svm` function, the overall error was reduced, but this resulted in a slightly higher `Type-II error rate`.
+For this reason, our basic model is actually better than the tuned model when it comes to meeting our goal of reducing the `Type-II error rate`.
+
+```{r austin_svm_plot2, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(svm2_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq)) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Figure H: Tuned Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
 
@@
 
-# Dannys's Classification Proccess and Analysis
+## Dannys's Classification Proccess and Analysis
 
@@
 
-*Insert text here*
+Following our analysis, we have identified a few potential improvements that could be made to our analysis to further improve our models.
+The obvious and first improvement that could be made is to **collect more data**. With more data, we would allow our models to see more examples of malignant cells and be able to better classify them.
+As you saw in the Exploratory Data Analysis section, we have more more information regarding the benign cells than the malignant cells. 
+If we had more data, we could balance out the data set and have more information regarding the malignant cells.
+Another possible improvement is to **collect different or more features** regarding each cell. 
+Most of the features included the mean, worst and standard error of geometry measurements of the cell.
+This leads to many variables being highly correlated with each other and could lead to multicollinearity.
+If we had more features, we could reduce the multicollinearity and diversify the information we have regarding each cell.
+Finally, if we had **access to better technology** and more time or money, we could use a more advanced machine learning model.
+Because this project was done for a class, we were limited to the models we could use as there was a time contraint, especially in a team enviroment.
 

### 2022-12-15 paper.rmd
Update paper.rmd
updated-paper

@@
 ---
-title: "Paper"
-author: "Austin Lackey, Ethan Powers and Danny Laposata"
+title: "Breast Cancer Diagnosis Classification Analysis"
+author: Austin Lackey, Ethan Powers and Danny Laposata
+date: "December 9th, 2022"
 output: 
-  pdf_document: default
-  html_document:
+  pdf_document:
+    toc: TRUE
+    toc_depth: 2
+    number_sections: TRUE
     df_print: paged
+    fig_caption: TRUE
+    highlight: tango
 ---
-
-## Sample Output
-```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+```{r setup, include=FALSE}
+# Import packages
 library(tidyverse)
 library(ggplot2)
-library(dplyr)
+library(GGally)
+library(knitr)
+library(boot)
 library(caret)
+library(randomForest)
+library(e1071)
+library(class)
 set.seed(420)
-data <- read.csv("breast_cancer_diagnostic_data.csv")
-data <- subset(data, select = -c(X))
-split <- sample(c(rep(0, 0.6 * nrow(data)), rep(1, 0.4 * nrow(data))))
-train <- data[split == 0, ]
-test <- data[split== 1, ]
-#write.csv(train,"training.csv", row.names = FALSE)
-#write.csv(test,"test.csv", row.names = FALSE)
+# Import data
+data <- read.csv("breast_cancer_diagnostic_data.csv") # Full Data
+training <- read.csv("training.csv") # Training Data
+test <- read.csv("test.csv")  # Test Data
+# Data Cleaning
+data <- data[,-1] # Remove ID column
+test <- test[,-1] # Remove ID column
+training <- training[,-1] # Remove ID column
+# Convert diagnosis to factor
+data$diagnosis <- as.factor(data$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+training$diagnosis <- as.factor(training$diagnosis)
 ```
 
-```{r}
-print(names(train))
+# Introduction (Should we do an Abstract?)
+
+## Background
+
+In our analysis, we will be using the Breast Cancer Wisconsin (Diagnostic) Data Set from Kaggle. 
+This data set contains `569 observations` of breast cancer cells with `32 variables` describing each cell. 
+Some of the variables include characteristics like `radius`, `texture`, `area`, `perimeter` of the cell.
+
+The goal of our analysis is to predict whether a cell is benign or malignant based on the `32 variables`.
+A cell is considered benign if it is not cancerous and malignant if it is cancerous.
+Normally in most machine learning models, we do our best to train the model to reduce the overall error rate.
+While this is an important goal, our group was more concerned with the `Type-II error rate`.
+By reducing the `Type-II error rate`, we can ensure that we are not making the mistake of classifying a malignant cell as benign.
+This is important in the world of Oncology because if a malignant cell is classified as benign, it could lead to a patient not receiving the proper treatment.
+Whereas if a benign cell is classified as malignant, the patient may be alarmed, but a false alarm is better than a missed diagnosis.
+
+In order to achieve our goal, we conducted the following steps:
+
+1. Data Cleaning
+2. Explatory Data Analysis
+3. Classification Analysis
+4. Regression Analysis
+5. Overall Analysis Summary
+
+## Data
+
+In order to properly train and test our models, we first had to split the data into a training and test set.
+We decided to use a `60/40 split` for our training and test data. This allows us to allocate more data to the training set, which will allow us to train our models more effectively.
+We also decided to remove the `ID` column from the data set because it was not relevant to our analysis.
+
+## Variable Descriptions
+
+*Insert text here*
+
+# Explatory Data Analysis
+
+To begin our analysis, we first wanted to get a better understanding of the data so we could properly prepare it for our models.
+As you can see in Figure A below, we have more information regarding the benign cells than the malignant cells.
+To be more specific, we have 357 benign cells and 212 malignant cells.
+Since we are worried about the `Type-II error rate`, we would want more information regarding the malignant cells in an ideal world.
+However since this is not the case, is important to note because it could lead to a bias in our model if we do not take this into account.
+
+```{r, fig.width= 4, fig.height= 4, echo=FALSE, message=FALSE, warning=FALSE}
+# Malignant vs Benign Pie Chart
+pieChart <- ggplot(data, aes(x = "", fill = diagnosis)) + 
+  geom_bar(width = 1) + 
+  coord_polar("y", start = 0) + 
+  labs(title = "Figure A: Malignant vs Benign Cell Counts", x= "", y="", fill="Diagnosis") +  
+  geom_text(stat='count', aes(label=..count..), vjust=-3, size=5)
+print(pieChart)
 ```
+
+*Insert text here regarding the GGPairs Plot*
+
+```{r, fig.width= 6, fig.height= 6, echo=FALSE, message=FALSE, warning=FALSE}
+training[,c(1,2,12,22)] |>
+  ggpairs(aes(color = diagnosis)) + labs(title = "Figure B: Pairwise Scatterplots of Features")
+```
+
+# Classification Analysis
+
+During our classification analysis, we attacked `Type-II error` in two different ways.
+The first thing we did was use different paramters for each model and to figure out which parameters yield the lowest `Type-II error rate`. 
+However, testing many different parameters can be time consuming and computationally expensive. 
+We also did a different approach by tuning his models to be more sensitive to Malignant cells. 
+This allows us to reduce the `Type-II error rate` by classifying more cells as Malignant. 
+However, on the downside, this also increases the `Type-I error rate` by classifying more cells as Malignant. 
+By harnessing both approaches, we were able to reduce the `Type-II error rate` to its fullest while also maintaining a good `Type-I error rate`.
+
+## Austin's Classification Proccess and Analysis
+
+```{r load_data, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+# Load the data
+set.seed(420)
+data <- read.csv("breast_cancer_diagnostic_data.csv") # Full Data
+training <- read.csv("training.csv") # Training Data
+test <- read.csv("test.csv")  # Test Data
+# Data Cleaning
+data <- data[,-1] # Remove ID column
+test <- test[,-1] # Remove ID column
+training <- training[,-1] # Remove ID column
+# Convert diagnosis to factor
+data$diagnosis <- as.factor(data$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+training$diagnosis <- as.factor(training$diagnosis)
+```
+
+```{r train_setup, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+train_control <- trainControl(method = "repeatedcv", summaryFunction = defaultSummary, classProbs = TRUE, number = 10, repeats = 10)
+```
+
+The following models used were tuned using a `10-fold cross validation` method that was repeated 10 times. 
+Cross validation is method that is used to train a model on a subset of the data and then test the model on the remaining data. 
+This process is repeated `10` times with each subset of data being used as the test set once. 
+By using this method we are able to effectively train our model while also testing it on data that it has not seen before.
+
+### KNN Model
+
+The first model that we used was a `K Nearest Neighbor's model`. In order to tune this model, we used the `tuneGrid` parameter to test different values of `k`.
+We tested `100` values of `k` ranging from `1` to `100` and then plotted the `Type-II error rate` (in purple) as well as the overall error rate (in black) for each value of `k` as shown in Figure C below.
+The model with the lowest `Type-II error rate` was the model with `k = 10`. Any value of `k` greater than `10` resulted in a higher `Type-II error rate` as well as a higher overall error rate.
+This can be attributed to the fact that the model is overfitting the data.
+
+```{r knn_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+type_two_error_knn <- data.frame(k = seq(1, 100, 1), error = NA, typetwoerror = NA)
+
+for (i in 1:nrow(type_two_error_knn)) {
+    knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = type_two_error_knn$k[i]))
+    knn_pred <- predict(knn, test)
+    knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
+    type_two_error_knn$typetwoerror[i] <- knn_confusion$table[1, 2]/sum(knn_confusion$table)
+    type_two_error_knn$error[i] <- (knn_confusion$table[1, 2] + knn_confusion$table[2, 1])/sum(knn_confusion$table)
+}
+print(which.min(type_two_error_knn$typetwoerror))
+knn1_k <- which.min(type_two_error_knn$typetwoerror)
+knn_tte <- min(type_two_error_knn$typetwoerror)
+```
+
+```{r knn_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot() + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$error), size=1) + 
+  geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$typetwoerror), size=1.5, color="#c82798") + 
+  labs(title = "Figure C: KNN Type II and Overall Error", x = "K", y = "Error")
+```
+
+### Tuned KNN Model
+
+```{r tuned_knn, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+set.seed(420)
+tuned_knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = which.min(type_two_error_knn$typetwoerror)))
+knn_pred <- predict(tuned_knn, test)
+knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
+```
+
+Using the information from the previous plot, we were able to tune our model to have a `k` value of `10`. 
+This tuned model resulted in a `Type-II error rate` of `r knn_tte` and an overall accuracy of `r knn_confusion$overall[1]`.
+The confusion matrix for this model is shown in Figure D below. 
+As you can see out of the total `r dim(training)[1]` training samples, `r knn_confusion$table[1, 2]` were misclassified as benign when they were actually malignant.
+
+```{r tuned_knn_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(knn_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq)) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Figure D: KNN Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+### Random Forrest Model
+
+```{r rf_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+type_two_error_rf <- data.frame(ntree = seq(1, 100, 10), error = NA, typetwoerror = NA)
+
+for (i in 1:nrow(type_two_error_rf)) {
+    rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "Accuracy", ntree = type_two_error_rf$ntree[i])
+    rf_pred <- predict(rf, test)
+    rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
+    type_two_error_rf$typetwoerror[i] <- rf_confusion$table[1, 2]/sum(rf_confusion$table)
+    type_two_error_rf$error[i] <- (rf_confusion$table[1, 2] + rf_confusion$table[2, 1])/sum(rf_confusion$table)
+}
+print(which.min(type_two_error_rf$typetwoerror))
+rf1_ntree <- type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)]
+rf_tte <- min(type_two_error_rf$typetwoerror)
+```
+
+The second model that we used was a `Random Forrest model`. In order to tune this model, we used the `tuneGrid` parameter to test different numbers of `trees`.
+We tested `10` values of `trees` ranging from `1` to `100` and then plotted the `Type-II error rate` (in purple) as well as the overall error rate (in black) for each value of `trees` as shown in Figure E below.
+The model with the lowest `Type-II error rate` was the model with `trees =` `r type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)]`. 
+Any value of `trees` greater than `r type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)]` resulted in a higher `Type-II error rate` or a higher computation time for the same `Type-II error rate`.
+
+```{r rf_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot() + 
+  geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$error), size=1) + 
+  geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$typetwoerror), size=1.5, color="#c82798") + 
+  labs(title = "Figure E: Random Forrest Type II and Overall Error", x = "Number of Trees", y = "Error")
+```
+
+### Tuned Random Forrest
+
+Using the information from the previous plot, we were able to tune our model to have a `trees` value of `r type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)]`.
+This tuned model resulted in a `Type-II error rate` of `r rf_tte` and an overall accuracy of `r rf_confusion$overall[1]`.
+The confusion matrix for this model is shown in Figure F below.
+As you can see out of the total `r dim(training)[1]` training samples, `r rf_confusion$table[1, 2]` were misclassified as benign when they were actually malignant.
+And we were able to further reduce our `Type-II error rate` from `r knn_tte` to `r rf_tte` when compared to the KNN model.
+
+```{r tuned_rf, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+tuned_rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "ROC", ntree = type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)])
+rf_pred_tuned <- predict(tuned_rf, test)
+rf_confusion <- confusionMatrix(rf_pred_tuned, test$diagnosis, positive = "M")
+```
+
+```{r tuned_rf_plot, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(rf_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq)) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Figure F: Random Forrest Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+### Radial Support Vector Machine
+
+```{r austin_svm, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+#Radial Support Vector Machine
+svm <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric="ROC")
+svm_pred <- predict(svm, test)
+svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive="M")
+
+svm_tte <- svm_confusion$table[1, 2]/sum(svm_confusion$table)
+```
+
+The third model that we used was a `Radial Support Vector Machine`.
+To begin we used a basic model with the default parameters and then tuned the model using the `tuneGrid` parameter.
+The basic SVM model resulted in a `Type-II error rate` of `r svm_tte` and an overall accuracy of `r svm_confusion$overall[1]`.
+The confusion matrix for this model is shown in Figure G below.
+As you can see out of the total `r dim(training)[1]` training samples, `r svm_confusion$table[1, 2]` was misclassified as benign when they it was actually malignant.
+This is a great improvement in our `Type-II error rate` when compared to the other models.
+
+```{r austin_svm_plot, echo=FALSE, message=FALSE, warning=FALSE}
+
+ggplot(data.frame(svm_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq)) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Figure G: Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+### Tuned Radial Support Vector Machine
+
+```{r austin_svm2, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+tuned_svm <- tune.svm(diagnosis ~., data = training, gamma = 10^(-5:-1), cost = 10^(-3:1))
+svm2 <- svm(diagnosis ~., data = training, gamma = 0.01, cost = 10, type="C-classification", kernel="radial")
+svm2_pred <- predict(svm2, test)
+svm2_confusion <- confusionMatrix(svm2_pred, test$diagnosis, positive="M")
+
+svm2_tte <- svm2_confusion$table[1, 2]/sum(svm2_confusion$table)
+
+```
+
+The final model that we used was a tuned `Radial Support Vector Machine`.
+The tuned model resulted in a `Type-II error rate` of `r svm2_tte` and an overall accuracy of `r svm2_confusion$overall[1]`.
+The confusion matrix for this model is shown in Figure H below.
+Out of the total `r dim(training)[1]` training samples, `r svm2_confusion$table[1, 2]` were misclassified as benign when they were actually malignant.
+As you can see, our `Type-II error rate` was actually increased to `r svm2_tte` when compared to the basic SVM model.
+Since this svm model was tuned using carets `tune.svm` function, the overall error was reduced, but this resulted in a slightly higher `Type-II error rate`.
+For this reason, our basic model is actually better than the tuned model when it comes to meeting our goal of reducing the `Type-II error rate`.
+
+```{r austin_svm_plot2, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(data.frame(svm2_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq)) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Figure H: Tuned Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
+  theme_minimal() +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
+```
+
+*Insert text here*
+
+## Dannys's Classification Proccess and Analysis
+
+*Insert text here*
+
+# Ethans's Regression Proccess and Analysis
+
+*Insert text here*
+
+# Analysis Summary
+
+*Insert text here*
+
+# Potential Improvements
+
+Following our analysis, we have identified a few potential improvements that could be made to our analysis to further improve our models.
+The obvious and first improvement that could be made is to **collect more data**. With more data, we would allow our models to see more examples of malignant cells and be able to better classify them.
+As you saw in the Exploratory Data Analysis section, we have more more information regarding the benign cells than the malignant cells. 
+If we had more data, we could balance out the data set and have more information regarding the malignant cells.
+Another possible improvement is to **collect different or more features** regarding each cell. 
+Most of the features included the mean, worst and standard error of geometry measurements of the cell.
+This leads to many variables being highly correlated with each other and could lead to multicollinearity.
+If we had more features, we could reduce the multicollinearity and diversify the information we have regarding each cell.
+Finally, if we had **access to better technology** and more time or money, we could use a more advanced machine learning model.
+Because this project was done for a class, we were limited to the models we could use as there was a time contraint, especially in a team enviroment.
+
+# Works Cited
+
+- Breast Cancer Wisconsin (Diagnostic) Data Set:
+    - https://www.kaggle.com/datasets/uciml/breast-cancer-wisconsin-data?datasetId=180&sortBy=voteCount
+- Definition of Features (Variables):
+    - https://www.causeweb.org/usproc/sites/default/files/usclap/2017-2/Evaluating_Benign_and_Malignant_Breast_Cancer_Cells_from_Fine-Needle_Aspirates.pdf

### 2022-12-15 ethans_temp_paper.rmd
Add files via upload
Ethan part of paper

--- 2022-12-10 dannys_temp_paper.rmd
@@
     highlight: tango
+    extra_dependencies: ["float"]
+
 ---
@@
 library(class)
+#
+library(dplyr)
+library(leaps)
+library(caTools)
+library(tidymodels)
+library(glmnet)
+library(forecast)
+library(plyr)
+library(readr)
+library(repr)
+library(kableExtra)
+
 set.seed(420)
+
+#
+knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+#tinytex::install_tinytex()
+
 ```
@@
 
-*Insert text here*
+After removing the `ID` column, we were left with `31 variables`. 
+For this project our response variable is `diagnosis` as this variable tells whether an observation is Malignant (cancerous cells): `M` or Benign (noncancerous cells): `B`.
+The predictor variables make up the remaining 30 columns; however there are only actually 10 variables that each have 3 measurements.
+These measurements are `mean`, `standard error (SE)`, and `worst` (the mean of the 3 largest values of a variables).
+
+Below is a brief explanation of what each of the 10 variables represent:
+
+- Radius: Average Distance from cell center to cell perimeter
+- Texture: Standard deviation of gray-scale values; brightness of pixel of cell
+- Perimeter: Distance around nucleus boundary
+- Area: Area of the nucleus
+- Smoothness: Variation in cell's radial lengths
+- Compactness: The Perimeter^2/Area
+- Concavity: Size of the indention in nucleus boundary
+- Concave Points: Number of points on indented section of nucleus boundary
+- Symmetry: Deviation of the nuclei shape from the ideal measurement
+- Fractal Dimension: Measurement of irregularity in nucleus boundary
 
@@
 
-*Insert text here*
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+
+dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
+datatemp2 <- dataTemp %>% select_if(~ !any(is.na(.)))
+dataTot <- datatemp2[,!names(datatemp2) %in% c("id")]
+
+testtemp <- read.csv("test.csv")
+testTot <- testtemp[,!names(testtemp) %in% c("id")]
+
+traintemp <- read.csv("training.csv")
+trainTot <- traintemp[,!names(traintemp) %in% c("id")]
+```
+
+While Austin and Danny did Classification Analysis on `Type-II error rate`, I decided to do a Regression Analysis using the variable importance from their analysis's to compute the `training` and `test` error rate of those variables.
+This is to see how much error exists in a variable that is important in calculating if a cell is Malignant or Benign.
+However, I will be focused on `test` error in my results as its more important to see how our results will be on future data rather than what we have already trained.
+The regression models that I will be using are:
+
+    - Linear Regression
+    - Ridge Regression
+    - Lasso Regression
+
+## Variables
+
+In order to find the variables to run the Regression Analysis on, I had Austin and Danny run a variable importance plot on the data they collected and took the top 5 variables from each of their results, which are listed below.
+
+- Austin Top 5 Importance Variables:
+
+  1. *`area_mean`
+  2. *`area_worst`
+  3. *`concave.points_mean`
+  4. `texture_mean`
+  5. *`radius_worst`
+
+- Danny Top 5 Importance Variables
+
+  1. `concave.points_worst`
+  2. *`concave.points_mean`
+  3. *`area_worst`
+  4. *`area_mean`
+  5. *`radius_worst`
+
+After I got both Austin's and Danny's top 5 importance variables, I noticed that there were 4 common variables in each of their results (designated with a '*' in the results above).
+I then decided to see how the 4 common importance variables correlate to one another which can be seen in the pairwise scatterplot below:
+
+```{r correlation_plot, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+corr_df <- dataTot[,c( 'area_worst', 'area_mean', 'concave.points_mean', 'radius_worst')]
+ggpairs(corr_df) + labs(title = "Figure 1: Pairwise Scatterplot of Common Importance Variables")
+```
+
+What I found interesting was that `area_worst` & `raduis_worst` were more correlated that `area_worst` & `area_mean`.
+What I initially thought before making `Figure 1` was that `area_worst` & `area_mean` would be more correlated than the other 2 variables as they are both under the same variable category.
+
+## area_mean
+
+```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+#Train
+am_lm <- lm(area_mean ~ ., data = trainTot)
+pred_amtrain<- predict(am_lm, trainTot)
+amtrain_log <- postResample(pred_amtrain, trainTot$area_mean)
+#amtrain_log
+```
+```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+#Test
+pred_amtest <- predict(am_lm, testTot)
+amtest_log <- postResample(pred_amtest, testTot$area_mean)
+#amtest_log
+#summary(am_lm)
+```
+
+```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+dummy <- dummyVars(area_mean~., data = dataTot[,-1])
+x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
+y_train = trainTot$area_mean
+
+x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
+y_test = testTot$area_mean
+
+lambdas <- 10^seq(2, -3, by = -.1)
+ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
+cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
+optimal_lambda <- cv_ridge$lambda.min
+
+eval_results <- function(true, predicted, df) {
+  SSE <- sum((predicted - true)^2)
+  SST <- sum((true - mean(true))^2)
+  R_square <- 1 - SSE / SST
+  RMSE = sqrt(SSE/nrow(df))
+
+  # Model performance metrics
+data.frame(
+  RMSE = RMSE,
+  Rsquared = R_square)
+
+}
+
+# Prediction and evaluation on train data
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
+amr_train <- eval_results(y_train, predictions_train, trainTot)
+
+# Prediction and evaluation on test data
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+amr_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(amr_train)
+#print(amr_test)
+```
+
+```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+lambdas <- 10^seq(2, -3, by = -.1)
+lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+
+lambda_best <- lasso_reg$lambda.min 
+lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
+aml_train <- eval_results(y_train, predictions_train, trainTot)
+
+predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
+aml_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(aml_train)
+#print(aml_test)
+```
+
+For `area_mean` after running the 3 regression models, I found the lowest `test error` is from the ridge regression model with an `RMSE` of `23.70795` with a `Rsquared` of `99.489%`. 
+This mean that using ridge regression, we can explain `99.489%` of the variance in `area_mean` in Austin's and Danny's models.
+The ridge regression corresponding `training error` is `17.87197` with a `Rsquared` of 	`99.759%`.
+This corresponds to that `99.759%` of our variance in `area_mean` was trained in Austin's and Danny's models.
+
+```{r area_mean_table, echo=FALSE, message=FALSE, warning=FALSE}
+table_am <- as_tibble(rbind(amtrain_log,
+      amtest_log,
+      amr_train,
+      amr_test,
+      aml_train,
+      aml_test)) %>%
+    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
+    select(model, RMSE, Rsquared)
+table_am %>% 
+  kbl(digits = 5 ) %>% 
+  kable_styling(bootstrap_options = "striped") %>%
+  kable_styling(latex_options = "HOLD_position")
+```
+
+## Area_worst
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+#Test
+aw_lm <- lm(area_worst ~ ., data = trainTot)
+pred_awtest <- predict(aw_lm, testTot)
+awtest_log <- postResample(pred_awtest, testTot$area_worst)
+#awtest_log
+#summary(am_lm)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+#Train
+pred_awtrain<- predict(aw_lm, trainTot)
+awtrain_log <- postResample(pred_awtrain, trainTot$area_worst)
+#awtrain_log
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+dummy <- dummyVars(area_worst~., data = dataTot[,-1])
+x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
+y_train = trainTot$area_worst
+
+x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
+y_test = testTot$area_worst
+
+lambdas <- 10^seq(2, -3, by = -.1)
+ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
+cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
+optimal_lambda <- cv_ridge$lambda.min
+
+eval_results <- function(true, predicted, df) {
+  SSE <- sum((predicted - true)^2)
+  SST <- sum((true - mean(true))^2)
+  R_square <- 1 - SSE / SST
+  RMSE = sqrt(SSE/nrow(df))
+
+  # Model performance metrics
+data.frame(
+  RMSE = RMSE,
+  Rsquared = R_square)
+
+}
+
+# Prediction and evaluation on train data
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
+awr_train <- eval_results(y_train, predictions_train, trainTot)
+
+# Prediction and evaluation on test data
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+awr_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(awr_train)
+#print(awr_test)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+lambdas <- 10^seq(2, -3, by = -.1)
+lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+
+lambda_best <- lasso_reg$lambda.min 
+lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
+awl_train <- eval_results(y_train, predictions_train, trainTot)
+
+predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
+awl_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(awl_train)
+#print(awl_test)
+```
+
+In `area_worst` after running the 3 regression models, I found the lowest `test error` is from the ridge regression model with an `RMSE` of `42.28132` with a `Rsquared` of `99.288%`. 
+This mean that using ridge regression, we can explain `99.288%` of the variance in `area_worst` in Austin's and Danny's models.
+The ridge regression corresponding `training error` is `27.36780` with a `Rsquared` of `99.799%`.
+This corresponds to that `99.799%` of our variance in `area_worst` was trained in Austin's and Danny's models.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+table_aw <- as_tibble(rbind(awtrain_log,
+      awtest_log,
+      awr_train,
+      awr_test,
+      awl_train,
+      awl_test)) %>%
+    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
+    select(model, RMSE, Rsquared)
+table_aw %>% 
+  kbl(digits = 5 ) %>% 
+  kable_styling(bootstrap_options = "striped") %>%
+  kable_styling(latex_options = "HOLD_position")
+```
+
+## concave.points_mean
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+#Train
+cvptsm_lm <- lm(concave.points_mean ~ ., data = trainTot)
+pred_cvptsm_train<- predict(cvptsm_lm, trainTot)
+cvptsm_train_log <- postResample(pred_cvptsm_train, trainTot$concave.points_mean)
+#cvptsm_train_log
+```
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+#Test
+pred_cvptsm_test <- predict(cvptsm_lm, testTot)
+cvptsm_test_log <- postResample(pred_cvptsm_test, testTot$concave.points_mean)
+#cvptsm_test_log
+#summary(am_lm)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+dummy <- dummyVars(concave.points_mean~., data = dataTot[,-1])
+x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
+y_train = trainTot$concave.points_mean
+
+x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
+y_test = testTot$concave.points_mean
+
+lambdas <- 10^seq(2, -3, by = -.1)
+ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
+cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
+optimal_lambda <- cv_ridge$lambda.min
+
+eval_results <- function(true, predicted, df) {
+  SSE <- sum((predicted - true)^2)
+  SST <- sum((true - mean(true))^2)
+  R_square <- 1 - SSE / SST
+  RMSE = sqrt(SSE/nrow(df))
+
+  # Model performance metrics
+data.frame(
+  RMSE = RMSE,
+  Rsquared = R_square)
+
+}
+
+# Prediction and evaluation on train data
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
+cvptsmr_train <- eval_results(y_train, predictions_train, trainTot)
+
+# Prediction and evaluation on test data
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+cvptsmr_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(cvptsmr_train)
+#print(cvptsmr_test)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+lambdas <- 10^seq(2, -3, by = -.1)
+lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+
+lambda_best <- lasso_reg$lambda.min 
+lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
+cvptsml_train <- eval_results(y_train, predictions_train, trainTot)
+
+predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
+cvptsml_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(cvptsml_train)
+#print(cvptsml_test)
+```
+
+With `concave.points_mean` after running the 3 regression models, I found the lowest `test error` is from the linear regression model with an `RMSE` of `0.00574` with a `Rsquared` of `97.847%`. 
+This mean that using ridge regression, we can explain `97.847%` of the variance in `concave.points_mean` in Austin's and Danny's models.
+The ridge regression corresponding `training error` is `0.00467` with a `Rsquared` of `98.532%`.
+This corresponds to that `98.532%` of our variance in `concave.points_mean` was trained in Austin's and Danny's models.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+table_am <- as_tibble(rbind(cvptsm_train_log,
+      cvptsm_test_log,
+      cvptsmr_train,
+      cvptsmr_test,
+      cvptsml_train,
+      cvptsml_test)) %>%
+    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
+    select(model, RMSE, Rsquared)
+table_am %>% 
+  kbl(digits = 5 ) %>% 
+  kable_styling(bootstrap_options = "striped") %>%
+  kable_styling(latex_options = "HOLD_position")
+```
+
+## radius_worst
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+#Train
+rw_lm <- lm(radius_worst ~ ., data = trainTot)
+pred_rw_train<- predict(rw_lm, trainTot)
+rw_train_log <- postResample(pred_rw_train, trainTot$radius_worst)
+#rw_train_log
+```
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+#Test
+pred_rw_test <- predict(rw_lm, testTot)
+rw_test_log <- postResample(pred_rw_test, testTot$radius_worst)
+#rw_test_log
+#summary(am_lm)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+dummy <- dummyVars(radius_worst~., data = dataTot[,-1])
+x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
+y_train = trainTot$radius_worst
+
+x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
+y_test = testTot$radius_worst
+
+lambdas <- 10^seq(2, -3, by = -.1)
+ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
+cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
+optimal_lambda <- cv_ridge$lambda.min
+
+eval_results <- function(true, predicted, df) {
+  SSE <- sum((predicted - true)^2)
+  SST <- sum((true - mean(true))^2)
+  R_square <- 1 - SSE / SST
+  RMSE = sqrt(SSE/nrow(df))
+
+  # Model performance metrics
+data.frame(
+  RMSE = RMSE,
+  Rsquared = R_square)
+
+}
+
+# Prediction and evaluation on train data
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
+rwr_train <- eval_results(y_train, predictions_train, trainTot)
+
+# Prediction and evaluation on test data
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+rwr_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(rwr_train)
+#print(rwr_test)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+lambdas <- 10^seq(2, -3, by = -.1)
+lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+
+lambda_best <- lasso_reg$lambda.min 
+lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
+rwl_train <- eval_results(y_train, predictions_train, trainTot)
+
+predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
+rwl_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(rwl_train)
+#print(rwl_test)
+```
+
+For `radius_worst` after running the 3 regression models, I found the lowest `test error` is from the lasso regression model with an `RMSE` of `0.24120	` with a `Rsquared` of `99.713%`. 
+This mean that using ridge regression, we can explain `99.713%` of the variance in `radius_worst` in Austin's and Danny's models.
+The ridge regression corresponding `training error` is `0.15907` with a `Rsquared` of `99.900%`.
+This corresponds to that `99.900%` of our variance in `radius_worst` was trained in Austin's and Danny's models.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+table_am <- as_tibble(rbind(rw_train_log,
+      rw_test_log,
+      rwr_train,
+      rwr_test,
+      rwl_train,
+      rwl_test)) %>%
+    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
+    select(model, RMSE, Rsquared)
+table_am %>% 
+  kbl(digits = 5 ) %>% 
+  kable_styling(bootstrap_options = "striped") %>%
+  kable_styling(latex_options = "HOLD_position")
+```
 

### 2022-12-15 dannys_temp_paper.rmd

@@
 
-*Insert text here*
+```{r, echo = F, warning = F, message = F}
+library(tidyverse)
+library(GGally)
+library(ggplot2)
+library(tidyr)
+library(tidymodels)
+library(vip)
+library(caTools)
+library(xgboost)
+library(discrim)
+library(glmnet)
+library(kableExtra)
+
+### Gathering Data
+
+train <- read.csv("training.csv")
+test <- read.csv("test.csv")
+
+train |>
+  select(-id) -> train
+
+test |>
+  select(-id) -> test
+
+train$diagnosis <- as.factor(train$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+```
+
+As the main purpose of this analysis is to reduce Type II error, we wanted to try different ways of recuding this within similar models.
+The following models all use classification to classify a tumor as benign or malignant however they have a lower prediction cutoff of 25% malignant to bias our results towards more malignant.
+For example, if a tumor is only 30% likely to be malignant, we still would classify that tumor as malignant to avoid false negatives.
+While this does lead to higher Type 1 error and sometimes lower overall accuracy, we aren't as worried about the false-positives.
+We fit a number of models on the training data using this approach, including Logistic Regression, K-Nearest Neighbors, Quadratic Discriminant Analysis (QDA), Random Forest, xgBoost, and lastly a support vector classifier.
+While all the most models performed quite decently, there were a few that stood out.
+
+### K-Nearest Neighbors
+
+```{r, echo = F, warning = F, message = F}
+set.seed(445)
+### KNN
+
+cancer_10foldcv <- vfold_cv(train, v = 10)
+neighbors_df <- data.frame(neighbors = c(1,3,5,7,10))
+
+knn_spec <- the_model <- nearest_neighbor(neighbors = tune("neighbors")) %>%
+  set_engine("kknn") %>% 
+  set_mode("classification")
+
+knn_rec <- recipe(diagnosis ~ ., data = train)
+
+workflow() |>
+  add_model(knn_spec) |>
+  add_recipe(knn_rec) -> knn_wf
+
+knn_wf |>
+  tune_grid(resamples = cancer_10foldcv, grid = neighbors_df) -> knn_tune
+
+knn_final <- finalize_workflow(knn_wf, select_best(knn_tune, metric = "accuracy"))
+knn_final_fit <- fit(knn_final, data = train)
+
+knn_final_fit |>
+  augment(new_data = test) -> knn_test_res
+
+knn_test_res |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  conf_mat(truth = diagnosis, estimate = pred_lower_cutoff) -> knn.cm
+
+knn_test_res |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  accuracy(truth = diagnosis, estimate = pred_lower_cutoff) |>
+  mutate(error = 1 - .estimate) |>
+  pull(error) -> knn.err
+
+100 - (round(knn.err, 4) * 100) -> knn.err
+
+neighbors_best <- select_best(knn_tune, metric = "accuracy") |> pull(neighbors)
+```
+
+K-Nearest Neighbors is a classifying method that estimates the bayes classifier using the closest training points to the test point that is being classified.
+It uses a neighbor parameter that is a big factor in determining the classifier.
+For this model, we used 10-fold cross-validation to tune the neighbor parameter based on the highest accuracy.
+Once the neighbors parameter was tuned to `r neighbors_best`, the model was fit and prediction ensued.
+The confusion matrix showcasing the classification prediction of this model is below.
+
+```{r, echo = F, warning = F, message = F}
+ggplot(data.frame(knn.cm$table), aes(x = Prediction, y = Truth, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "K-Nearest Neighbors Confusion Matrix", x = "Prediction", y = "Truth") +
+  theme_minimal() +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798") +
+  guides(fill = F)
+```
+
+As shown in the confusion matrix, this KNN model only predicted two false-negatives out of 223 test observations and 90 true test malignant tumors.
+The KNN model predicted with an overall accuracy of `r knn.err`.
+While this model did perform quite well, we still want to improve these results.
+
+### Support Vector Classifier
+
+```{r, echo = F, warning = F, message = F}
+set.seed(445)
+### SVC
+
+df_cost <- grid_regular(cost(), levels = 10)
+
+svm_linear_tune_spec <- svm_poly(degree = 1, cost = tune("cost")) %>%
+  set_mode("classification") %>%
+  set_engine("kernlab", scaled = FALSE)
+
+svm_linear_rec <- recipe(diagnosis ~ ., data = train)
+
+svm_linear_wf <- workflow() |>
+  add_model(svm_linear_tune_spec) |>
+  add_recipe(svm_linear_rec)
+
+tune_fit <- svm_linear_wf |>
+  tune_grid(resamples = cancer_10foldcv, grid = df_cost)
+
+svm_linear_final <- finalize_workflow(svm_linear_wf, select_best(tune_fit, metric = "accuracy"))
+
+svm_linear_final |>
+  fit(data = train) -> svm_linear_final_fit
+
+svm_linear_final_fit |>
+  extract_fit_engine() -> svm_linear_final_fit_engine
+
+svm_linear_final_fit |>
+  augment(new_data = test) |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  conf_mat(truth = diagnosis, estimate = pred_lower_cutoff) -> svm.cm
+
+svm_linear_final_fit |>
+  augment(new_data = test) |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  accuracy(truth = diagnosis, estimate = pred_lower_cutoff) |>
+  mutate(error = 1 - .estimate) |>
+  pull(error) -> svm.err
+
+100 - (round(svm.err, 4) * 100) -> svm.err
+
+cost_best <- select_best(tune_fit, metric = "accuracy") |> pull(cost)
+```
+
+A support vector classifier uses a softened margin classifier to classify the observations but with less sensitivity than a maximal margin classifier.
+The support vector classifier will not necessarily classify all training points accurately, but this allows for better test prediction.
+We tuned our support vector classifier to take a cost value of `r cost_best` using 10-fold cross-validation.
+The confusion matrix showcasing the prediciton accuracy of the support vector classifier model can be seen below.
+
+```{r, echo = F, warning = F, message = F}
+ggplot(data.frame(svm.cm$table), aes(x = Prediction, y = Truth, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Support Vector Classifier Confusion Matrix", x = "Prediction", y = "Truth") +
+  theme_minimal() + 
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798") +  
+  guides(fill = F)
+```
+
+As shown above, the support vector classifier predicted the exact same as the KNN model.
+With only 2 false negatives and an overall accuracy of `r svm.err`, this model was not an improvement from the KNN model, but it is reassuring to see consistent high accuracy amongst different models.
+The last model we will look at is a Random Forest model.
+
+### Random Forest
+
+```{r, echo = F, warning = F, message = F}
+### Random Forest
+set.seed(420)
+rf_spec <- rand_forest(mtry = sqrt(.cols())) |>
+  set_engine("randomForest", importance = TRUE) |>
+  set_mode("classification")
+
+rf_fit <- rf_spec |>
+  fit(diagnosis ~ ., data = train)
+
+rf_fit |>
+  augment(new_data = test) |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  conf_mat(truth = diagnosis, estimate = pred_lower_cutoff) -> rf.cm
+
+rf_fit |>
+  augment(new_data = test) |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  accuracy(truth = diagnosis, estimate = pred_lower_cutoff) |>
+  mutate(error = 1 - .estimate) |>
+  pull(error) -> rf.err
+
+100 - (round(rf.err, 4) * 100) -> rf.err
+```
+
+Random Forest creates many decision trees from bootstrapped sample training points and uses a subset of predictors for each tree.
+The main difference between Random Forest and bagging decision trees is this feature subset that helps decorrelate the trees.
+For this Random Forest model, a default of 500 trees was used as tuning did not improve the model.
+The confusion matrix showcasing the prediction accuracy of the Random Forest model is shown below.
+
+```{r, echo = F, echo = F, warning = F, message = F}
+ggplot(data.frame(rf.cm$table), aes(x = Prediction, y = Truth, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Random Forest Confusion Matrix", x = "Prediction", y = "Truth") +
+  theme_minimal() +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798") +  
+  guides(fill = F)
+```
+
+As shown above, the Random Forest model predicted the best out of all our models according to the standards we set.
+With only one false-negative, this is the lowest out of all the models and with an overall accuracy of `r rf.err` it holds up against the other models and does not gain too much Type I error.
+This confusion matrix portrays the predicition cutoff very well.
+Because the model biases towards classifying as malignant, there is much higher Type I error than Type II error, which is exactly what we were aiming for.
 

### 2022-12-15 paper.rmd

@@
 ```
+
+\newpage
 
@@
 
-*Insert text here*
+\newpage
 
@@
 
-*Insert text here*
+```{r, echo = F, warning = F, message = F}
+library(tidyverse)
+library(GGally)
+library(ggplot2)
+library(tidyr)
+library(tidymodels)
+library(vip)
+library(caTools)
+library(xgboost)
+library(discrim)
+library(glmnet)
+library(kableExtra)
+
+### Gathering Data
+
+train <- read.csv("training.csv")
+test <- read.csv("test.csv")
+
+train |>
+  select(-id) -> train
+
+test |>
+  select(-id) -> test
+
+train$diagnosis <- as.factor(train$diagnosis)
+test$diagnosis <- as.factor(test$diagnosis)
+```
+
+As the main purpose of this analysis is to reduce Type II error, we wanted to try different ways of recuding this within similar models.
+The following models all use classification to classify a tumor as benign or malignant however they have a lower prediction cutoff of 25% malignant to bias our results towards more malignant.
+For example, if a tumor is only 30% likely to be malignant, we still would classify that tumor as malignant to avoid false negatives.
+While this does lead to higher Type 1 error and sometimes lower overall accuracy, we aren't as worried about the false-positives.
+We fit a number of models on the training data using this approach, including Logistic Regression, K-Nearest Neighbors, Quadratic Discriminant Analysis (QDA), Random Forest, xgBoost, and lastly a support vector classifier.
+While all the most models performed quite decently, there were a few that stood out.
+
+### K-Nearest Neighbors
+
+```{r, echo = F, warning = F, message = F}
+set.seed(445)
+### KNN
+
+cancer_10foldcv <- vfold_cv(train, v = 10)
+neighbors_df <- data.frame(neighbors = c(1,3,5,7,10))
+
+knn_spec <- the_model <- nearest_neighbor(neighbors = tune("neighbors")) %>%
+  set_engine("kknn") %>% 
+  set_mode("classification")
+
+knn_rec <- recipe(diagnosis ~ ., data = train)
+
+workflow() |>
+  add_model(knn_spec) |>
+  add_recipe(knn_rec) -> knn_wf
+
+knn_wf |>
+  tune_grid(resamples = cancer_10foldcv, grid = neighbors_df) -> knn_tune
+
+knn_final <- finalize_workflow(knn_wf, select_best(knn_tune, metric = "accuracy"))
+knn_final_fit <- fit(knn_final, data = train)
+
+knn_final_fit |>
+  augment(new_data = test) -> knn_test_res
+
+knn_test_res |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  conf_mat(truth = diagnosis, estimate = pred_lower_cutoff) -> knn.cm
+
+knn_test_res |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  accuracy(truth = diagnosis, estimate = pred_lower_cutoff) |>
+  mutate(error = 1 - .estimate) |>
+  pull(error) -> knn.err
+
+100 - (round(knn.err, 4) * 100) -> knn.err
+
+neighbors_best <- select_best(knn_tune, metric = "accuracy") |> pull(neighbors)
+```
+
+K-Nearest Neighbors is a classifying method that estimates the bayes classifier using the closest training points to the test point that is being classified.
+It uses a neighbor parameter that is a big factor in determining the classifier.
+For this model, we used 10-fold cross-validation to tune the neighbor parameter based on the highest accuracy.
+Once the neighbors parameter was tuned to `r neighbors_best`, the model was fit and prediction ensued.
+The confusion matrix showcasing the classification prediction of this model is below.
+
+```{r, echo = F, warning = F, message = F}
+ggplot(data.frame(knn.cm$table), aes(x = Prediction, y = Truth, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "K-Nearest Neighbors Confusion Matrix", x = "Prediction", y = "Truth") +
+  theme_minimal() +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798") +
+  guides(fill = F)
+```
+
+As shown in the confusion matrix, this KNN model only predicted two false-negatives out of 223 test observations and 90 true test malignant tumors.
+The KNN model predicted with an overall accuracy of `r knn.err`.
+While this model did perform quite well, we still want to improve these results.
+
+### Support Vector Classifier
+
+```{r, echo = F, warning = F, message = F}
+set.seed(445)
+### SVC
+
+df_cost <- grid_regular(cost(), levels = 10)
+
+svm_linear_tune_spec <- svm_poly(degree = 1, cost = tune("cost")) %>%
+  set_mode("classification") %>%
+  set_engine("kernlab", scaled = FALSE)
+
+svm_linear_rec <- recipe(diagnosis ~ ., data = train)
+
+svm_linear_wf <- workflow() |>
+  add_model(svm_linear_tune_spec) |>
+  add_recipe(svm_linear_rec)
+
+tune_fit <- svm_linear_wf |>
+  tune_grid(resamples = cancer_10foldcv, grid = df_cost)
+
+svm_linear_final <- finalize_workflow(svm_linear_wf, select_best(tune_fit, metric = "accuracy"))
+
+svm_linear_final |>
+  fit(data = train) -> svm_linear_final_fit
+
+svm_linear_final_fit |>
+  extract_fit_engine() -> svm_linear_final_fit_engine
+
+svm_linear_final_fit |>
+  augment(new_data = test) |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  conf_mat(truth = diagnosis, estimate = pred_lower_cutoff) -> svm.cm
+
+svm_linear_final_fit |>
+  augment(new_data = test) |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  accuracy(truth = diagnosis, estimate = pred_lower_cutoff) |>
+  mutate(error = 1 - .estimate) |>
+  pull(error) -> svm.err
+
+100 - (round(svm.err, 4) * 100) -> svm.err
+
+cost_best <- select_best(tune_fit, metric = "accuracy") |> pull(cost)
+```
+
+A support vector classifier uses a softened margin classifier to classify the observations but with less sensitivity than a maximal margin classifier.
+The support vector classifier will not necessarily classify all training points accurately, but this allows for better test prediction.
+We tuned our support vector classifier to take a cost value of `r cost_best` using 10-fold cross-validation.
+The confusion matrix showcasing the prediciton accuracy of the support vector classifier model can be seen below.
+
+```{r, echo = F, warning = F, message = F}
+ggplot(data.frame(svm.cm$table), aes(x = Prediction, y = Truth, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Support Vector Classifier Confusion Matrix", x = "Prediction", y = "Truth") +
+  theme_minimal() + 
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798") +  
+  guides(fill = F)
+```
+
+As shown above, the support vector classifier predicted the exact same as the KNN model.
+With only 2 false negatives and an overall accuracy of `r svm.err`, this model was not an improvement from the KNN model, but it is reassuring to see consistent high accuracy amongst different models.
+The last model we will look at is a Random Forest model.
+
+### Random Forest
+
+```{r, echo = F, warning = F, message = F}
+### Random Forest
+set.seed(420)
+rf_spec <- rand_forest(mtry = sqrt(.cols())) |>
+  set_engine("randomForest", importance = TRUE) |>
+  set_mode("classification")
+
+rf_fit <- rf_spec |>
+  fit(diagnosis ~ ., data = train)
+
+rf_fit |>
+  augment(new_data = test) |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  conf_mat(truth = diagnosis, estimate = pred_lower_cutoff) -> rf.cm
+
+rf_fit |>
+  augment(new_data = test) |>
+  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
+  accuracy(truth = diagnosis, estimate = pred_lower_cutoff) |>
+  mutate(error = 1 - .estimate) |>
+  pull(error) -> rf.err
+
+100 - (round(rf.err, 4) * 100) -> rf.err
+```
+
+Random Forest creates many decision trees from bootstrapped sample training points and uses a subset of predictors for each tree.
+The main difference between Random Forest and bagging decision trees is this feature subset that helps decorrelate the trees.
+For this Random Forest model, a default of 500 trees was used as tuning did not improve the model.
+The confusion matrix showcasing the prediction accuracy of the Random Forest model is shown below.
+
+```{r, echo = F, echo = F, warning = F, message = F}
+ggplot(data.frame(rf.cm$table), aes(x = Prediction, y = Truth, fill = Freq)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), size = 12) +
+  scale_fill_gradient(low = "white", high = "#c82798") +
+  labs(title = "Random Forest Confusion Matrix", x = "Prediction", y = "Truth") +
+  theme_minimal() +
+  annotate("path",
+   x=1+.25*cos(seq(0,2*pi,length.out=100)),
+   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798") +  
+  guides(fill = F)
+```
+
+As shown above, the Random Forest model predicted the best out of all our models according to the standards we set.
+With only one false-negative, this is the lowest out of all the models and with an overall accuracy of `r rf.err` it holds up against the other models and does not gain too much Type I error.
+This confusion matrix portrays the predicition cutoff very well.
+Because the model biases towards classifying as malignant, there is much higher Type I error than Type II error, which is exactly what we were aiming for.
+
+\newpage
 
@@
 *Insert text here*
+
+\newpage
 
@@
 
+\newpage
+
 # Works Cited

### 2022-12-15 ethans_temp_paper.rmd
Update ethans_temp_paper.rmd

@@
 ```{r setup, include=FALSE}
-# Import packages
-library(tidyverse)
-library(ggplot2)
-library(GGally)
-library(knitr)
-library(boot)
-library(caret)
-library(randomForest)
-library(e1071)
-library(class)
 #
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+
+set.seed(420)
 

### 2022-12-15 paper.rmd

@@
 
-*Insert text here regarding the GGPairs Plot*
+Another important thing to note about the data is the variables included.
+There are three different measurements for the ten predictor variables: mean value, standard error, and "worst" value (mean of 3 largest measurements).
+As shown the correlation plot below, these three measurements are all correlated to each other, but still hold a strong relationship with the diagnosis response variable.
+The relation between the mean value and "worst" value is almost completely correlated which makes sense as both are averages however, the "worst" value only uses the three largest measurements.
+This is important to know when trying to interpret feature effect, importance and selection.
 
@@
 
-knn_spec <- the_model <- nearest_neighbor(neighbors = tune("neighbors")) %>%
+knn_spec <- nearest_neighbor(neighbors = tune("neighbors")) %>%
   set_engine("kknn") %>% 
@@
 
-*Insert text here*
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+# Import packages
+library(tidyverse)
+library(ggplot2)
+library(GGally)
+library(knitr)
+library(boot)
+library(caret)
+library(randomForest)
+library(e1071)
+library(class)
+#
+library(dplyr)
+library(leaps)
+library(caTools)
+library(tidymodels)
+library(glmnet)
+library(forecast)
+library(plyr)
+library(readr)
+library(repr)
+library(kableExtra)
+
+dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
+datatemp2 <- dataTemp %>% select_if(~ !any(is.na(.)))
+dataTot <- datatemp2[,!names(datatemp2) %in% c("id")]
+
+testtemp <- read.csv("test.csv")
+testTot <- testtemp[,!names(testtemp) %in% c("id")]
+
+traintemp <- read.csv("training.csv")
+trainTot <- traintemp[,!names(traintemp) %in% c("id")]
+```
+
+While Austin and Danny did Classification Analysis on `Type-II error rate`, I decided to do a Regression Analysis using the variable importance from their analysis's to compute the `training` and `test` error rate of those variables.
+This is to see how much error exists in a variable that is important in calculating if a cell is Malignant or Benign.
+However, I will be focused on `test` error in my results as its more important to see how our results will be on future data rather than what we have already trained.
+The regression models that I will be using are:
+
+    - Linear Regression
+    - Ridge Regression
+    - Lasso Regression
+
+## Variables
+
+In order to find the variables to run the Regression Analysis on, I had Austin and Danny run a variable importance plot on the data they collected and took the top 5 variables from each of their results, which are listed below.
+
+- Austin Top 5 Importance Variables:
+
+  1. *`area_mean`
+  2. *`area_worst`
+  3. *`concave.points_mean`
+  4. `texture_mean`
+  5. *`radius_worst`
+
+- Danny Top 5 Importance Variables
+
+  1. `concave.points_worst`
+  2. *`concave.points_mean`
+  3. *`area_worst`
+  4. *`area_mean`
+  5. *`radius_worst`
+
+After I got both Austin's and Danny's top 5 importance variables, I noticed that there were 4 common variables in each of their results (designated with a '*' in the results above).
+I then decided to see how the 4 common importance variables correlate to one another which can be seen in the pairwise scatterplot below:
+
+```{r correlation_plot, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+corr_df <- dataTot[,c( 'area_worst', 'area_mean', 'concave.points_mean', 'radius_worst')]
+ggpairs(corr_df) + labs(title = "Figure 1: Pairwise Scatterplot of Common Importance Variables")
+```
+
+What I found interesting was that `area_worst` & `raduis_worst` were more correlated that `area_worst` & `area_mean`.
+What I initially thought before making `Figure 1` was that `area_worst` & `area_mean` would be more correlated than the other 2 variables as they are both under the same variable category.
+
+## area_mean
+
+```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+#Train
+am_lm <- lm(area_mean ~ ., data = trainTot)
+pred_amtrain<- predict(am_lm, trainTot)
+amtrain_log <- postResample(pred_amtrain, trainTot$area_mean)
+#amtrain_log
+```
+```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+#Test
+pred_amtest <- predict(am_lm, testTot)
+amtest_log <- postResample(pred_amtest, testTot$area_mean)
+#amtest_log
+#summary(am_lm)
+```
+
+```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+dummy <- dummyVars(area_mean~., data = dataTot[,-1])
+x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
+y_train = trainTot$area_mean
+
+x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
+y_test = testTot$area_mean
+
+lambdas <- 10^seq(2, -3, by = -.1)
+ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
+cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
+optimal_lambda <- cv_ridge$lambda.min
+
+eval_results <- function(true, predicted, df) {
+  SSE <- sum((predicted - true)^2)
+  SST <- sum((true - mean(true))^2)
+  R_square <- 1 - SSE / SST
+  RMSE = sqrt(SSE/nrow(df))
+
+  # Model performance metrics
+data.frame(
+  RMSE = RMSE,
+  Rsquared = R_square)
+
+}
+
+# Prediction and evaluation on train data
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
+amr_train <- eval_results(y_train, predictions_train, trainTot)
+
+# Prediction and evaluation on test data
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+amr_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(amr_train)
+#print(amr_test)
+```
+
+```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+lambdas <- 10^seq(2, -3, by = -.1)
+lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+
+lambda_best <- lasso_reg$lambda.min 
+lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
+aml_train <- eval_results(y_train, predictions_train, trainTot)
+
+predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
+aml_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(aml_train)
+#print(aml_test)
+```
+
+For `area_mean` after running the 3 regression models, I found the lowest `test error` is from the ridge regression model with an `RMSE` of `23.70795` with a `Rsquared` of `99.489%`. 
+This mean that using ridge regression, we can explain `99.489%` of the variance in `area_mean` in Austin's and Danny's models.
+The ridge regression corresponding `training error` is `17.87197` with a `Rsquared` of 	`99.759%`.
+This corresponds to that `99.759%` of our variance in `area_mean` was trained in Austin's and Danny's models.
+
+```{r area_mean_table, echo=FALSE, message=FALSE, warning=FALSE}
+table_am <- as_tibble(rbind(amtrain_log,
+      amtest_log,
+      amr_train,
+      amr_test,
+      aml_train,
+      aml_test)) %>%
+    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
+    select(model, RMSE, Rsquared)
+table_am %>% 
+  kbl(digits = 5 ) %>% 
+  kable_styling(bootstrap_options = "striped") %>%
+  kable_styling(latex_options = "HOLD_position")
+```
+
+## Area_worst
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+#Test
+aw_lm <- lm(area_worst ~ ., data = trainTot)
+pred_awtest <- predict(aw_lm, testTot)
+awtest_log <- postResample(pred_awtest, testTot$area_worst)
+#awtest_log
+#summary(am_lm)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+#Train
+pred_awtrain<- predict(aw_lm, trainTot)
+awtrain_log <- postResample(pred_awtrain, trainTot$area_worst)
+#awtrain_log
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+dummy <- dummyVars(area_worst~., data = dataTot[,-1])
+x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
+y_train = trainTot$area_worst
+
+x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
+y_test = testTot$area_worst
+
+lambdas <- 10^seq(2, -3, by = -.1)
+ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
+cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
+optimal_lambda <- cv_ridge$lambda.min
+
+eval_results <- function(true, predicted, df) {
+  SSE <- sum((predicted - true)^2)
+  SST <- sum((true - mean(true))^2)
+  R_square <- 1 - SSE / SST
+  RMSE = sqrt(SSE/nrow(df))
+
+  # Model performance metrics
+data.frame(
+  RMSE = RMSE,
+  Rsquared = R_square)
+
+}
+
+# Prediction and evaluation on train data
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
+awr_train <- eval_results(y_train, predictions_train, trainTot)
+
+# Prediction and evaluation on test data
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+awr_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(awr_train)
+#print(awr_test)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+lambdas <- 10^seq(2, -3, by = -.1)
+lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+
+lambda_best <- lasso_reg$lambda.min 
+lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
+awl_train <- eval_results(y_train, predictions_train, trainTot)
+
+predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
+awl_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(awl_train)
+#print(awl_test)
+```
+
+In `area_worst` after running the 3 regression models, I found the lowest `test error` is from the ridge regression model with an `RMSE` of `42.28132` with a `Rsquared` of `99.288%`. 
+This mean that using ridge regression, we can explain `99.288%` of the variance in `area_worst` in Austin's and Danny's models.
+The ridge regression corresponding `training error` is `27.36780` with a `Rsquared` of `99.799%`.
+This corresponds to that `99.799%` of our variance in `area_worst` was trained in Austin's and Danny's models.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = "H", out.extra = ""}
+table_aw <- as_tibble(rbind(awtrain_log,
+      awtest_log,
+      awr_train,
+      awr_test,
+      awl_train,
+      awl_test)) %>%
+    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
+    select(model, RMSE, Rsquared)
+table_aw %>% 
+  kbl(digits = 5 ) %>% 
+  kable_styling(bootstrap_options = "striped") %>%
+  kable_styling(latex_options = "HOLD_position")
+```
+
+## concave.points_mean
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+#Train
+cvptsm_lm <- lm(concave.points_mean ~ ., data = trainTot)
+pred_cvptsm_train<- predict(cvptsm_lm, trainTot)
+cvptsm_train_log <- postResample(pred_cvptsm_train, trainTot$concave.points_mean)
+#cvptsm_train_log
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+#Test
+pred_cvptsm_test <- predict(cvptsm_lm, testTot)
+cvptsm_test_log <- postResample(pred_cvptsm_test, testTot$concave.points_mean)
+#cvptsm_test_log
+#summary(am_lm)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE, fig.pos = "H", out.extra = ""}
+dummy <- dummyVars(concave.points_mean~., data = dataTot[,-1])
+x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
+y_train = trainTot$concave.points_mean
+
+x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
+y_test = testTot$concave.points_mean
+
+lambdas <- 10^seq(2, -3, by = -.1)
+ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
+cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
+optimal_lambda <- cv_ridge$lambda.min
+
+eval_results <- function(true, predicted, df) {
+  SSE <- sum((predicted - true)^2)
+  SST <- sum((true - mean(true))^2)
+  R_square <- 1 - SSE / SST
+  RMSE = sqrt(SSE/nrow(df))
+
+  # Model performance metrics
+data.frame(
+  RMSE = RMSE,
+  Rsquared = R_square)
+
+}
+
+# Prediction and evaluation on train data
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
+cvptsmr_train <- eval_results(y_train, predictions_train, trainTot)
+
+# Prediction and evaluation on test data
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+cvptsmr_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(cvptsmr_train)
+#print(cvptsmr_test)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+lambdas <- 10^seq(2, -3, by = -.1)
+lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+
+lambda_best <- lasso_reg$lambda.min 
+lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
+cvptsml_train <- eval_results(y_train, predictions_train, trainTot)
+
+predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
+cvptsml_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(cvptsml_train)
+#print(cvptsml_test)
+```
+
+With `concave.points_mean` after running the 3 regression models, I found the lowest `test error` is from the linear regression model with an `RMSE` of `0.00574` with a `Rsquared` of `97.847%`. 
+This mean that using ridge regression, we can explain `97.847%` of the variance in `concave.points_mean` in Austin's and Danny's models.
+The ridge regression corresponding `training error` is `0.00467` with a `Rsquared` of `98.532%`.
+This corresponds to that `98.532%` of our variance in `concave.points_mean` was trained in Austin's and Danny's models.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = "H", out.extra = ""}
+table_am <- as_tibble(rbind(cvptsm_train_log,
+      cvptsm_test_log,
+      cvptsmr_train,
+      cvptsmr_test,
+      cvptsml_train,
+      cvptsml_test)) %>%
+    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
+    select(model, RMSE, Rsquared)
+table_am %>% 
+  kbl(digits = 5 ) %>% 
+  kable_styling(bootstrap_options = "striped") %>%
+  kable_styling(latex_options = "HOLD_position")
+```
+
+## radius_worst
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+#Train
+rw_lm <- lm(radius_worst ~ ., data = trainTot)
+pred_rw_train<- predict(rw_lm, trainTot)
+rw_train_log <- postResample(pred_rw_train, trainTot$radius_worst)
+#rw_train_log
+```
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+#Test
+pred_rw_test <- predict(rw_lm, testTot)
+rw_test_log <- postResample(pred_rw_test, testTot$radius_worst)
+#rw_test_log
+#summary(am_lm)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE, fig.pos = "H", out.extra = ""}
+dummy <- dummyVars(radius_worst~., data = dataTot[,-1])
+x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
+y_train = trainTot$radius_worst
+
+x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
+y_test = testTot$radius_worst
+
+lambdas <- 10^seq(2, -3, by = -.1)
+ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
+cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
+optimal_lambda <- cv_ridge$lambda.min
+
+eval_results <- function(true, predicted, df) {
+  SSE <- sum((predicted - true)^2)
+  SST <- sum((true - mean(true))^2)
+  R_square <- 1 - SSE / SST
+  RMSE = sqrt(SSE/nrow(df))
+
+  # Model performance metrics
+data.frame(
+  RMSE = RMSE,
+  Rsquared = R_square)
+
+}
+
+# Prediction and evaluation on train data
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
+rwr_train <- eval_results(y_train, predictions_train, trainTot)
+
+# Prediction and evaluation on test data
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+rwr_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(rwr_train)
+#print(rwr_test)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
+lambdas <- 10^seq(2, -3, by = -.1)
+lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+
+lambda_best <- lasso_reg$lambda.min 
+lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
+rwl_train <- eval_results(y_train, predictions_train, trainTot)
+
+predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
+rwl_test <- eval_results(y_test, predictions_test, testTot)
+
+#print(rwl_train)
+#print(rwl_test)
+```
+
+For `radius_worst` after running the 3 regression models, I found the lowest `test error` is from the lasso regression model with an `RMSE` of `0.24120	` with a `Rsquared` of `99.713%`. 
+This mean that using ridge regression, we can explain `99.713%` of the variance in `radius_worst` in Austin's and Danny's models.
+The ridge regression corresponding `training error` is `0.15907` with a `Rsquared` of `99.900%`.
+This corresponds to that `99.900%` of our variance in `radius_worst` was trained in Austin's and Danny's models.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = "H", out.extra = ""}
+table_am <- as_tibble(rbind(rw_train_log,
+      rw_test_log,
+      rwr_train,
+      rwr_test,
+      rwl_train,
+      rwl_test)) %>%
+    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
+    select(model, RMSE, Rsquared)
+table_am %>% 
+  kbl(digits = 5 ) %>% 
+  kable_styling(bootstrap_options = "striped") %>%
+  kable_styling(latex_options = "HOLD_position")
+```
 
@@
 
-*Insert text here*
+From the results of our classification analysis, we are confident not only in our predictions of tumor diagnosis, but reducing the amount of false-negative diagnoses.
+Through our methods of tuning specifically for Type II error, and lowering the prediction cutoff for malignant tumors, we have shown two successful ways of battling classification data with a preffered error class.
+This is extremely important in cancer diagnosis, as well as other analyses, as we do not want to tell someone they don't have cancer when they do, as that could delay treatment and further risk their life.
+
+In terms of feature importance and feature selection, there were many limitations in interpretations due to the correlation of the measurement types.
+We did find that there were a number of the same features outlined by different models, so we can conclude that the variables, regardless of the measurement type, hold importance in classifying a tumor diagnosis.
 

### 2022-12-15 dannys_temp_paper.rmd

@@
 *Insert text here*
+
+Another important thing to note about the data is the variables included.
+There are three different measurements for the ten predictor variables: mean value, standard error, and "worst" value (mean of 3 largest measurements).
+As shown the correlation plot below, these three measurements are all correlated to each other, but still hold a strong relationship with the diagnosis response variable.
+The relation between the mean value and "worst" value is almost completely correlated which makes sense as both are averages however, the "worst" value only uses the three largest measurements.
+This is important to know when trying to interpret feature effect, importance and selection.
 
@@
 
+From the results of our classification analysis, we are confident not only in our predictions of tumor diagnosis, but reducing the amount of false-negative diagnoses.
+Through our methods of tuning specifically for Type II error, and lowering the prediction cutoff for malignant tumors, we have shown two successful ways of battling classification data with a preffered error class.
+This is extremely important in cancer diagnosis, as well as other analyses, as we do not want to tell someone they don't have cancer when they do, as that could delay treatment and further risk their life.
+
+In terms of feature importance and feature selection, there were many limitations in interpretations due to the correlation of the measurement types.
+We did find that there were a number of the same features outlined by different models, so we can conclude that the variables, regardless of the measurement type, hold importance in classifying a tumor diagnosis.
+
 # Potential Improvements

### 2022-12-15 paper.rmd

@@
 
-# Introduction (Should we do an Abstract?)
+# Introduction
+
+## Abstract
+
+According to the American Cancer Society breast cancer accounts for approximately 30% of female cancer diagnoses in the United States each year.
+They also state that they average risk of a woman in the United States developing breast cancer at some point in their life is around 13%.
+This means that about 1 in 8 women will get breast cancer at some point in their lifetime.
+Although breast cancer is not viewed as terminal as other types of cancer, women in the United States have about a 2.5% of dying from breast cancer.
+While breast cancer is treatable, it is very important to recognize the tumor early on in order to treat it effectively.
 
@@
 
+- American Cancer Society Breast Cancer Facts:
+    - https://www.cancer.org/cancer/breast-cancer/about/how-common-is-breast-cancer.html
 - Breast Cancer Wisconsin (Diagnostic) Data Set:

### 2022-12-15 powerpoint.rmd

@@
 ### Random Forest
-set.seed(445)
+set.seed(420)
 rf_spec <- rand_forest(mtry = sqrt(.cols())) |>
